### 2025-03-29 Paper/Paper_draft1.tex
add objects

\documentclass[12pt]{article}

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{papercitations.bib}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}
\usepackage{subcaption}
\usepackage{caption}

\onehalfspacing
\textwidth 6.5in \oddsidemargin 0in \evensidemargin -0.6in
\textheight 8.5in \topmargin -0.2in

\newcolumntype{L}[1]{>{\raggedright\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{P}[1]{>{\raggedright\tabularxbackslash}p{#1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\captionsetup{justification=centering,singlelinecheck=false}

\newcommand{\xsub}[1]{%
	\mbox{\scriptsize\begin{tabular}{@{}c@{}}#1\end{tabular}}%
}

%\renewcommand{\thetable}{\Roman{table}}

\begin{document}

	\linespread{1.2}\title{\vspace{-0.5in} Common Ownership in the Nonprofit Sector:\newline Evidence from the Hospital Industry}

	\date{\today}

	\author{\vspace{10mm}Hanna Glenn\footnote{Department of Economics, Emory University, 1602 Fishburne Drive, Atlanta, GA 30322, hanna.glenn@emory.edu.} }

	\maketitle
	%\setlength{\droptitle}{-10pt}

	\vspace{-0.2in}

	\singlespacing\maketitle

 \vspace{3mm}

    \begin{abstract}
		{\small

		} 
	\end{abstract}

	\onehalfspacing

	\newpage

    \section{Introduction}

    As consolidation in investment companies and in many industries has increased over time, so has the prevalence of common ownership among publicly traded firms. That is, firms within the same industry and market are increasingly owned by the same investment firms. Whether this type of common ownership affects competition is a widely debated theme in the economics literature. Theoretical models predict that common ownership will lead to lower equilibrium output and higher markups through firms placing non-zero weight on competitors' profit (\cite{rubinstein1983competitive}; \cite{rotemberg1984financial}; \cite{azar2012new}). Empirical evidence yields mixed findings, with various studies documenting less competition due to common ownership (\cite{he2017product}; \cite{azar2018anticompetitive}; \cite{azar2022ultimate}; \cite{newham2018common}), but several recent studies that carefully consider identification find either no effect or increased competition due to common ownership (\cite{chen2023does}; \cite{kini2024common}). Additionally, recent literature has considered common ownership in the context of private firms with venture capitalists, where consolidated ownership is even more prevalent, and has found that competition decreases and information sharing increases in this setting (\cite{lindsey2008blurring}; \cite{gonzalez2020exchanges}; \cite{li2023common}; \cite{eldar2024common}). I extend this theme of common ownership and firm behavior to a setting which exhibits similar patterns of affiliation with competing firms in a way that has not yet been considered: board of directors of nonprofits.

    Unlike publicly traded firms, nonprofits do not have shareholders that receive dividend payments. Instead, the main governing body is a board of directors: a group of volunteers responsible for selecting executives and holding them accountable. In some ways, a board functions similarly to shareholders; they influence the strategic direction of the organization through information sharing and voting on firm decisions without actively managing day-to-day operations. As with investor common ownership, there are no legal restrictions preventing directors from serving on multiple boards, even if those firms are competing in the same industry and market. Therefore, if board members are affiliated with multiple firms and they have the power to influence decisions of the firm, such affiliation might affect firm decisions on product design, output, research and development, and other factors that affect consumers.

    In this paper, I focus specifically on board affiliations among nonprofit hospitals in the US. Nonprofit hospitals play a dominant role in US health care. As of 2023, 50\% of hospitals were private nonprofits, compared to 36\% for-profit. Moreover, nonprofit hospitals operate larger facilities, with an average of 207 staffed beds, compared to 107 in for-profits (\cite{ASPE_2023}). Therefore, nonprofit hospital behavior directly impacts the patient population in the US. Additionally, hospital mergers and consolidations are under increasing scrutiny as competition in the sector has declined significantly in recent years (\cite{levinson2024ten}). It may be that hospital affiliations through board of directors provides a means of informal affiliation that has the power to raise profits but doesn't raise anti-trust concerns. In light of this, the purpose of this paper is two-fold. First, I document the extent of board-level affiliation among plausibly independent nonprofit hospitals, capturing a nonprofit equivalent to common ownership. Second, exploiting timing in board connections formed or lost between hospitals, I explore several potential effects of board affiliations on hospital behavior. While many hospital behaviors are potentially relevant to board member affiliations, I focus specifically on how hospitals with shared board members alter their service offerings and specialization. Board members likely influence hospital decisions through knowledge sharing, which can shape strategic choices about which services to provide or which patient populations to target. This mechanism is particularly relevant if there is asymmetric information about competitors, as hospitals may gain insights from shared board members.

    To document shared board members of nonprofit hospitals in the US, I compile information on hospital board of director members from publicly available Tax Form 990s, filed each year by nonprofits in the US. One section of this form requires declaration of the names of each member of the board, executives, and highest compensated employees. Extracting these identities and matching to other sources of hospital information, I form a network of hospital board members from 2017-2022. I show that nearly 10\% of the nonprofit hospitals in the sample are affiliated with another hospital through shared board members, even excluding system or network affiliations. Two-thirds of affiliated pairs are made up of two general hospitals, with the remaining pairs consisting of a general hospital affiliated with a specialty hospital.

    I then construct measures of service offerings and patient populations from two sources of data: the American Hospital Association (AHA) survey, which records the number of beds devoted to a large range of services in each year, and the Center for Medicaid and Medicare Services (CMS) Provider Utilization Files, which record the percent of Medicare patients treated with certain conditions, along with aggregate demographics of Medicare patients seen in the hospital. I construct measures of the concentration of services provided (both in terms of bed allocation and patient percentages) at the hospital-year level using a Herfindahl-Herschmann index calculation, where a highly concentrated hospital specializes in a smaller number of services. I show descriptively that general hospitals unaffiliated through board members have a constant service offering concentration over time, but hospitals with board member affiliations have an increasing service concentration over time.

    To gain a better understanding of the causal effects of shared board member affiliations, I employ a difference-in-differences strategy, comparing hospitals that become affiliated through board members to similar hospitals that do not become affiliated before and after the affiliation begins. Specifically, I employ four different specifications leveraging (1) whether a hospital loses or gains board affiliation with another hospital, and (2) comparison groups of hospitals that are unaffiliated through board, but either belong to a hospital system or not. I estimate the effect of losing or gaining board affiliation on four outcomes: the concentration indexes described above, the total number of Medicare patients seen, and the average risk score of Medicare patients seen. While concentration and risk scores do not seem to affected by board affiliations, the total number of Medicare patients seen increases by 250 patients per year in the second year after becoming affiliated through board members. This indicates potential patient redistribution, possibly through referrals, when more information is gained about competing hospitals.

    This finding has important implications for hospital competition and patient access to care. If board affiliations facilitate patient redistribution through improved information flow, policymakers should consider how governance structures influence market dynamics. Increased patient volume at affiliated hospitals could improve efficiency and resource allocation, but it may also raise concerns about reduced competition if patient referrals become concentrated within certain networks. To ensure that these affiliations benefit patients without distorting competition, regulators could enhance transparency around board connections and referral practices.

    This study contributes to the literature on common ownership and firm behavior by examining how shared board affiliations among nonprofit hospitals influence competition and patient distribution. While prior work has largely focused on common ownership in publicly traded firms and venture capital-backed private firms, I extend this line of inquiry to nonprofit hospitals, where governance structures differ but affiliation mechanisms may still shape competitive dynamics. By leveraging a staggered difference-in-differences design, I provide causal evidence on how board affiliations affect hospital service offerings and patient flows, offering novel insights into the role of corporate governance in healthcare markets. [REWRITE THIS TO BE BETTER LATER]

    \section{Data}\label{sec:data}

   I compile data on nonprofit hospital board of directors from publicly available Tax Form 990s, which all sufficiently large nonprofits must file with the Internal Revenue Service (IRS) each year. These forms contain a section in which firms declare their board, executives, and highest compensated employees. I limit to firms that have a completed Schedule H in the 990, indicating that the firm runs a hospital. PDF version of the 990s are made public by Nonprofit Explorer dating back earlier than 2009. However, using optical character recognition text extraction algorithms on these PDFs is both time consuming and can lead to small errors in names extracted. Thus, I analyze the data from 2017 onward, when the IRS began publishing XML versions of these files, ensuring more accurate names that can be matched across hospitals and time. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics. There are 2,096 hospitals with information extracted directly from the XML files.

   To observe hospital outcomes regarding service offerings, I must match hospitals in the tax form data to hospitals in other publicly available data sources. The only common information about hospitals across these sources is the name and location. Therefore, I use fuzzy string matching methods to create a crosswalk of EINs to identification numbers in the AHA data. Specifically, I compute the Jaro-Winkler distance between hospital names in each data set. Then, I record matches with a sufficiently small distance and sufficiently similar address. Using this method, I have a sample of 1,623 hospitals in the tax data that can be linked to the AHA survey. From the AHA survey, I gather demographic information about hospitals as well as the allocation of their beds to specific services. Additionally, I merge the CMS Provider Utilization files, which provides hospital-level information about the types of Medicare patients seen in each year, including percent of patients of a certain race or age, and the percent of patients treated for specific conditions.

    Next, I clean the column of board member names by removing common titles (Dr., Reverand, etc.), and combining names within the same firm that have slight spelling differences. For example, if John Matthews is a board member of hospital A in 2017, and John Mathews is a board member of hospital A in 2018, I assume these names represent the same person. Additionally, I combine names within the same HRR that are identical apart from small mis-characters such as the presence of a middle initial. Finally, to avoid issues in whether first or last names are listed first, I sort each individual's names alphabetically. I define a hospital as being affiliated if the same individual is a board member of two otherwise unaffiliated hospitals. That is, two hospitals that do not belong to the same system or network can be affiliated through board members, but same-system or same-network hospitals can not. I make this distinction to separate the variation in affiliation by board member alone versus simply having shared board members due to being in the same system, which is plausibly already correlated with service offerings. I also focus on affiliations within the same hospital market, as the outcomes of interest focus on competing for patients in a specific geographic region. I measure hospital markets using Hospital Referral Regions (HRRs) as defined by Dartmouth Atlas of Healthcare.

    \import{Objects}{all_summarystats.tex}

    \section{Documenting Board of Director Affiliation}

    Of the approximately 1400 hospitals in the final sample, approximately 7\% are affiliated with another hospital through a shared board member. As shown in Figure \ref{fig:connected_percent}, this percentage remains relatively constant over time, decreasing from 132 affiliated hospitals in 2017 to 117 affiliated hospitals in 2021.

    \begin{figure}[ht!]
        \centering
        \caption{Percent of hospitals that have a board affiliation}
        \includegraphics[width=.8\textwidth]{Objects/connected_percent.pdf}
        \label{fig:connected_percent}
    \end{figure}

    As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population. Of the 306 HRRs in the US, approximately 17\% have at least one pair of board affiliated hospitals within the market. As shown in Figure \ref{fig:connected_HRR_percent}, this proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation.

    \begin{figure}[ht!]
        \centering
        \caption{Percent of HRRs containing board affiliated hospitals}
        \includegraphics[width=.8\textwidth]{Objects/connected_HRR_percent.pdf}
        \label{fig:connected_HRR_percent}
    \end{figure}

    The geographic distribution of these pairs is shown in Figure \ref{fig:connected_maps}, where the blue dots represent hospitals that share a common board member. Connected hospitals are distributed fairly regularly across the US, apart from relatively few connections on the west coast relative to the population. Over time, the connections remain fairly consistent geographically.

    \begin{figure}[ht!]
        \centering
        \caption{Geographic distribution of board affiliated hospitals over time}
        \includegraphics[width=.8\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    Next, I show the proportion of the type of hospital affiliations in Table \ref{tab:hospital_pair_types}. Each hospital is general or specialty, children or adult, and independent or system, all characteristics found in the AHA data. Two-thirds of board affiliated pairs are between two general hospitals, and 10\% of pairs are between a specialty and a general hospital. In the main analysis, I focus only on the affiliations between general hospitals, as this is the most prominent pair type and the most direct competition. 
    There is also significant variation in the ownership type of pairs. Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system. In total, there are 217 hospitals that are affiliated with at least one other hospital through shared board members at some point in the sample.

    \import{Objects}{hospital_pair_types.tex}

    I now shift towards examining summary statistics of hospitals with different types of affiliation relationships. First, in Table \ref{tab:hospital_general_summarystats}, I focus on general hospitals in four categories: affiliated through shared board members with another general hospital, affiliated through shared board members with a specialty hospital, unaffiliated through board members but belonging to a system, and completely unaffiliated. I present means of hospitals in each of these categories for the number of beds, the number of patients seen, whether it is an academic medical center, and whether the hospital is located in a metropolitan area. On these dimensions, hospitals affiliated with other general hospitals (column 1) are relatively similar to unaffiliated hospitals that are part of a system (column 3). Unaffiliated hospitals that are not part of a system are slightly smaller in terms of bed size and number of patients seen, and less likely to be in a metropolitan area. The hospitals in column (2) are outliers, with a large number of beds and patients seen, and mainly in metropolitan areas. There are only 12 hospitals in this sample that are likely very different from the average general hospital.

    \import{Objects}{hospital_general_summarystats.tex}

    I also present means of variables to capture aspects of the service offerings of the different types of hospitals. First, from the AHA survey, I include an indicator for whether the hospital offers services in a Neonatal Intensive Care Unit (NICU) or Cardiac catheterization lab, both specialized and profitable services. General hospitals that are affiliated with specialty hospitals are the most likely to offer these services. Hospitals without any board member affiliations but belonging to a system are more likely to offer these services than unaffiliated hospitals that do not belong to a system. Hospitals that are affiliated with another general hospitals fall somewhere in between these two, with slightly less likelihood than the unaffiliated hospitals belonging to a system. I also include a measure of the concentration of services offered by the hospital using a Herfindahl-Hirschman Index calculation of the number of beds devoted to each service category in the AHA data. Specifically, for each hospital, I compute the proportion of total beds assigned to each service category and then square these proportions before summing them. A higher HHI value indicates that a hospital's bed capacity is concentrated in fewer service categories, whereas a lower HHI suggests a more diversified service offering, which provides insight into the degree of specialization or diversification in a hospital’s service provision. I also compute a similar concentration metric using the percent of patients in several service categories from the CMS Provider Utilization Files, including cancer, Chronic Obstructive Pulmonary Disease (COPD), kidney, and heart-related services. For both measures, there seems to be little variation in service concentration across affiliation types.

    Next, I present the same summary statistics for specialty hospitals with different types of affiliations: affiliated with a general hospital, unaffiliated through board but belonging to a system, and unaffiliated through board or system, shown in Table \ref{tab:hospital_specialty_summarystats}. Overall, there are less specialty hospitals included in the sample, and very few of them are affiliated through board members. Specialty hospitals are more concentrated than general hospitals, as expected. In the AHA measure of concentration, all specialty hospitals are similarly concentrated. However, when using the services included in the CMS files, the concentration of unaffiliated system hospitals is much lower than that of any other specialty hospital. This could be due to the limited services included in the CMS data.

    \import{Objects}{hospital_specialty_summarystats.tex}

    In Figure \ref{fig:concentration_services_time}, I show the AHA measure of service concentration based on the number of beds over time for general hospitals with different levels of affiliation. General hospitals that are unaffiliated through board members, whether they are affiliated with a system or not, remain relatively constant in their service offerings over time. However, general hospitals that are affiliated with another general hospital through a shared board member display increasing service concentration over time, suggesting that there may be some correlation between board affiliations and the services offered in hospitals.

    \begin{figure}[ht!]
        \centering
        \caption{Concentration of services offered over time}
        \includegraphics[width=0.8\textwidth]{Objects/concentration_services_time.pdf}
        \label{fig:concentration_services_time}
    \end{figure}

    \section{Effect of Board Affiliation on Hospital Services and Patients}

    I estimate the average effect of either losing or gaining board affiliation on the outcomes discussed in Section \ref{sec:data} using a staggered timing difference in differences strategy. For simplicity, I only consider general hospitals in both the treatment and control groups, as they are the majority of connections and the most closely related competitors within a market. Further, I leverage the variation in different types of affiliation, whether via board members or system ownership, by defining separate control groups of unaffiliated hospitals into system or non-system categories. Therefore, I define four relevant subsets of hospitals. First, I construct two groups of ``treated" hospitals, where treatment is defined as either losing an established board affiliation for the remainder of the time period, or gaining a shared board affiliation that remains for the rest of the time period. After dropping hospitals that lose or gain affiliation multiple times, there are 112 hospitals who are affiliated at some point, 59 0f which lose affiliation at some point and 53 of which gain affiliation at some point. Then, I define two ``control" groups of hospitals, those that are a part of a formal hospital system or not.

    Since hospitals can lose or gain affiliated board members at different times, negative weighting issues can arise when employing a classic two way fixed effects specification. Thus, I estimate average treatment effects for a specific group $g$ at time $t$:

    $$ATT(g,t)=\mathbbm{E}[Y_t(g)-Y_t(0)|G_g=1],$$

    \noindent where $G_g=1$ for those in group $g$. A group represents hospitals that are treated, defined as gaining or losing board affiliation, in the same year. I employ the doubly robust estimator established in \citeauthor{sant2020doubly} (\citeyear{sant2020doubly}). I aggregate the group-time estimates into a more conventional event study plot, incorporating simultaneous confidence bands derived from bootstrapped standard errors.

    There are several necessary assumptions to identify the parameters of interest, $ATT(g,t)$. First, I assume that once a hospital gains or loses board affiliation with another hospital, this affiliation is not reversed. I ensure this mechanically by only including hospitals in the treated sample who establish or lose a board affiliation once for the entire time period. Second, I assume there is no anticipation of treatment. That is, hospitals do not change behavior based on the expectation of gaining or losing board affiliation. As is usual in a difference-in-differences framework, I also assume a version of parallel trends based on never-treated units. I assume that average outcomes for hospitals treated in group $g$ would have followed a parallel trend as hospitals that were never board affiliated in the absence of treatment.

    \subsection{Concentration of Bed Allocation}

    In Figure \ref{fig:aha_hhi_did}, I present the estimated effect of either losing board affiliation or gaining board affiliation with another hospital on the concentration of bed allocation. As described in Section \ref{sec:data}, I construct this concentration measure using the number of beds a hospital allocates towards different services. While the magnitude itself does not have a direct interpretation, relative magnitudes indicate more or less specialized hospitals.

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Board Affiliation on Concentration of Bed Allocation}
        \includegraphics[width=0.75\textwidth]{Objects/aha_hhi_did.pdf}
        \label{fig:aha_hhi_did}
    \end{figure}

    The top panel of Figure \ref{fig:aha_hhi_did} compares hospitals who lose a board affiliation at some point in time with hospitals who never had a board affiliation, and the bottom panel compares hospitals who gain a board affiliation to hospitals that never have an affiliation. Relative time $t=0$ is the point at which shared board members were gained or lost. In yellow, the comparison group contains hospitals that do not belong to a system. In gray, the comparison group contains hospitals that do belong to a system but are unaffiliated by board members outside of their system. In all specifications, there is no statistically significant difference between board affiliated and non-board-affiliated hospitals in terms of how specialized their bed allocation is. It could be that this measure of service offerings is sticky in the short run, and that hospitals do not make drastic changes to their bed allocations. However, even with constant bed allocations, the types of patients that a hospital treats might change in the short run.

    \subsection{Concentration of Patient Treatment}

    A more adjustable measure in the short run could be patient-driven specialization as opposed to bed allocation. The type of patients seen in the hospital can be altered by referral patterns, patient demand, or advertising. Therefore, I now present the effect of losing or gaining board affiliation on the concentration of patients seen for specific conditions. The calculation is similar to how I constructed the bed allocation measure, although I now calculate the HHI index using the share of patients treated for various conditions, as described in Section \ref{sec:data}. I present the estimated average treatment effects in Figure \ref{fig:cms_hhi_did}.

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Board Affiliation on Concentration of Patient Treatment}
        \includegraphics[width=0.75\textwidth]{Objects/cms_hhi_did.pdf}
        \label{fig:cms_hhi_did}
    \end{figure}

    I compare hospitals who lose board affiliation to those who were never affiliated in the top panel of Figure \ref{fig:cms_hhi_did}. There is slight evidence that hospitals who lose board affiliation have relatively more concentrated patient services than unaffiliated hospitals, although the magnitude is quite small relative to the mean patient concentration of around .5. Additionally, there is no effect of gaining board affiliation on patient treatment concentration, shown in the bottom panel of Figure \ref{fig:cms_hhi_did}. These results, taken with the null effect on the bed allocation concentration, indicate that the services offered to patients do not change much as a result of shared board members.

    \subsection{Medicare and Medicaid Patients}

    A common outcome studied in the common ownership literature is output. In hospitals, output is roughly equivalent to the patients treated in the hospital. However, there are differences in the types of patients a hospital can treat. Private patients are the most profitable, and therefore the most desirable to hospitals, while publicly insured patients are often seen as less profitable. Thus, board affiliations might affect the flow of different types of patients seen in different hospitals. For example, a hospital might direct patient referrals to a board-affiliated hospital over a hospital they have no affiliation with. Therefore, I consider three outcomes of patient composition that I observe in the AHA or CMS data. First, I show the effect of losing or gaining common ownership on the total number of Medicare patients seen in the hospital from the CMS Provider Utilization Files in Figure \ref{fig:benes_did}.

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Board Affiliation on Number of Medicare Patients}
        \includegraphics[width=0.75\textwidth]{Objects/benes_did.pdf}
        \label{fig:benes_did}
    \end{figure}

    The top panel of Figure \ref{fig:benes_did} shows the relative change of the number of Medicare patients when a hospital loses board affiliation. In yellow, the comparison group consists of non-system hospitals. In gray, the comparison group consists of hospitals belonging to systems. There is a slight increase in Medicare patients for hospitals losing affiliation compared to system hospitals, but this magnitude is small and only persists for one year. However, the bottom panel shows that hospitals that gain a board affiliation have an increased number of Medicare patients compared to non-system, non-board-affiliated hospitals in both the second and third year after becoming board affiliated. This magnitude of approximately 250 additional Medicare patients is meaningful compared to the average number of Medicare patients, approximately 3,000. This increase indicates a potential change in patient volume or patient composition once a hospital becomes affiliated through board members.

    To dig into this a little deeper, I also estimate the effect of gaining or losing board affiliation on Medicare and Medicaid discharges relative to the total number of beds in the hospital.

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Board Affiliation on Medicare/Bed}
        \includegraphics[width=0.75\textwidth]{Objects/mcare_did.pdf}
        \label{fig:mcare_did}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Board Affiliation on Medicaid/Bed}
        \includegraphics[width=0.75\textwidth]{Objects/mcaid_did.pdf}
        \label{fig:mcaid_did}
    \end{figure}

    While not a perfect measure of patient volume, this measure will give an indication of whether total volume or the types of patients a hospital admits might change due to board affiliation.

    \subsection{Risk Score of Medicare Patients}

    Finally, I present estimates of how board affiliation affects the average risk score of patients in the hospital in Figure \ref{fig:risk_did}. A higher risk score indicates that patients seen in that hospital have, on average, more complex conditions. This outcome might indicate whether patient composition changes as a result of board affiliation.

    \begin{figure}[ht!]
        \centering
        \caption{Effect of Board Affiliation on Risk Score}
        \includegraphics[width=0.75\textwidth]{Objects/risk_did.pdf}
        \label{fig:risk_did}
    \end{figure}

    Again, the top panel of Figure \ref{fig:risk_did} includes the sample of hospitals who lose their board affiliation at some point as the treated group and hospitals who were never board affiliated as the comparison group. The comparison group is also divided into those who belong to a system, in gray, and those who do not belong to a system, in yellow. Hospitals that lose their board affiliation have a slight increase in the risk score of their Medicare patients after losing board affiliation relative to both types of comparison hospitals. For non-system hospitals, this happens in the first year after losing affiliation, whereas for system hospitals this happens in the second year. However, this increase in risk score might indicate that losing affiliation affects the patients coming through the hospital, potentially through referrals. Interestingly, the bottom panel of Figure \ref{fig:risk_did} shows that hospitals who gain board affiliation also have an increased risk score relative to the comparison non-system hospitals. [EXPLANATION??]

    \newpage

    \printbibliography

\end{document}

### 2025-03-29 Scripts/Read990_scheduleH.R

@@ -1,0 +1,130 @@
+library(xml2)
+library(XML)
+library(dplyr)
+library(tidyr)
+library(purrr)
+library(stringr)
+
+# write a script to loop through all xml files published by the IRS
+# Here, I will check whether it's a hospital by checking if there is a Schedule H section
+
+replace_null <- function(x) {
+  x <- purrr::map(x, ~replace(.x, is.null(.x), NA_character_))
+  purrr::map(x, ~if(is.list(.x)) replace_null(.x) else .x)
+}
+
+# make a list of all folders to search through
+folders <- list.files(path = paste0(drive_path, "IRS_XML_990s/"), full.names = FALSE)
+
+# create empty data frames to store data
+all_ein_data <- data.frame()
+all_people_data <- data.frame()
+
+# read in temp data sets
+all_ein_data <- readRDS(paste0(created_data_path, "all_ein_data_scheduleH.rds"))
+all_people_data <- readRDS(paste0(created_data_path, "all_people_data_scheduleH.rds"))
+
+# loop through each folder
+for (folder in folders[25:75]) {
+  # make a list of all files in the folder
+  files <- list.files(path = paste0(drive_path, "IRS_XML_990s/",folder), pattern = ".xml", full.names = FALSE)
+  
+  # loop through files
+  for (file in files){
+    # read in the xml file
+    # See if the file is non-empty
+    if (file.size(paste0(drive_path, "IRS_XML_990s/",folder,"/",file)) > 0){
+      xml <- xmlParse(paste0(drive_path, "IRS_XML_990s/",folder,"/",file))
+      xml_list <- xmlToList(xml, addAttributes = TRUE)
+      
+      # Return TRUE if there is a Schedule H in the return
+      hosp <- ifelse("IRS990ScheduleH" %in% names(xml_list[["ReturnData"]]),1,0)
+    
+      # check if the ein is in the list of eins
+      if (hosp==1){
+        ein <- xml_list[["ReturnHeader"]][["Filer"]][["EIN"]]
+        # convert to list format
+        xml_list <- xmlToList(xml, addAttributes = TRUE)
+        # fill all missing values with NA
+        xml_list <- replace_null(xml_list)
+        
+        header <- as.data.frame(xml_list[["ReturnHeader"]])
+        
+        data <- list_flatten(xml_list[["ReturnData"]], name_repair = "minimal")
+        data <- list_flatten(data, name_repair = "minimal")
[…]

### 2025-03-29 Scripts/clean_matched_names.R

@@ -1,0 +1,366 @@
+library(readr)
+library(stringr)
+library(dplyr)
+
+created_data_path <- "CreatedData/"
+
+# Script to clean up the name columns of people data scraped from 990s
+
+all_people_data <- readRDS(paste0(created_data_path, "/all_people_data_scheduleH.rds"))
+cw <- readRDS(paste0(created_data_path, "/updated_ein_aha_cw.rds"))
+
+cw <- cw %>%
+  filter(!is.na(ID))
+
+# only keep observations where EIN is found in either ein_hosp or ein_sys
+all_people_data <- all_people_data %>%
+  filter(Filer.EIN %in% cw$Filer.EIN)
+
+# look at unique values of section
+unique(all_people_data$Section)
+
+# only keep relevant sections
+all_people_data <- all_people_data %>%
+  filter(Section %in% c("Form990PartVIISectionAGrp", "RltdOrgOfficerTrstKeyEmplGrp"))
+
+# fill personNm with other lines
+all_people_data <- all_people_data %>%
+  mutate(PersonNm = ifelse(is.na(PersonNm), BusinessName_BusinessNameLine1Txt, PersonNm)) %>%
+  filter(!is.na(PersonNm) & PersonNm != "'")
+  # this fills all missing person names
+
+# Look at missing titles
+observe <- all_people_data %>%
+  filter(is.na(TitleTxt))
+
+# remove observations with no title text
+all_people_data <- all_people_data %>%
+  filter(!is.na(TitleTxt))
+
+# tabulate the number of unique EINs for every value of TaxYr
+all_people_data %>%
+  group_by(TaxYr) %>%
+  summarise(n = n_distinct(Filer.EIN))
+
+# First goal is to clean up the names. ####
+
+# create data set with only relevant variables
+people_data <- all_people_data %>%
+  select(TaxYr, Filer.EIN, PersonNm, TitleTxt)
+
+# convert to all lowercase
+people_data <- people_data %>%
+  mutate(PersonNm = tolower(PersonNm),
+         TitleTxt = tolower(TitleTxt))
+
+# create a column for md
+people_data <- people_data %>%
+  mutate(md = ifelse(str_detect(PersonNm, "\\smd$|\\smd\\s"),1,0)) %>%
+  mutate(md = ifelse(str_detect(TitleTxt, "\\smd$|\\smd\\s"),1,md))
+
[…]

### 2025-03-29 Scripts/descriptive_tables_and_figures.R

@@ -1,0 +1,344 @@
+# load libraries 
+library(dplyr)
+library(fuzzyjoin)
+library(stringdist)
+library(tidyr)
+library(stringr)
+library(ggplot2)
+library(patchwork)
+library(readr)
+library(data.table)
+library(kableExtra)
+library(purrr)
+library(ggpubr)
+library(did)
+
+created_data_path <- "CreatedData/"
+
+
+## SCRIPT CREATING SUMMARY STATISTICS FOR THIRD CHAPTER ################
+
+# Part 1: Clean the Data ---------------------------------------------------------
+# Read in cleaned up version of names found in 990 tax forms
+people <- read_rds(paste0(created_data_path, "cleaned_people_data.rds"))
+# Read is crosswalk of AHA ID - EIN, only keeping those who match to a standalone hospital
+cw <- readRDS(paste0(created_data_path, "updated_ein_aha_cw.rds")) 
+
+
+
+# only keep people whose EIN is in the crosswalk as a standalone hospital
+people <- people %>% filter(Filer.EIN %in% cw$Filer.EIN)
+
+# only keep people who are board members
+board_people <- people %>%
+  filter(board_member==1) %>%
+  select(TaxYr, Filer.EIN, name_cleaned)
+
+# Only keep EINs present in 2017-2021
+ein_keep <- board_people %>%
+  distinct(TaxYr, Filer.EIN) %>%
+  mutate(count = ifelse(TaxYr %in% 2017:2021, 1, 0)) %>%
+  group_by(Filer.EIN) %>%
+  summarise(count = sum(count)) %>%
+  filter(count>=5)
+
+board_people <- board_people %>%
+  filter(Filer.EIN %in% ein_keep$Filer.EIN)
+
+rm(ein_keep)
+
+# clean up extras in names
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "deceased")
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\seff$")
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\scsjp$")
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\scisa$")
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "^mrs\\s")
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "^reverend\\s")
+board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\scns$")
[…]

### 2025-03-29 Scripts/function1_standardize_names.R

@@ -1,0 +1,81 @@
+standardize_names_optimized <- function(data, max_dist) {
+  setDT(data)  # Convert to data.table for efficiency
+  
+  # Split by EIN to avoid excessive joins
+  firm_list <- split(data, by = "Filer.EIN", keep.by = TRUE)
+  
+  # Function to process each firm's data
+  process_firm <- function(firm_data) {
+    if (nrow(firm_data) <= 1) return(firm_data)  # Skip firms with only 1 name
+    
+    unique_names <- unique(firm_data[, .(name_cleaned)])  # Reduce size
+    
+    # Compute approximate name matches only within this firm
+    matches <- stringdist_inner_join(
+      unique_names, unique_names, 
+      by = "name_cleaned", 
+      max_dist = max_dist, 
+      method = "osa"
+    ) %>%
+      filter(name_cleaned.x < name_cleaned.y) %>%
+      group_by(name_cleaned.y) %>%
+      summarise(correct_name = first(name_cleaned.x), .groups = "drop")
+    
+    # Merge corrected names
+    setDT(matches)
+    firm_data <- merge(firm_data, matches, by.x = "name_cleaned", by.y = "name_cleaned.y", all.x = TRUE)
+    firm_data[, name_cleaned := fcoalesce(correct_name, name_cleaned)]
+    firm_data[, correct_name := NULL]
+    
+    return(firm_data)
+  }
+  
+  # Process firms in parallel (reduces RAM usage)
+  corrected_list <- lapply(firm_list, process_firm)
+  
+  # Recombine data
+  corrected_data <- rbindlist(corrected_list, use.names = TRUE, fill = TRUE)
+  
+  return(corrected_data)
+}
+
+standardize_names_by_hrr <- function(data, max_dist) {
+  setDT(data)  # Convert to data.table for efficiency
+  
+  # Split data by HRRCODE (instead of Filer.EIN)
+  hrr_list <- split(data, by = "HRRCODE", keep.by = TRUE)
+  
+  # Function to process each HRR group
+  process_hrr <- function(hrr_data) {
+    if (nrow(hrr_data) <= 1) return(hrr_data)  # Skip HRRs with only 1 name
+    
+    unique_names <- unique(hrr_data[, .(name_cleaned)])  # Reduce size
+    
+    # Compute approximate name matches only within this HRR
+    matches <- stringdist_inner_join(
+      unique_names, unique_names, 
[…]

### 2025-03-29 Scripts/function2_identify_common_members.R

@@ -1,0 +1,40 @@
+identify_common_members <- function(data) {
+  # Collect all EINs for each name within the same year
+  ein_mapping <- data %>%
+    group_by(name_cleaned, TaxYr) %>%
+    summarise(all_eins = list(unique(Filer.EIN)), .groups = "drop")
+  
+  # Merge back and correctly compute other EINs within the same year
+  df_with_others <- data %>%
+    left_join(ein_mapping, by = c("name_cleaned", "TaxYr")) %>%
+    rowwise() %>%
+    mutate(other_eins = paste(setdiff(unlist(all_eins), Filer.EIN), collapse = ", ")) %>%
+    ungroup() %>%
+    select(-all_eins)  # Drop intermediate column
+  
+  return(df_with_others)
+}
+
+identify_common_members_ceo <- function(data) {
+  # Collect all EINs and positions for each name within the same year
+  ein_mapping <- data %>%
+    group_by(name_cleaned, TaxYr) %>%
+    summarise(
+      all_eins = list(unique(Filer.EIN)),
+      positions = list(unique(position)),  # Store unique positions
+      .groups = "drop"
+    )
+  
+  # Merge back and correctly compute other EINs within the same year
+  df_with_others <- data %>%
+    left_join(ein_mapping, by = c("name_cleaned", "TaxYr")) %>%
+    rowwise() %>%
+    mutate(
+      other_eins = paste(setdiff(unlist(all_eins), Filer.EIN), collapse = ", "),
+      other_positions = ifelse(other_eins == "", NA, paste(unique(unlist(positions)), collapse = ", "))
+    ) %>%
+    ungroup() %>%
+    select(-all_eins, -positions)  # Drop intermediate columns
+  
+  return(df_with_others)
+}

### 2025-03-29 Scripts/updated_match_hosp.R

@@ -1,0 +1,556 @@
+## Match AHA IDs to tax identifiers
+
+library(dplyr)
+library(stringr)
+library(readr)
+library(cdlTools)
+
+# Read in EIN data from XMLs
+all_ein_data <- readRDS("CreatedData/all_ein_data_scheduleH.rds")
+
+# How many EINs are in this data?
+length(unique(all_ein_data$Filer.EIN))
+
+# Get rid of EINs that don't appear from at least 2018-2021
+all_ein_data <- all_ein_data %>%
+  mutate(present = ifelse(TaxYr>=2018 & TaxYr<=2021, 1, 0)) %>%
+  group_by(Filer.EIN) %>%
+  filter(sum(present)>3) %>%
+  ungroup()
+
+# Combine the variables that end with "BusinessNameLine1Txt"
+all_ein_data <- all_ein_data %>%
+  mutate(BusinessName1 = Filer.BusinessName.BusinessNameLine1Txt) %>%
+  mutate(BusinessName1 = ifelse(is.na(BusinessName1), Filer.BusinessNameLine1Txt, BusinessName1))
+
+# Combine Line2 of business name
+all_ein_data <- all_ein_data %>%
+  mutate(BusinessName2 = Filer.BusinessName.BusinessNameLine2Txt)
+
+# Change businessname to all caps
+all_ein_data <- all_ein_data %>%
+  mutate(BusinessName1 = toupper(BusinessName1),
+         BusinessName2 = toupper(BusinessName2))
+
+# Select and rename relevant variables
+# keep all variables with USAddress in the variable name
+ein_data <- all_ein_data %>%
+  select(TaxYr, Filer.EIN, BusinessName1, BusinessName2, Filer.USAddress.AddressLine1Txt, Filer.USAddress.StateAbbreviationCd,
+         Filer.USAddress.ZIPCd, Filer.USAddress.CityNm)
+
+# Work through any missing business names to determine most accurate name
+ein_data <- ein_data %>%
+  mutate(BusinessName = ifelse(is.na(BusinessName2), BusinessName1, NA))
+ein_data <- ein_data %>%
+  mutate(BusinessName = ifelse(grepl("^DBA|^D/B/A", BusinessName2), BusinessName2, BusinessName)) %>%
+  mutate(BusinessName = str_remove(BusinessName, "^DBA|^D/B/A"))
+ein_data <- ein_data %>%
[…]

### 2025-03-29 Paper/descriptive_paper.tex
Updates from Overleaf

\documentclass[12pt]{article}

\usepackage{tgtermes}
\usepackage{epsf}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}
\usepackage{dcolumn}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mwe}
\usepackage{url}
%\usepackage{harvard}
\usepackage{fancyheadings}
\usepackage{longtable}
\usepackage{authblk}
\usepackage{setspace}
%\usepackage[nomarkers]{endfloat}
\usepackage{float}
\usepackage{bbm}
%\usepackage{titling}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{import}
\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{papercitations.bib}
%\usepackage[nomarkers,nofiglist,notablist]{endfloat}
\usepackage{subcaption}
\usepackage{caption}

\onehalfspacing
\textwidth 6.5in \oddsidemargin 0in \evensidemargin -0.6in
\textheight 8.5in \topmargin -0.2in

\newcolumntype{L}[1]{>{\raggedright\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\
		\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{P}[1]{>{\raggedright\tabularxbackslash}p{#1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}

\captionsetup{justification=centering,singlelinecheck=false}

\newcommand{\xsub}[1]{%
	\mbox{\scriptsize\begin{tabular}{@{}c@{}}#1\end{tabular}}%
}

%\renewcommand{\thetable}{\Roman{table}}

\begin{document}

	\linespread{1.2}\title{\vspace{-0.5in} Overlapping Board Members in the Hospital Industry}

	\date{\today}

	\author{\vspace{10mm}Hanna Glenn\footnote{Department of Economics, Emory University, 1602 Fishburne Drive, Atlanta, GA 30322, hanna.glenn@emory.edu.} }

	\maketitle
	%\setlength{\droptitle}{-10pt}

	\vspace{-0.2in}

	\singlespacing\maketitle

 \vspace{3mm}

    \begin{abstract}
		{\small
    Consolidation in health care has been a heavily discussed topic in research and regulation, as more concentration has been linked to higher prices and lower quality. Aside from formal mergers and acquisitions, there is opportunity for hospitals to form affiliations with other competing hospitals through sharing common board members, which has only been studied in the context of large public firms. This paper establishes the existence of overlapping board members in competing nonprofit hospitals in the US. Using a data set constructed from IRS Tax Form 990s, I show that 7\% of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market. These findings highlight an understudied mechanism through which hospitals can collude, potentially affecting patient and market outcomes. 
		} 
	\end{abstract}

	\onehalfspacing

	\newpage

    \section{Introduction}

    Consolidation in health care remains a forefront concern for regulators in the US, as both horizontal and vertical integration have substantially increased over the last several decades (\cite{levinson2024ten}). Researchers have documented higher prices (\cite{brot2024pays}), and mixed effects on quality (\cite{kessler2000hospital}; \cite{cooper2011does}) as a result of health care becoming more concentrated through mergers and acquisitions. There is significant antitrust scrutiny in this sector, and proposed mergers are regularly denied due to expected harmful effects to competition (\cite{Meyer2022}). Firms in other industries have been shown to engage in methods of informal affiliation besides formal mergers and acquisitions. For example, instances of board overlap in public firms have been cited to violate the Clayton Act due to their anti-competitive nature. Yet, despite concerns of low competition, this type of connectedness among firms has not been given much consideration in health care. In this paper, I document the extent of overlapping board members in nonprofit hospitals in the US and explore avenues of future research to investigate 
    potential anti-competitive effects.

    The presence of overlapping board members across firms, also referred to as board interlock, has been heavily studied in corporate finance, particularly for large publicly traded firms. By the Clayton Act of 1914, board overlap among sufficiently large competing firms is illegal. Yet, it is known to exist even among the largest publicly traded firms. Researchers have found that board overlap is correlated with governance practices, performance, investment in research and development, and product differentiation (\cite{cai2014board}; \cite{lamb2016ties}; \cite{geng2021does}). While the Department of Justice has increased scrutiny of board overlap, their focus has remained on tech and private firms (\cite{Morse2023}). Board overlap does occur in health care with instances of organizations denied merger permission due to boards not acting independently (\cite{huberfeld2006tackling}), of hospitals declaring informal affiliations through leadership (\cite{barnett_babcock_2012}), and of life sciences companies with substantial documented overlap (\cite{manjunath2024illegal}). However, little is known about how prevalent board overlap is in the hospital industry, and whether it is correlated with other hospital characteristics.

    In this paper, I form networks of hospital board of director members using Tax Form 990s, which requires nonprofits in the US to declare the identities and compensation information of their board. Extracting these identities and matching to other sources of hospital information, I show that approximately 7\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations. [Add more results]

    Establishing descriptive evidence of board overlap in hospitals, this paper encourages future research to take seriously this behavior as potentially having downstream effects for patients. While formal integration in health care should not be overlooked, there is a call for future research to scrutinize other forms of firm allegiance in policy and regulation.

    \section{Related Literature}

    \subsection{Overlapping Board Members}\label{sec:boardoverlaplit}

    One of the first studies of board overlap argues that firms are not independent, self-sufficient organisms in the economy, but are actually heavily interconnected via overlapping board members (\cite{dooley1969interlocking}). Dooley frames board interlock as ``the root of many evils", and finds several aspects of corporations to be correlated with whether they share board members with another corporation: size, financial relationships with other companies, and management control. This paper led to a number of studies focusing on how board interlock affects corporate governance, including whether board overlap stretches individual board members too thin (\cite{ferris2003too}; \cite{field2013busy}), contributes to increases in backdating stock options (\cite{bizjak2009option}), or led to convergence in governance strategy and behavior (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}).

    The literature thus establishes a strong link between board interlock and governance practices. However, the relationship between board overlap and firm behavior, particularly its impact on competition, remains less understood. Theoretically, the direction of this effect is ambiguous. On one hand, board interlock is often viewed as an anti-competitive practice that enables implicit collusion. Alternatively, firms may seek highly experienced and connected individuals for their boards, leading to valuable knowledge sharing and enhanced decision-making. Most studies in this area have focused on the financial performance implications of board overlap, with firm behavior considered primarily as a mediating mechanism.

    Two main studies analyze of the effect of board overlap on financial performance. First, \citeauthor{baran2017director} (\citeyear{baran2017director}) leverages entry and exit from the S\&P 500, along with variation in the local supply of directors, and finds that well-connected board members contribute positively to firm value. Similarly, leveraging policies in the US that remove barriers to board overlap, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that board overlap leads to higher profitability.

    Beyond financial performance, several studies explore the relationship between board overlap and firm investment decisions, including mergers and acquisitions, research and development (R\&D), and technology investment. \citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections benefit the acquiring firm. One paper examines whether interconnected firms are more likely to invest in the same information technology, and finds a positive association between board overlap and investment in IT among firms with active boards, indicating that timely communication is an important factor in whether board overlap matters (\cite{cheng2021social}). Finally, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that lower research and development expenditure and more product differentiation are key mechanisms through which overlapping firms improve profitability.

    Despite these insights, significant gaps remain in our understanding of board overlap. Research has largely focused on governance or firm-level performance, while firm behavior and the broader market implications of board overlap remain underexplored. Additionally, most studies have been limited to large, publicly traded firms. This focus stems from the assumption that nonprofits engage in less anti-competitive behavior (\cite{baer2014clayton}; \cite{aai2013section7}), making board overlap in such organizations less concerning. However, this assumption is flawed, particularly in industries like health care, where nonprofit hospitals dominate and have been shown to engage in anti-competitive practices (\cite{hulver2023ftc}). Further research is needed to assess the competitive effects of board overlap across different market settings, particularly in nonprofit sectors where such dynamics may have substantial implications for industry competition and consumer welfare. I draw from this literature to inform potential outcomes of interest when considering board overlap in hospitals: financial performance, mergers and acquisitions, investment into the hospital, and product differentiation. I discuss in Section \ref{sec:hospbehaviors} how these outcomes related to hospitals specifically.

    \subsection{Common Ownership}\label{sec:commonownerlit}

    As consolidation in investment companies and in many industries has increased over time, so has the prevalence of common ownership among publicly traded firms. That is, firms within the same industry and market are increasingly owned by the same investment companies. While a separate occurrence and literature than that of board overlap, there are numerous parallels between the two concepts. Unlike publicly traded firms, nonprofits do not have shareholders that receive dividend payments. However, in some ways, a board functions similarly to shareholders: they influence the strategic direction of the organization through information sharing and voting on firm decisions without actively managing day-to-day operations. Further, the literature on common ownership in relation to market concentration and economic incentives is more developed than that of board overlap. Thus, I draw on this literature to inform outcomes that are potentially related to board overlap in the hospital industry.

    Whether common ownership affects competition is a widely debated theme in the economics literature. Theoretical models predict that common ownership will lead to lower equilibrium output and higher markups through firms placing non-zero weight on competitors' profit (\cite{rubinstein1983competitive}; \cite{rotemberg1984financial}; \cite{azar2012new}). Empirical evidence has grown tremendously since the seminal studies by \citeauthor{he2017product} (\citeyear{he2017product}) and \citeauthor{azar2018anticompetitive} (\citeyear{azar2018anticompetitive}), who found that common ownership leads to greater consolidation (acquisitions, joint ventures, strategic alliances), and that common ownership led to higher prices in the airline industry, respectively. However, these findings have been critiqued for how they measure common ownership and endogeneity concerns (\cite{kennedy2017competitive}; \cite{lewellen2021does}).

    A number of studies have since attempted to estimate the effect of common ownership on a range of outcomes including prices, research and development, merger activity, and patent durability (\cite{gerardi2023critical}). Many of which document a statistically significant relationship between common ownership and the relevant outcome. However, recent studies using alternative identification strategies find that common ownership can actually increase competition through lower insider trading profits, more product development, and higher investments (\cite{chen2023does}; \cite{kini2024common}).

    Additionally, recent literature has considered common ownership in the context of private firms with venture capitalists, where consolidated ownership is even more prevalent, and has found that competition decreases and information sharing increases in this setting (\cite{lindsey2008blurring}; \cite{gonzalez2020exchanges}; \cite{li2023common}; \cite{eldar2024common}).

    The literature on common ownership yields several conclusions for a study on board overlap. The first is a theoretical foundation confirming the importance of particular outcomes relevant to competition and consumer welfare, including merger and acquisition activity and investment decisions. Second, the common ownership literature focuses much more on how prices are impacted, which is also relevant in the health care setting. Finally, this literature introduces important concerns about the endogenous choice of owners or board members. While this study is not causally estimating the effect of board overlap, this is an important consideration for future research.

    \subsection{Consolidation in Hospitals}

    There is a rich empirical literature examining the effects of consolidation in health care that includes both vertical and horizontal integration between hospitals, practices, and physicians. As the focus of this paper is affiliation among competing hospitals, I draw on the literature examining hospital consolidation through mergers and systems to inform hospital behaviors that might be correlated with board overlap but are unique to the hospital setting and thus weren't discussed in Section \ref{sec:boardoverlaplit}. The main outcomes of interest in this area of research have centered around price and quality of care.

    Early studies on hospital consolidation found very little evidence of impacts on operating costs, efficiency, or quality of care (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}). From 2000-2020, the occurrence of hospital mergers increased dramatically, where a quarter of hospital markets included zero independent hospitals by 2020 (\cite{ElevanceHealth2023}). This increased concentration has led to a vast number of studies estimating its effect on quality, operating costs, and prices.

    Studies typically document cost and efficiency gains experienced by hospitals who merge to become part of a larger system (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}), even if the effects are small (\cite{craig2021mergers}). Naturally, this leads to the question of whether such efficiency gains are passed through to patients by lower prices. Across various settings, data sources, and methodology, hospital mergers are associated with significant increases in price due to hospitals having greater negotiating power with insurance companies (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}).

    There is mixed evidence on whether such gains in efficiency translate to better quality of care.  Several studies comparing hospitals who became a part of a system with observably similar hospitals that did not merge find little to no effect on readmission or mortality rates on average (\cite{haas2011mergers}; \cite{beaulieu2020changes}). However, one study finds that among heart disease patients, treatment intensity and mortality rates both increase after hospital consolidation (\cite{hayford2012impact}), and another finds increased readmission rates across all types of patients (\cite{andreyeva2024corporatization}).

    A select number of studies investigate other dimensions that could be affected by more system affiliation over time. In a slightly different setting of acquisitions of dialysis centers, \citeauthor{eliason2020acquisitions} (\citeyear{eliason2020acquisitions}) finds that decreases in quality are driven primarily by the convergence of treatment styles among acquired centers and their acquirer. Additional studies have suggested that consolidated service offerings is a potential mechanism for effects on quality (\cite{mariani2022impact}). Finally, \citeauthor{desai2023hospital} (\citeyear{desai2023hospital}) examine admission patterns of low-income patients as a result of increased hospital concentration, and find that Medicaid admissions decrease as a result of increased concentration. That is, highly concentrated markets redistribute patients within the market (\cite{desai2023hospital}).

    This literature informs ways that hospital affiliations could affect firm behaviors and outcomes specific to hospitals. In addition to the behaviors discussed in Sections \ref{sec:boardoverlaplit} and \ref{sec:commonownerlit}, measures of quality of care, treatment styles, service offerings, and admission patterns could also be important aspects driven by affiliation with other hospitals. I discuss these in more detail in Section \ref{sec:hospbehaviors}.

    \section{Documenting Overlapping Board Members in US Hospitals}\label{sec:data}

    \subsection{Data Construction}

   I compile data on nonprofit hospital board of directors from publicly available Tax Form 990s, which all sufficiently large nonprofits must file with the Internal Revenue Service (IRS) each year. These forms contain a section in which nonprofits declare their board, executives, and highest compensated employees. This paper focuses only on the board of director identities. I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital. PDF versions of the 990s are made publicly available by Nonprofit Explorer dating back at least 20 years. However, using optical character recognition text extraction methods on these PDFs is both time consuming and can lead to small errors in names extracted. Since I rely heavily on the accuracy of matching individuals, I analyze the data from 2017 onward, when the IRS began publishing XML versions of these files, which ensures accurate names that can be matched across hospitals and time. This also provides a more current picture of the prevalence of board overlap in this industry. I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees". I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics. There are 2,096 hospitals in the initial extraction of names.

   Next, I match hospitals in the tax form data to hospitals in other publicly available data sources. The only common information about hospitals across these sources is the name and location. Therefore, I use fuzzy string matching methods to create a crosswalk of EINs to identification numbers in the American Hospital Association (AHA) survey data. Specifically, I compute the Jaro-Winkler distance between hospital names in each data set. Then, I record matches with sufficiently similar names and addresses. I have a sample of 1,623 hospitals in the tax data that can be linked to the AHA survey.

    I limit the individuals in the sample to those on the board of directors, excluding executives that also serve on the board. I clean the board member names by removing common titles (Dr., Reverand, etc.), and combining names within the same firm that have slight spelling differences. For example, if John Matthews is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person. Finally, to avoid issues in whether first or last names are listed first, I sort each individual's names alphabetically.

    I define a hospital as having an overlapping board member if any of the individuals on their board are also on the board of a hospital in the same hospital referral region (HRR). HRRs are market designations created by Dartmouth Atlas using patient referral data which are meant to capture geographic regions of the average patient's choice set of health care facilities. I make this restriction first because board overlap is predicted to matter among competing firms, and hospitals compete for patients within the same geographic region. Additionally, hospital mergers have been shown to have a more significant effect when the entities involved are located geographically close (\cite{cooper2019price}). Therefore, the relevant measure for board overlap in hospitals should be those within the same market.

    Another restriction that I place for two hospitals to share overlapping board members is that they must not be affiliated by system or network. It is common for within-system hospitals to either share common board members or be governed by the same over-arching board. However, systems are already formally affiliated, and this affiliation can dictate many of the behaviors I discuss in Section \ref{sec:hospbehaviors}. Therefore, I rely on variation in board overlap that does not include system affiliation. Therefore, any hospitals with board overlap in my sample are otherwise unaffiliated. There still may exist board overlap across systems. That is, I still consider hospital A belonging to System 1 as eligible to share board members with hospital B in System 2.

    \subsection{Summary Statistics}

    First, I present a summary statistics table showing characteristics of all hospitals in the final sample. These hospitals are all nonprofit due to the restriction of only nonprofits filing Tax Form 990s.

    Of the approximately 1400 hospitals in the final sample, 7\% are affiliated with another hospital through a shared board member. In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decrease from 132 affiliated hospitals in 2017 to 117 affiliated hospitals in 2021. As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population. Therefore, in Figure \ref{fig:connected_HRR_percent}, I show the prevalence of having shared board members within a hospital market. Of the 306 HRRs in the US, approximately 17\% have at least one pair of board affiliated hospitals within the market. This proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation.

    \begin{figure}[ht!]
        \centering
        \vspace{9mm}
        \caption{Percent of hospitals that have a board affiliation}
        \includegraphics[width=.8\textwidth]{Objects/connected_percent.pdf}
        \label{fig:connected_percent}
    \end{figure}

    \begin{figure}[ht!]
        \centering
        \caption{Percent of HRRs containing board affiliated hospitals}
        \includegraphics[width=.8\textwidth]{Objects/connected_HRR_percent.pdf}
        \label{fig:connected_HRR_percent}
    \end{figure}

    Next, I present he geographic distribution of these pairs in Figure \ref{fig:connected_maps}. The blue dots represent hospitals that share a common board member, and the red dots are all other hospitals. Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population. Over time, the connections remain fairly consistent geographically.

    \begin{figure}[ht!]
        \centering
        \caption{Geographic distribution of board affiliated hospitals over time}
        \includegraphics[width=\textwidth]{Objects/connected_maps.pdf}
        \label{fig:connected_maps}
    \end{figure}

    I present the proportion of the type of hospital affiliations in Table \ref{tab:hospital_pair_types}. Each hospital is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or system, all characteristics found in the AHA data. Two-thirds of board affiliated pairs are between two adult general hospitals, and 10\% of pairs are between a specialty and a general hospital. In the main analysis, I focus only on the affiliations between adult general hospitals, as this is the most prominent pair type and the most direct competition. There is also significant variation in the ownership type of pairs. Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system. In total, there are 217 hospitals that are affiliated with at least one other hospital through shared board members at some point in the sample.

    \import{Objects}{hospital_pair_types.tex}

    I now shift towards examining summary statistics of hospitals with different types of affiliation relationships. First, in Table \ref{tab:hospital_general_summarystats}, I focus on general hospitals in four categories: board affiliated with another general hospital, board affiliated with a specialty hospital, unaffiliated through board members but belonging to a system, and completely unaffiliated. I present means of hospitals in each of these categories for the number of beds, the number of patients seen, whether it is an academic medical center, whether the hospital is located in a metropolitan area, and whether the hospital belongs to a system. On these dimensions, hospitals affiliated with other general hospitals (column 1) are relatively similar to unaffiliated hospitals that are part of a system (column 3). Unaffiliated hospitals that are not part of a system are slightly smaller in terms of bed size and number of patients seen, and less likely to be in a metropolitan area. The hospitals in column (2) are outliers, with a large number of beds and patients seen, and mainly in metropolitan areas. There are only 12 hospitals in this sample that are likely very different from the average general hospital.

    \import{Objects}{hospital_general_summarystats.tex}

    I also present means of various measures of patient demographics, include the number of Medicare and Medicaid patients normalized by total beds, the total number of Medicare patients from CMS, the average Medicare patient age, and the average Medicare patient risk score. These measures yield similar descriptive results, that the unaffiliated hospitals that are not part of a system are slightly smaller, but they see similar patients in terms of age and complexity. Finally, I show means of variables to capture aspects of the service offerings of the different types of hospitals. First, from the AHA survey, I include an indicator for whether the hospital offers services in a Neonatal Intensive Care Unit (NICU) or Cardiac catheterization lab, both specialized and profitable services. General hospitals that are affiliated with specialty hospitals are the most likely to offer these services. Hospitals without any board member affiliations but belonging to a system are more likely to offer these services than unaffiliated hospitals that do not belong to a system. Hospitals that are affiliated with another general hospitals fall somewhere in between these two, with slightly less likelihood than the unaffiliated hospitals belonging to a system. I also include a measure of the concentration of bed allocation, where a higher concentration indicates that a hospital's bed capacity is concentrated in fewer service categories, whereas a lower concentration suggests a more diversified service offering. I also compute a similar concentration metric using the percent of patients in several service categories from the CMS Provider Utilization Files. For both measures, there seems to be little variation in service concentration across affiliation types.

    Next, I present the same summary statistics for specialty hospitals with different types of affiliations: affiliated with a general hospital, unaffiliated through board but belonging to a system, and unaffiliated through board or system, shown in Table \ref{tab:hospital_specialty_summarystats}. Overall, there are less specialty hospitals included in the sample, and very few of them are affiliated through board members. Specialty hospitals are more concentrated than general hospitals, as expected. In the AHA measure of concentration, all specialty hospitals are similarly concentrated. However, when using the services included in the CMS files, the concentration of unaffiliated system hospitals is much lower than that of any other specialty hospital. This could be due to the limited services included in the CMS data. Since this is such a small and specialized sample of hospitals, I do not include them in the main analysis.

    \import{Objects}{hospital_specialty_summarystats.tex}

    \section{Hospital Behaviors}\label{sec:hospbehaviors}

    I consider the literature on overlapping boards, common ownership, and hospital consolidation discussed in Section to theoretically found the firm behaviors that may be associated with overlapping boards in this particular industry.

    \subsection{Variable Definitions}

    First, I define an indicator variable for whether a hospital has a shared board member with another hospital within the same market, which I refer to as being board affiliated. Specifically, a hospital is board affiliated if the same individual is a board member of two otherwise unaffiliated hospitals. That is, two hospitals that do not belong to the same system or network can be affiliated through board members, but same-system or same-network hospitals can not. I make this distinction to separate the variation in affiliation by board member alone versus simply having shared board members due to being in the same system, which is likely already correlated with service offerings. I also focus on board affiliations within the same hospital market, as the outcomes of interest focus on competing for patients in a specific geographic region. I measure hospital markets using Hospital Referral Regions (HRRs) as defined by Dartmouth Atlas of Healthcare.

    I will now define the five outcomes that I analyze in Section \ref{sec:boardanalysis}. I construct two measures of specialization based on modified Herfindahl-Hirschman Index (HHI) calculations. The first uses the number of beds allocated to specific services, variables found in the AHA data. For services $j$ in the list of all possible services\footnote{General medical and surgical (adult), General medical and surgical (pediatric), Obstetric care, Medical/surgical intensive care, Cardiac intensive care, Neonatal intensive care, Neonatal intermediate care, Pediatric intensive care beds, Burn care, Other Special Care, Other intensive care, Physical Rehabilitation care, Alcohol/drug abuse or dependency inpatient care, Psychiatric care, Skilled nursing care, Intermediate nursing care, Acute long term care, Other long-term care, Other care}, the concentration of bed allocation is defined as

    $$HHI_{bed} = \sum_{j}\left(\frac{beds_j}{total beds}\right)^2.$$

    The second concentration measure focuses on patients seen for certain conditions in each hospital. The CMS Provider Utilization files contain measures of the percent of patients seen for different conditions. Therefore, for services $k$ in the list of conditions in the CMS data\footnote{Percent Of Beneficiaries Identified With Combined Cancer Flag For 6 Cancer Indicators, Percent Of Beneficiaries Identified With Chronic Obstructive Pulmonary Disease, Percent Of Beneficiaries Identified With Chronic Kidney Disease, Percent Of Beneficiaries Identified With Heart Failure And Non-Ischemic Heart Disease}, I define the concentration of patient conditions in each hospital as

    $$HHI_{patient} = \sum_{k}\left(\frac{patients_k}{total patients}\right)^2.$$

    The other outcomes I consider are the total number of Medicare beneficiaries seen, from the CMS data, the total number of Medicaid discharges, from the AHA data, and the average risk score of Medicare beneficiaries, from the CMS data. I present a table of summary statistics for these outcomes, as well as demographics, of the hospitals in the sample, in Table \ref{all_sumstats}.

    \import{Objects}{all_summarystats.tex}

    \section{Documenting Board of Director Affiliation}

    In Figure \ref{fig:concentration_services_time}, I show the AHA measure of service concentration based on the number of beds over time for general hospitals with different levels of affiliation. General hospitals that are unaffiliated through board members, whether they are affiliated with a system or not, remain relatively constant in their service offerings over time. However, general hospitals that are affiliated with another general hospital through a shared board member display increasing service concentration over time, providing descriptive evidence that there may be some correlation between board affiliations and the services offered in hospitals.

    \begin{figure}[ht!]
        \centering
        \caption{Concentration of services offered over time}
        \includegraphics[width=0.8\textwidth]{Objects/concentration_services_time.pdf}
        \label{fig:concentration_services_time}
    \end{figure}

    \newpage

    \printbibliography

\end{document}

### 2025-03-29 Paper/descriptive_paper.tex
Updates from Overleaf

@@
     In this paper, I form networks of hospital board of director members using Tax Form 990s, which requires nonprofits in the US to declare the identities and compensation information of their board.
-Extracting these identities and matching to other sources of hospital information, I show that approximately 7\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations.
-[Add more results]
+Extracting these identities and matching to other sources of hospital information, I show that approximately 8\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations.
+Among hospitals that share board members, two thirds of the hospital pairs are between two general adult hospitals.
+That is, a majority of the shared board members serve on directly competing hospitals in the same geographic market.
 
@@
 
-    \section{Related Literature}
+    \section{Related Literature}\label{sec:relatedlit}
 
@@
 
-    \section{Documenting Overlapping Board Members in US Hospitals}\label{sec:data}
+    \subsection{Informing Hospital Behaviors}\label{sec:hospbehaviors}
+
+    The main purpose of this paper is to document the prevalence of board overlap in US hospitals in Section \ref{sec:document}.
+While I do not establish any causal relationship between board overlap and hospital behaviors, I consider the literature on board overlap, common ownership, and hospital consolidation discussed in Section \ref{sec:relatedlit} to inform a set of hospital behaviors that may be relevant for future research.
+I divide these behaviors into three categories: price/quality, investment decisions, and differentiation.
+
+    The hospital consolidation literature highlights price and quality of care as key factors in the industry.
+Mergers often grant hospitals greater market power, enabling them to negotiate higher prices with insurance companies.
+It is unclear whether shared board members would lead to similar price increases.
+If board overlap facilitates information sharing about negotiations, it could indirectly influence pricing.
+However, direct board involvement in pricing decisions seems less likely.
+Nevertheless, with access to data on actual prices charged by hospitals, this remains an interesting outcome for future research to explore.
+
+    It seems more likely that quality of care could be impacted by overlapping board members among hospitals.
+There are direct actions that could lead to changes in quality such as convergence in treatment styles or investment in technology, which I will discuss in detail in the following sections.
+There is also an indirect path through which overlapping board members might impact quality, which is that hospitals might compete for the same pool of experienced board members.
+If the board members with the most expertise are in high demand, sharing that expertise among multiple hospitals could facilitate greater improvements.
+The most commonly used measurements of quality of care in hospitals are mortality and readmission rates.
+
+    In both board overlap and common ownership, several investment decisions are shown to be associated with established firm affiliation.
+First, hospitals with board affiliations might be more likely to merge in the future.
+Second, sharing board members could facilitate engaging in the same types of technology or capital.
+Therefore, an area of future research is to look at outcomes such as health information technology investments, specific medical technology advancements in the hospital, and capital investments.
+
+    Finally, previous research has shown that product differentiation is associated with board overlap in other industries.
+That is, sharing information among competing hospitals leads to more specialization.
+In hospitals, there are two distinct types of differentiation: either through services offered, or the types of patients admitted.
+Both of these are also key findings in the hospital merger literature, indicating that affiliations through board members could also facilitate changes in these areas.
+
+    \section{Documenting Overlapping Board Members in US Hospitals}\label{sec:document}
 
@@
 
-    First, I present a summary statistics table showing characteristics of all hospitals in the final sample.
-These hospitals are all nonprofit due to the restriction of only nonprofits filing Tax Form 990s.
+    First, I present a summary statistics table showing characteristics of all hospitals in the final sample, which consists of 1,466 nonprofit hospitals, almost all of which are adult general hospitals.
+They have, on average, 178 beds and 14 board members.
+Almost 40\% of the hospitals in the sample are academic medical centers, almost 70\% of the sample are located in metropolitan areas, and 56\% belong to a system.
 
-    Of the approximately 1400 hospitals in the final sample, 7\% are affiliated with another hospital through a shared board member.
-In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decrease from 132 affiliated hospitals in 2017 to 117 affiliated hospitals in 2021.
+    \import{Objects}{all_summarystats.tex}
+
+    Eight percent of hospitals in the sample have overlapping board members with an otherwise unaffiliated hospital.
+In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decrease from 129 affiliated hospitals in 2017 to 114 affiliated hospitals in 2021.
+I include 2022 values, but there are less total firms in the sample due to lag in tax forms being made available publicly.
 As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population.
@@
 This proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation.
+This is likely due to increased system affiliations, which mechanically decreases board overlap according to the definition I use here.
 
@@
 
-    Next, I present he geographic distribution of these pairs in Figure \ref{fig:connected_maps}.
-The blue dots represent hospitals that share a common board member, and the red dots are all other hospitals.
+    Next, I present geographic distribution of these pairs in Figure \ref{fig:connected_maps}.
+The blue lines represent hospitals that share a common board member, and the red dots are all other hospitals.
 Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population.
 Over time, the connections remain fairly consistent geographically.
+Taken together, these show that board overlap is a prevalent occurrence among nonprofit hospitals in the US from 2017 onward.
 
@@
 
-    I present the proportion of the type of hospital affiliations in Table \ref{tab:hospital_pair_types}.
-Each hospital is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or system, all characteristics found in the AHA data.
+    I now shift to examining characteristics of hospitals with board overlap.
+I present the proportion of the type of hospital pairs in Table \ref{tab:hospital_pair_types}.
+Each hospital is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data.
 Two-thirds of board affiliated pairs are between two adult general hospitals, and 10\% of pairs are between a specialty and a general hospital.
-In the main analysis, I focus only on the affiliations between adult general hospitals, as this is the most prominent pair type and the most direct competition.
-There is also significant variation in the ownership type of pairs.
-Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system.
-In total, there are 217 hospitals that are affiliated with at least one other hospital through shared board members at some point in the sample.
+Thus, for the majority of hospitals that have board interlock, the hospital they share a board member with is a direct competitor.
 
     \import{Objects}{hospital_pair_types.tex}
+
+    There is also significant variation in the ownership type of hospitals within affiliated pairs.
+Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system.
+The lowest proportion of shared board member pairs is made up of two hospitals belonging to different systems.
+In total, there are 217 hospitals that share a board member with at least one other hospital at some point in the sample.
+
+    In Table
 
@@
 
-    \section{Hospital Behaviors}\label{sec:hospbehaviors}
-
-    I consider the literature on overlapping boards, common ownership, and hospital consolidation discussed in Section to theoretically found the firm behaviors that may be associated with overlapping boards in this particular industry.
-
-    \subsection{Variable Definitions}
-
-    First, I define an indicator variable for whether a hospital has a shared board member with another hospital within the same market, which I refer to as being board affiliated.
-Specifically, a hospital is board affiliated if the same individual is a board member of two otherwise unaffiliated hospitals.
-That is, two hospitals that do not belong to the same system or network can be affiliated through board members, but same-system or same-network hospitals can not.
-I make this distinction to separate the variation in affiliation by board member alone versus simply having shared board members due to being in the same system, which is likely already correlated with service offerings.
-I also focus on board affiliations within the same hospital market, as the outcomes of interest focus on competing for patients in a specific geographic region.
-I measure hospital markets using Hospital Referral Regions (HRRs) as defined by Dartmouth Atlas of Healthcare.
-
-    I will now define the five outcomes that I analyze in Section \ref{sec:boardanalysis}.
-I construct two measures of specialization based on modified Herfindahl-Hirschman Index (HHI) calculations.
-The first uses the number of beds allocated to specific services, variables found in the AHA data.
-For services $j$ in the list of all possible services\footnote{General medical and surgical (adult), General medical and surgical (pediatric), Obstetric care, Medical/surgical intensive care, Cardiac intensive care, Neonatal intensive care, Neonatal intermediate care, Pediatric intensive care beds, Burn care, Other Special Care, Other intensive care, Physical Rehabilitation care, Alcohol/drug abuse or dependency inpatient care, Psychiatric care, Skilled nursing care, Intermediate nursing care, Acute long term care, Other long-term care, Other care}, the concentration of bed allocation is defined as
-
-    $$HHI_{bed} = \sum_{j}\left(\frac{beds_j}{total beds}\right)^2.$$
-
-    The second concentration measure focuses on patients seen for certain conditions in each hospital.
-The CMS Provider Utilization files contain measures of the percent of patients seen for different conditions.
-Therefore, for services $k$ in the list of conditions in the CMS data\footnote{Percent Of Beneficiaries Identified With Combined Cancer Flag For 6 Cancer Indicators, Percent Of Beneficiaries Identified With Chronic Obstructive Pulmonary Disease, Percent Of Beneficiaries Identified With Chronic Kidney Disease, Percent Of Beneficiaries Identified With Heart Failure And Non-Ischemic Heart Disease}, I define the concentration of patient conditions in each hospital as
-
-    $$HHI_{patient} = \sum_{k}\left(\frac{patients_k}{total patients}\right)^2.$$
-
-    The other outcomes I consider are the total number of Medicare beneficiaries seen, from the CMS data, the total number of Medicaid discharges, from the AHA data, and the average risk score of Medicare beneficiaries, from the CMS data.
-I present a table of summary statistics for these outcomes, as well as demographics, of the hospitals in the sample, in Table \ref{all_sumstats}.
-
-    \import{Objects}{all_summarystats.tex}
-
-    \section{Documenting Board of Director Affiliation}
-
-    In Figure \ref{fig:concentration_services_time}, I show the AHA measure of service concentration based on the number of beds over time for general hospitals with different levels of affiliation.
-General hospitals that are unaffiliated through board members, whether they are affiliated with a system or not, remain relatively constant in their service offerings over time.
-However, general hospitals that are affiliated with another general hospital through a shared board member display increasing service concentration over time, providing descriptive evidence that there may be some correlation between board affiliations and the services offered in hospitals.
-
-    \begin{figure}[ht!]
-        \centering
-        \caption{Concentration of services offered over time}
-        \includegraphics[width=0.8\textwidth]{Objects/concentration_services_time.pdf}
-        \label{fig:concentration_services_time}
-    \end{figure}
-
     \newpage

### 2025-03-29 Paper/descriptive_paper.tex
Updates from Overleaf

@@
     Consolidation in health care has been a heavily discussed topic in research and regulation, as more concentration has been linked to higher prices and lower quality.
-Aside from formal mergers and acquisitions, there is opportunity for hospitals to form affiliations with other competing hospitals through sharing common board members, which has only been studied in the context of large public firms.
+Aside from formal mergers and acquisitions, there is opportunity for hospitals to form affiliations with other competing hospitals through sharing common board members, which has mainly been studied in the context of large public firms.
 This paper establishes the existence of overlapping board members in competing nonprofit hospitals in the US.
-Using a data set constructed from IRS Tax Form 990s, I show that 7\% of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market.
-These findings highlight an understudied mechanism through which hospitals can collude, potentially affecting patient and market outcomes.
+Using a data set constructed from IRS Tax Form 990s, I show that 8\% of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market.
+These findings highlight an understudied mechanism through which hospitals form affiliations, potentially affecting patient and market outcomes.
 		} 
@@
 Yet, despite concerns of low competition, this type of connectedness among firms has not been given much consideration in health care.
-In this paper, I document the extent of overlapping board members in nonprofit hospitals in the US and explore avenues of future research to investigate
-    potential anti-competitive effects.
+In this paper, I document the extent of overlapping board members in nonprofit hospitals in the US, and explore avenues of future research to investigate
+    potential anti-competitive effects of this type of affiliation.
 
@@
     Beyond financial performance, several studies explore the relationship between board overlap and firm investment decisions, including mergers and acquisitions, research and development (R\&D), and technology investment.
-\citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections benefit the acquiring firm.
+\citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections lead to better financial performance for the acquirer after the transaction.
 One paper examines whether interconnected firms are more likely to invest in the same information technology, and finds a positive association between board overlap and investment in IT among firms with active boards, indicating that timely communication is an important factor in whether board overlap matters (\cite{cheng2021social}).
@@
 Further research is needed to assess the competitive effects of board overlap across different market settings, particularly in nonprofit sectors where such dynamics may have substantial implications for industry competition and consumer welfare.
-I draw from this literature to inform potential outcomes of interest when considering board overlap in hospitals: financial performance, mergers and acquisitions, investment into the hospital, and product differentiation.
+I draw from this literature to inform potential outcomes of interest when considering board overlap in hospitals: financial performance, mergers and acquisitions, investment, and product differentiation.
 I discuss in Section \ref{sec:hospbehaviors} how these outcomes related to hospitals specifically.
@@
 However, recent studies using alternative identification strategies find that common ownership can actually increase competition through lower insider trading profits, more product development, and higher investments (\cite{chen2023does}; \cite{kini2024common}).
-
 Additionally, recent literature has considered common ownership in the context of private firms with venture capitalists, where consolidated ownership is even more prevalent, and has found that competition decreases and information sharing increases in this setting (\cite{lindsey2008blurring}; \cite{gonzalez2020exchanges}; \cite{li2023common}; \cite{eldar2024common}).
@@
 Finally, this literature introduces important concerns about the endogenous choice of owners or board members.
-While this study is not causally estimating the effect of board overlap, this is an important consideration for future research.
+While this study does not seek to causally estimate the effects of board overlap in hospitals, this is an important consideration for future research.
 
@@
     In both board overlap and common ownership, several investment decisions are shown to be associated with established firm affiliation.
-First, hospitals with board affiliations might be more likely to merge in the future.
+First, hospitals with board shared board members might be more likely to merge in the future.
 Second, sharing board members could facilitate engaging in the same types of technology or capital.
@@
     Finally, previous research has shown that product differentiation is associated with board overlap in other industries.
-That is, sharing information among competing hospitals leads to more specialization.
+That is, sharing information among competing hospitals can lead to more specialization.
 In hospitals, there are two distinct types of differentiation: either through services offered, or the types of patients admitted.
-Both of these are also key findings in the hospital merger literature, indicating that affiliations through board members could also facilitate changes in these areas.
+Both of these behaviors are key findings in the hospital merger literature, indicating that affiliations through board members could also facilitate changes in these areas.
+In future research, one could think about measuring the likelihood of specialized services being offered, such as a Neonatal Intensive Care Unit or cardiac catheterization laboratory, where hospitals might shift investments towards services not offered by their affiliated hospital.
+Additionally, patient composition in the hospital, measured by the payer mix of admitted patients or complexity of their conditions, could be compared among hospitals with and without shared board members.
 
@@
 I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics.
-There are 2,096 hospitals in the initial extraction of names.
+There are 2,096 hospitals in the initial extraction of firms and identities.
 

### 2025-03-29 Scripts/descriptive_tables_and_figures.R
add size aspects to hospital pair types

@@ -16,16 +16,15 @@ library(did)
 created_data_path <- "CreatedData/"
 
 
-## SCRIPT CREATING SUMMARY STATISTICS FOR THIRD CHAPTER ################
+###### PART 1: CLEAN DATA ###############################################################################################################################
 
-# Part 1: Clean the Data ---------------------------------------------------------
 # Read in cleaned up version of names found in 990 tax forms
 people <- read_rds(paste0(created_data_path, "cleaned_people_data.rds"))
+
 # Read is crosswalk of AHA ID - EIN, only keeping those who match to a standalone hospital
 cw <- readRDS(paste0(created_data_path, "updated_ein_aha_cw.rds")) 
 
-
-
+# Clean up the people data 
 # only keep people whose EIN is in the crosswalk as a standalone hospital
 people <- people %>% filter(Filer.EIN %in% cw$Filer.EIN)
 
@@ -44,8 +43,6 @@ ein_keep <- board_people %>%
 
 board_people <- board_people %>%
   filter(Filer.EIN %in% ein_keep$Filer.EIN)
-
-rm(ein_keep)
 
 # clean up extras in names
 board_people$name_cleaned <- str_remove(board_people$name_cleaned, "deceased")
@@ -113,6 +110,12 @@ board_people <- standardize_names_optimized(board_people, max_dist = 3)
 board_people <- board_people %>%
   standardize_names_by_hrr(., max_dist = 2)
 
+# create variable for number of people on board
+number_board_people <- board_people %>%
+  group_by(Filer.EIN, TaxYr) %>%
+  summarise(num_board = n()) %>%
+  ungroup()
+
 
 # run the script titled "function2_find_common_board_members"
 source("Scripts//function2_identify_common_members.R")
@@ -144,11 +147,16 @@ board_people <- board_people %>%
 # trim the white space on other_ein
 board_people$other_ein <- str_trim(board_people$other_ein)
 
-# Merge in information about the root and connected hospital
+# Merge in AHA geographic information
 AHA_geog <- AHA %>%
   select(ID, YEAR, SYSID, NETNAME, FSTCD, LONG, LAT, HRRCODE, MNAME) %>%
   filter(YEAR>=2016 & YEAR<=2022)
-  # no duplicates in ID, YEAR
+
[…]

### 2025-03-30 Scripts/descriptive_tables_and_figures.R
add hospital behavior tables for descriptive paper

@@ -249,11 +249,12 @@ filer_data <- hospital_connections %>%
   mutate(connected = ifelse(other_ein!="" & !is.na(other_ein), 1, NA)) %>%
   group_by(TaxYr, filer_id) %>%
   fill(connected, .direction = "downup") %>%
+  mutate(num_connections = sum(connected, na.rm = TRUE)) %>%
   ungroup() %>%
   mutate(connected = ifelse(is.na(connected), 0, connected)) %>%
   distinct(TaxYr, filer_id, connected, .keep_all = TRUE) 
 
-# join number of people people
+# join number of people
 filer_data <- filer_data %>%
   mutate(TaxYr = as.character(TaxYr)) %>%
   left_join(number_board_people, by = c("TaxYr"="TaxYr", "Filer.EIN"="Filer.EIN"))
@@ -303,7 +304,8 @@ tab1_data <- filer_data %>%
                  "Academic" = "academic",
                  "In Metrop. Area"="metro",
                  "Number of Board Members" = "num_board",
-                 "Overlapping Board"="connected"), 
+                 "Overlapping Board"="connected",
+                 "Number of Connections" = "num_connections"), 
                list(m=mean,min=min,max=max,sd=sd), na.rm=T) %>%
   mutate_if(is.numeric, ~ifelse(abs(.)==Inf,NA,.))  %>%
   gather(key=var,value=value) %>%
@@ -316,7 +318,7 @@ tab1_data <- filer_data %>%
          max = ifelse(max>9, round(max,0), round(max,2)),
          sd = ifelse(sd>9, round(sd,0), round(sd,2))) 
 
-knitr::kable(tab1_data[c(7,4,10,2,3,1,5,6,8,9,11),],
+knitr::kable(tab1_data[c(7,4,11,2,3,1,5,6,8,10,9,12),],
                format = "latex",
                col.names = c("Variable", "Mean", "Min", "Max", "SD"),
                caption = "Full Sample Summary Statistics\\label{all_sumstats}",
@@ -450,20 +452,26 @@ ggsave("Objects//connected_maps.pdf", width = 6, height = 6.5, units = "in")
 
 
 # Summarise hospital pairs
-hospital_connected_pairs <- hospital_pairs %>%
+hospital_connected_pairs <- hospital_connections %>%
   select(TaxYr, filer_id, other_id) %>%
[…]

### 2025-11-25 Objects/old/hospital_general_summarystats.tex
update name cleaning and work on descriptive tables and figures

\begin{table}[ht!]

\caption{Summary Statistics of General Hospitals}
\centering
\begin{tabular}[t]{lcccc}
\toprule
Variable & Affiliated w/ General & Affiliated w/ Specialty & Unaffiliated, Part of System & Unaffiliated, Not Part of System\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Characteristics}}\\
\hspace{1em}Number of Beds & 228.79 & 689.27 & 202.39 & 117.02\\
\hspace{1em}Academic Medical Center & 0.39 & 0.76 & 0.40 & 0.29\\
\hspace{1em}In Metropolitan Area & 0.69 & 0.91 & 0.70 & 0.53\\
\hspace{1em}In System & 0.43 & 0.91 & 0.98 & 0.00\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Patients}}\\
\hspace{1em}Medicaid Patients per Bed (AHA) & 6.92 & 8.27 & 7.63 & 6.02\\
\hspace{1em}Medicare Patients per Bed (AHA) & 17.78 & 22.29 & 20.56 & 15.89\\
\hspace{1em}Number of Medicare Patients (CMS) & 3033.37 & 6281.22 & 2654.91 & 1766.75\\
\hspace{1em}Avg Medicare Patient Age & 74.43 & 74.44 & 74.13 & 74.28\\
\hspace{1em}Avg Patient Risk Score & 2.04 & 2.09 & 2.11 & 2.03\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Services}}\\
\hspace{1em}Has a Cath Lab & 0.52 & 0.88 & 0.57 & 0.36\\
\hspace{1em}Has a NICU & 0.24 & 0.61 & 0.29 & 0.14\\
\hspace{1em}Service Concentration (beds) & 0.51 & 0.48 & 0.50 & 0.56\\
\hspace{1em}Service Concentration (conditions) & 0.52 & 0.47 & 0.57 & 0.54\\
Num. Hospitals & 157.00 & 12.00 & 836.00 & 462.00\\
\bottomrule
\end{tabular}
\end{table}

### 2025-11-25 Scripts/Read990_scheduleH.R

@@ -13,6 +13,7 @@ replace_null <- function(x) {
   purrr::map(x, ~if(is.list(.x)) replace_null(.x) else .x)
 }
 
+
 # make a list of all folders to search through
 folders <- list.files(path = paste0(drive_path, "IRS_XML_990s/"), full.names = FALSE)
 
@@ -25,7 +26,7 @@ all_ein_data <- readRDS(paste0(created_data_path, "all_ein_data_scheduleH.rds"))
 all_people_data <- readRDS(paste0(created_data_path, "all_people_data_scheduleH.rds"))
 
 # loop through each folder
-for (folder in folders[25:75]) {
+for (folder in folders) {
   # make a list of all files in the folder
   files <- list.files(path = paste0(drive_path, "IRS_XML_990s/",folder), pattern = ".xml", full.names = FALSE)
   

### 2025-11-25 Scripts/clean_matched_names_updated.R

@@ -1,0 +1,322 @@
+library(readr)
+library(stringr)
+library(dplyr)
+library(gender)
+
+source("./Scripts/paths.R")
+
+# Script to clean up the name columns of people data scraped from 990s
+
+all_people_data <- readRDS(paste0(created_data_path, "/all_people_data_scheduleH.rds"))
+cw <- readRDS(paste0(created_data_path, "/updated_ein_aha_cw.rds"))
+
+# only keep observations where EIN is found in either ein_hosp or ein_sys
+all_people_data <- all_people_data %>%
+  filter(Filer.EIN %in% cw$Filer.EIN)
+
+# look at unique values of section
+unique(all_people_data$Section)
+
+# only keep relevant sections
+all_people_data <- all_people_data %>%
+  filter(Section %in% c("Form990PartVIISectionAGrp", "RltdOrgOfficerTrstKeyEmplGrp"))
+
+# fill personNm with other lines
+all_people_data <- all_people_data %>%
+  mutate(PersonNm = ifelse(is.na(PersonNm), BusinessName_BusinessNameLine1Txt, PersonNm)) %>%
+  filter(!is.na(PersonNm) & PersonNm != "'")
+# this fills all missing person names
+
+# Look at missing titles
+observe <- all_people_data %>%
+  filter(is.na(TitleTxt))
+
+# remove observations with no title text
+all_people_data <- all_people_data %>%
+  filter(!is.na(TitleTxt))
+
+# tabulate the number of unique EINs for every value of TaxYr
+all_people_data %>%
+  group_by(TaxYr) %>%
+  summarise(n = n_distinct(Filer.EIN))
+
+# Remove 2015 and 2023
+all_people_data <- all_people_data %>%
+  filter(TaxYr!=2015 & TaxYr!=2023)
+
+# create data set with only relevant variables
+people_data <- all_people_data %>%
+  select(TaxYr, Filer.EIN, PersonNm, TitleTxt)
+
+# convert to all lowercase
+people_data <- people_data %>%
+  mutate(PersonNm = tolower(PersonNm),
+         TitleTxt = tolower(TitleTxt))
+
+# GOAL 1: GET ANY RELEVANT TITLES FROM THE NAME COLUMN #################
+# I got these lists by going through all of the names with at least 3 words. Don't do this again unless you add more names!!
[…]

### 2025-11-25 Scripts/create_connections_boardandexec.R

@@ -1,0 +1,523 @@
+# load libraries 
+library(dplyr)
+library(fuzzyjoin)
+library(stringdist)
+library(tidyr)
+library(stringr)
+library(readr)
+library(data.table)
+library(purrr)
+library(igraph)
+library(future.apply)
+
+source("./Scripts/paths.R")
+
+# functions ------------------------------------------------------------------------------------------------------
+
+combine_fuzzy_matches <- function(data, max_dist) {
+  setDT(data)  # Convert to data.table for efficiency
+  
+  # Split by EIN to avoid excessive joins
+  firm_list <- split(data, by = "Filer.EIN", keep.by = TRUE)
+  
+  # Function to process each firm's data
+  process_firm <- function(firm_data) {
+    if (nrow(firm_data) <= 1) return(firm_data)  # Skip firms with only 1 name
+    
+    unique_names <- unique(firm_data[, .(name_cleaned)])  # Reduce size
+    
+    # Compute approximate name matches only within this firm
+    matches <- stringdist_inner_join(
+      unique_names, unique_names, 
+      by = "name_cleaned", 
+      max_dist = max_dist, 
+      method = "osa"
+    ) %>%
+      filter(name_cleaned.x < name_cleaned.y) %>%
+      group_by(name_cleaned.y) %>%
+      summarise(correct_name = first(name_cleaned.x), .groups = "drop")
+    
+    # Merge corrected names
+    setDT(matches)
+    firm_data <- merge(firm_data, matches, by.x = "name_cleaned", by.y = "name_cleaned.y", all.x = TRUE)
+    firm_data[, name_cleaned := fcoalesce(correct_name, name_cleaned)]
+    firm_data[, correct_name := NULL]
+    
+    return(firm_data)
+  }
+  
+  # Process firms in parallel (reduces RAM usage)
+  corrected_list <- lapply(firm_list, process_firm)
+  
+  # Recombine data
+  corrected_data <- rbindlist(corrected_list, use.names = TRUE, fill = TRUE)
+  
+  return(corrected_data)
+}
+
+# nicknames mapping
+nickname_map <- list(
+  "jeff" = "jeffrey",
+  "bob" = "robert",
+  "rob" = "robert",
+  "liz" = "elizabeth",
+  "kate" = "katherine",
+  "mike" = "michael",
+  "tom" = "thomas",
+  "tim" = "timothy",
+  "pat" = "patricia",
[…]

### 2025-11-25 Scripts/create_connections_boardonly.R

@@ -1,0 +1,590 @@
+# load libraries 
+library(dplyr)
+library(fuzzyjoin)
+library(stringdist)
+library(tidyr)
+library(stringr)
+library(readr)
+library(data.table)
+library(purrr)
+library(igraph)
+library(future.apply)
+
+source("./Scripts/paths.R")
+
+# functions ------------------------------------------------------------------------------------------------------
+
+combine_fuzzy_matches <- function(data, max_dist) {
+  setDT(data)  # Convert to data.table for efficiency
+  
+  # Split by EIN to avoid excessive joins
+  firm_list <- split(data, by = "Filer.EIN", keep.by = TRUE)
+  
+  # Function to process each firm's data
+  process_firm <- function(firm_data) {
+    if (nrow(firm_data) <= 1) return(firm_data)  # Skip firms with only 1 name
+    
+    unique_names <- unique(firm_data[, .(name_cleaned)])  # Reduce size
+    
+    # Compute approximate name matches only within this firm
+    matches <- stringdist_inner_join(
+      unique_names, unique_names, 
+      by = "name_cleaned", 
+      max_dist = max_dist, 
+      method = "osa"
+    ) %>%
+      filter(name_cleaned.x < name_cleaned.y) %>%
+      group_by(name_cleaned.y) %>%
+      summarise(correct_name = first(name_cleaned.x), .groups = "drop")
+    
+    # Merge corrected names
+    setDT(matches)
+    firm_data <- merge(firm_data, matches, by.x = "name_cleaned", by.y = "name_cleaned.y", all.x = TRUE)
+    firm_data[, name_cleaned := fcoalesce(correct_name, name_cleaned)]
+    firm_data[, correct_name := NULL]
+    
+    return(firm_data)
+  }
+  
+  # Process firms in parallel (reduces RAM usage)
+  corrected_list <- lapply(firm_list, process_firm)
+  
+  # Recombine data
+  corrected_data <- rbindlist(corrected_list, use.names = TRUE, fill = TRUE)
+  
+  return(corrected_data)
+}
+
+# nicknames mapping
+nickname_map <- list(
+  "jeff" = "jeffrey",
+  "bob" = "robert",
+  "rob" = "robert",
+  "liz" = "elizabeth",
+  "kate" = "katherine",
+  "mike" = "michael",
+  "tom" = "thomas",
+  "tim" = "timothy",
+  "pat" = "patricia",
+  "chris" = "christopher",
[…]

### 2025-11-25 Scripts/descriptive_tables_and_figures.R

@@ -13,359 +13,93 @@ library(purrr)
 library(ggpubr)
 library(did)
 
-created_data_path <- "CreatedData/"
-
-
-###### PART 1: CLEAN DATA ###############################################################################################################################
-
-# Read in cleaned up version of names found in 990 tax forms
-people <- read_rds(paste0(created_data_path, "cleaned_people_data.rds"))
-
-# Read is crosswalk of AHA ID - EIN, only keeping those who match to a standalone hospital
-cw <- readRDS(paste0(created_data_path, "updated_ein_aha_cw.rds")) 
-
-# Clean up the people data 
-# only keep people whose EIN is in the crosswalk as a standalone hospital
-people <- people %>% filter(Filer.EIN %in% cw$Filer.EIN)
-
-# only keep people who are board members
-board_people <- people %>%
-  filter(board_member==1) %>%
-  select(TaxYr, Filer.EIN, name_cleaned)
-
-# Only keep EINs present in 2017-2021
-ein_keep <- board_people %>%
-  distinct(TaxYr, Filer.EIN) %>%
-  mutate(count = ifelse(TaxYr %in% 2017:2021, 1, 0)) %>%
-  group_by(Filer.EIN) %>%
-  summarise(count = sum(count)) %>%
-  filter(count>=5)
-
-board_people <- board_people %>%
-  filter(Filer.EIN %in% ein_keep$Filer.EIN)
-
-# clean up extras in names
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "deceased")
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\seff$")
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\scsjp$")
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\scisa$")
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "^mrs\\s")
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "^reverend\\s")
-board_people$name_cleaned <- str_remove(board_people$name_cleaned, "\\scns$")
-board_people$name_cleaned <- str_remove_all(board_people$name_cleaned, "january|february|march|june|july|august|september|october|november|december|dec\\b")
-
[…]

### 2025-11-25 Scripts/function2_identify_common_members.R

@@ -1,12 +1,12 @@
 identify_common_members <- function(data) {
   # Collect all EINs for each name within the same year
   ein_mapping <- data %>%
-    group_by(name_cleaned, TaxYr) %>%
+    group_by(name_id, TaxYr) %>%
     summarise(all_eins = list(unique(Filer.EIN)), .groups = "drop")
   
   # Merge back and correctly compute other EINs within the same year
   df_with_others <- data %>%
-    left_join(ein_mapping, by = c("name_cleaned", "TaxYr")) %>%
+    left_join(ein_mapping, by = c("name_id", "TaxYr")) %>%
     rowwise() %>%
     mutate(other_eins = paste(setdiff(unlist(all_eins), Filer.EIN), collapse = ", ")) %>%
     ungroup() %>%

### 2025-11-25 Scripts/paths.R

@@ -1,0 +1,3 @@
+created_data_path <- "C:\\Users\\uqhglen2\\Documents\\ResearchProjects\\hospital_boardoverlap\\CreatedData\\"
+raw_data_path <- "C:\\Users\\uqhglen2\\Documents\\ResearchProjects\\RawData"
+objects_path <- "C:\\Users\\uqhglen2\\Documents\\ResearchProjects\\hospital_boardoverlap\\Objects\\"

### 2025-11-25 Scripts/updated_match_hosp.R

@@ -5,18 +5,26 @@ library(stringr)
 library(readr)
 library(cdlTools)
 
+source("./Scripts/paths.R")
+
 # Read in EIN data from XMLs
 all_ein_data <- readRDS("CreatedData/all_ein_data_scheduleH.rds")
 
 # How many EINs are in this data?
 length(unique(all_ein_data$Filer.EIN))
 
-# Get rid of EINs that don't appear from at least 2018-2021
+# Make a variable capturing years present for each EIN
 all_ein_data <- all_ein_data %>%
-  mutate(present = ifelse(TaxYr>=2018 & TaxYr<=2021, 1, 0)) %>%
   group_by(Filer.EIN) %>%
-  filter(sum(present)>3) %>%
+  arrange(TaxYr) %>%
+  mutate(allyears = list(unique(TaxYr))) %>%
   ungroup()
+
+observe <- all_ein_data %>%
+  distinct(Filer.EIN, allyears)
+
+# frequency table of which years are included
+freq <- all_ein_data %>% distinct(Filer.EIN, allyears) %>% count(allyears)
 
 # Combine the variables that end with "BusinessNameLine1Txt"
 all_ein_data <- all_ein_data %>%
@@ -36,7 +44,8 @@ all_ein_data <- all_ein_data %>%
 # keep all variables with USAddress in the variable name
 ein_data <- all_ein_data %>%
   select(TaxYr, Filer.EIN, BusinessName1, BusinessName2, Filer.USAddress.AddressLine1Txt, Filer.USAddress.StateAbbreviationCd,
-         Filer.USAddress.ZIPCd, Filer.USAddress.CityNm)
+         Filer.USAddress.ZIPCd, Filer.USAddress.CityNm, allyears)
+
 
 # Work through any missing business names to determine most accurate name
 ein_data <- ein_data %>%
@@ -95,15 +104,20 @@ ein_data <- ein_data %>%
 ein_data <- ein_data %>%
   mutate(BusinessName = ifelse(is.na(BusinessName), paste(BusinessName1, BusinessName2, sep=" "), BusinessName))
 
+# Create indicator for group return
+ein_data <- ein_data %>%
+  mutate(group_return = ifelse(!is.na(BusinessName2) & str_detect(BusinessName2, "GROUP"),1,0)) %>%
[…]

### 2025-11-25 Paper/descriptive_paper.tex
Updates from Overleaf

@@
 \usepackage[backend=biber,style=authoryear,
-sorting=ynt,citestyle=authoryear]{biblatex}
+sorting=nyt,citestyle=authoryear]{biblatex}
 \addbibresource{papercitations.bib}
@@
 		{\small
-    Consolidation in health care has been a heavily discussed topic in research and regulation, as more concentration has been linked to higher prices and lower quality.
+    Consolidation in health care has been a heavily discussed topic in research and regulation, as more concentration is linked to higher prices and lower quality.
 Aside from formal mergers and acquisitions, there is opportunity for hospitals to form affiliations with other competing hospitals through sharing common board members, which has mainly been studied in the context of large public firms.
 This paper establishes the existence of overlapping board members in competing nonprofit hospitals in the US.
-Using a data set constructed from IRS Tax Form 990s, I show that 8\% of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market.
+Using a data set constructed from IRS Tax Form 990s, I show that 8\% of this sample of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market.
+Having an overlapping board member is correlated with quality metrics and investment decisions, though I do not draw a causal link.
 These findings highlight an understudied mechanism through which hospitals form affiliations, potentially affecting patient and market outcomes.
@@
     The presence of overlapping board members across firms, also referred to as board interlock, has been heavily studied in corporate finance, particularly for large publicly traded firms.
-By the Clayton Act of 1914, board overlap among sufficiently large competing firms is illegal.
+Board overlap can facilitate indirect collusion on pricing and output strategies, and thus by the Clayton Act of 1914, board overlap among sufficiently large competing firms is illegal.
 Yet, it is known to exist even among the largest publicly traded firms.
 Researchers have found that board overlap is correlated with governance practices, performance, investment in research and development, and product differentiation (\cite{cai2014board}; \cite{lamb2016ties}; \cite{geng2021does}).
+However, it is still unclear whether board overlap really is anti-competitive, as it could also facilitate valuable knowledge sharing and allow multiple firms to gain from expertise of high quality board members.
 While the Department of Justice has increased scrutiny of board overlap, their focus has remained on tech and private firms (\cite{Morse2023}).
@@
 
-    In this paper, I form networks of hospital board of director members using Tax Form 990s, which requires nonprofits in the US to declare the identities and compensation information of their board.
+    In this paper, I form networks of hospital board of director members using Tax Form 990s, which are yearly declarations of the financial state of nonprofits in the US that also include information about who is leading the firm.
 Extracting these identities and matching to other sources of hospital information, I show that approximately 8\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations.
@@
 That is, a majority of the shared board members serve on directly competing hospitals in the same geographic market.
+In addition, half of the hospital pairs are among one large and one small hospital, and half of the hospital pairs are between two hospitals that do not belong to a system.
 
-    Establishing descriptive evidence of board overlap in hospitals, this paper encourages future research to take seriously this behavior as potentially having downstream effects for patients.
-While formal integration in health care should not be overlooked, there is a call for future research to scrutinize other forms of firm allegiance in policy and regulation.
+    Informed by established research on board overlap and hospital consolidation, I describe categories of hospital behaviors that may be affected by board overlap: price/quality, investment decisions, and differentiation.
+I construct various measures of behavior in these categories, and show how they are correlated with board overlap for small/large and system/independent hospitals.
+While the purpose of this paper is not to establish a causal relationship between board overlap and hospital behavior, there are key observable differences in the average quality and investment decisions of hospitals with and without board overlap that motivate future research on the topic.
+
+     This paper contributes to our understanding of board overlap in a setting not previously considered and our understanding of hospital affiliations apart from formal mergers and acquisitions.
+The findings have implications for policymakers in assessing potential antitrust violations in certain nonprofit settings, and for future research in identifying the causal effects of board overlap in hospitals.
 
@@
 
-    One of the first studies of board overlap argues that firms are not independent, self-sufficient organisms in the economy, but are actually heavily interconnected via overlapping board members (\cite{dooley1969interlocking}).
-Dooley frames board interlock as ``the root of many evils", and finds several aspects of corporations to be correlated with whether they share board members with another corporation: size, financial relationships with other companies, and management control.
-This paper led to a number of studies focusing on how board interlock affects corporate governance, including whether board overlap stretches individual board members too thin (\cite{ferris2003too}; \cite{field2013busy}), contributes to increases in backdating stock options (\cite{bizjak2009option}), or led to convergence in governance strategy and behavior (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}).
+    One of the first studies of board overlap argues that firms are not independent, self-sufficient entities as typically assumed in economics, but are actually heavily interconnected via overlapping board members (\cite{dooley1969interlocking}).
+Dooley frames board interlock as ``the root of many evils", and finds several characteristics to be correlated with the likelihood of sharing a board member with another corporation: size, financial relationships with other companies, and management control.
+This led to a number of studies focusing on how board interlock affects corporate governance, including whether board overlap stretches individual board members too thin (\cite{ferris2003too}; \cite{field2013busy}), contributes to increases in backdating stock options (\cite{bizjak2009option}), or led to convergence in governance strategy and behavior (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}).
 
@@
     Two main studies analyze of the effect of board overlap on financial performance.
-First, \citeauthor{baran2017director} (\citeyear{baran2017director}) leverages entry and exit from the S\&P 500, along with variation in the local supply of directors, and finds that well-connected board members contribute positively to firm value.
-Similarly, leveraging policies in the US that remove barriers to board overlap, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that board overlap leads to higher profitability.
+First, \citeauthor{baran2017director} (\citeyear{baran2017director}) leverages entry and exit from the S\&P 500, along with variation in the local supply of directors, and finds that well-connected board members contribute positively to firm value (\cite{baran2017director}).
+Similarly, leveraging policies in the US that remove barriers to board overlap, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that board overlap leads to higher profitability (\cite{geng2021does}).
+\citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections lead to better financial performance for the acquirer after the transaction (\cite{schonlau2009board}; \cite{cai2012board}).
 
-    Beyond financial performance, several studies explore the relationship between board overlap and firm investment decisions, including mergers and acquisitions, research and development (R\&D), and technology investment.
-\citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections lead to better financial performance for the acquirer after the transaction.
+    Beyond financial performance, several studies explore the relationship between board overlap and firm investment decisions as a mechanism for changes in financial performance.
 One paper examines whether interconnected firms are more likely to invest in the same information technology, and finds a positive association between board overlap and investment in IT among firms with active boards, indicating that timely communication is an important factor in whether board overlap matters (\cite{cheng2021social}).
-Finally, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that lower research and development expenditure and more product differentiation are key mechanisms through which overlapping firms improve profitability.
+Finally, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that lower research and development expenditure and more product differentiation are key mechanisms through which overlapping firms improve profitability (\cite{geng2021does}).
 
@@
 I draw from this literature to inform potential outcomes of interest when considering board overlap in hospitals: financial performance, mergers and acquisitions, investment, and product differentiation.
-I discuss in Section \ref{sec:hospbehaviors} how these outcomes related to hospitals specifically.
+I discuss in Section \ref{sec:hospbehaviors} how these outcomes relate to hospitals specifically.
 
@@
 Theoretical models predict that common ownership will lead to lower equilibrium output and higher markups through firms placing non-zero weight on competitors' profit (\cite{rubinstein1983competitive}; \cite{rotemberg1984financial}; \cite{azar2012new}).
-Empirical evidence has grown tremendously since the seminal studies by \citeauthor{he2017product} (\citeyear{he2017product}) and \citeauthor{azar2018anticompetitive} (\citeyear{azar2018anticompetitive}), who found that common ownership leads to greater consolidation (acquisitions, joint ventures, strategic alliances), and that common ownership led to higher prices in the airline industry, respectively.
+Empirical evidence has grown tremendously since the seminal studies by \citeauthor{he2017product} (\citeyear{he2017product}) and \citeauthor{azar2018anticompetitive} (\citeyear{azar2018anticompetitive}), who found that common ownership leads to greater consolidation (acquisitions, joint ventures, strategic alliances), and that common ownership led to higher prices in the airline industry, respectively (\cite{he2017product}; \cite{azar2018anticompetitive}).
 However, these findings have been critiqued for how they measure common ownership and endogeneity concerns (\cite{kennedy2017competitive}; \cite{lewellen2021does}).
@@
 Finally, this literature introduces important concerns about the endogenous choice of owners or board members.
-While this study does not seek to causally estimate the effects of board overlap in hospitals, this is an important consideration for future research.
+While I do not seek to causally estimate the effects of board overlap in hospitals in this paper, this is an important consideration for future research.
 
-    \subsection{Consolidation in Hospitals}
+    \subsection{Consolidation in Hospitals}\label{sec:hospconsol}
 
     There is a rich empirical literature examining the effects of consolidation in health care that includes both vertical and horizontal integration between hospitals, practices, and physicians.
-As the focus of this paper is affiliation among competing hospitals, I draw on the literature examining hospital consolidation through mergers and systems to inform hospital behaviors that might be correlated with board overlap but are unique to the hospital setting and thus weren't discussed in Section \ref{sec:boardoverlaplit}.
+As the focus of this paper is affiliation among competing hospitals, I draw on the literature examining hospital consolidation in particular through mergers and systems.
+This will inform hospital behaviors that might be correlated with board overlap but are unique to the hospital setting and thus weren't discussed in Section \ref{sec:boardoverlaplit}.
 The main outcomes of interest in this area of research have centered around price and quality of care.
@@
 From 2000-2020, the occurrence of hospital mergers increased dramatically, where a quarter of hospital markets included zero independent hospitals by 2020 (\cite{ElevanceHealth2023}).
-This increased concentration has led to a vast number of studies estimating its effect on quality, operating costs, and prices.
+This increase in concentration has led to a vast number of studies estimating its effect on quality, operating costs, and prices.
 
@@
     There is mixed evidence on whether such gains in efficiency translate to better quality of care.
-Several studies comparing hospitals who became a part of a system with observably similar hospitals that did not merge find little to no effect on readmission or mortality rates on average (\cite{haas2011mergers}; \cite{beaulieu2020changes}).
+Several studies comparing hospitals that became a part of a system with observably similar hospitals that did not merge find little to no effect on readmission or mortality rates on average (\cite{haas2011mergers}; \cite{beaulieu2020changes}).
 However, one study finds that among heart disease patients, treatment intensity and mortality rates both increase after hospital consolidation (\cite{hayford2012impact}), and another finds increased readmission rates across all types of patients (\cite{andreyeva2024corporatization}).
@@
     A select number of studies investigate other dimensions that could be affected by more system affiliation over time.
-In a slightly different setting of acquisitions of dialysis centers, \citeauthor{eliason2020acquisitions} (\citeyear{eliason2020acquisitions}) finds that decreases in quality are driven primarily by the convergence of treatment styles among acquired centers and their acquirer.
+In a slightly different setting of acquisitions of dialysis centers, \citeauthor{eliason2020acquisitions} (\citeyear{eliason2020acquisitions}) finds that decreases in quality are driven primarily by the convergence of treatment styles among acquired centers and their acquirer (\cite{eliason2020acquisitions}).
 Additional studies have suggested that consolidated service offerings is a potential mechanism for effects on quality (\cite{mariani2022impact}).
 Finally, \citeauthor{desai2023hospital} (\citeyear{desai2023hospital}) examine admission patterns of low-income patients as a result of increased hospital concentration, and find that Medicaid admissions decrease as a result of increased concentration.
-That is, highly concentrated markets redistribute patients within the market (\cite{desai2023hospital}).
+That is, highly concentrated markets exhibit redistribution patterns of patients within the market (\cite{desai2023hospital}).
 
-    This literature informs ways that hospital affiliations could affect firm behaviors and outcomes specific to hospitals.
+    This literature informs ways that hospital affiliations might affect behaviors and outcomes specific to hospitals.
 In addition to the behaviors discussed in Sections \ref{sec:boardoverlaplit} and \ref{sec:commonownerlit}, measures of quality of care, treatment styles, service offerings, and admission patterns could also be important aspects driven by affiliation with other hospitals.
@@
 
-    The main purpose of this paper is to document the prevalence of board overlap in US hospitals in Section \ref{sec:document}.
-While I do not establish any causal relationship between board overlap and hospital behaviors, I consider the literature on board overlap, common ownership, and hospital consolidation discussed in Section \ref{sec:relatedlit} to inform a set of hospital behaviors that may be relevant for future research.
+    The main purpose of this paper is to document the prevalence of board overlap in US hospitals in recent years.
+While I do not establish any causal relationship between board overlap and hospital behaviors, I consider the literature on board overlap, common ownership, and hospital consolidation discussed in Sections \ref{sec:boardoverlaplit}-\ref{sec:hospconsol} to inform a set of hospital behaviors that may be relevant for future research.
 I divide these behaviors into three categories: price/quality, investment decisions, and differentiation.
@@
 
-    It seems more likely that quality of care could be impacted by overlapping board members among hospitals.
+    It may be more likely that quality of care could be impacted by overlapping board members among hospitals.
 There are direct actions that could lead to changes in quality such as convergence in treatment styles or investment in technology, which I will discuss in detail in the following sections.
 There is also an indirect path through which overlapping board members might impact quality, which is that hospitals might compete for the same pool of experienced board members.
-If the board members with the most expertise are in high demand, sharing that expertise among multiple hospitals could facilitate greater improvements.
-The most commonly used measurements of quality of care in hospitals are mortality and readmission rates.
+If the board members with the most expertise are in high demand, sharing that expertise among multiple hospitals could facilitate greater improvements in quality.
 
@@
 First, hospitals with board shared board members might be more likely to merge in the future.
-Second, sharing board members could facilitate engaging in the same types of technology or capital.
+Second, sharing board members could facilitate engagement in the same types of technology or capital.
 Therefore, an area of future research is to look at outcomes such as health information technology investments, specific medical technology advancements in the hospital, and capital investments.
@@
     Finally, previous research has shown that product differentiation is associated with board overlap in other industries.
-That is, sharing information among competing hospitals can lead to more specialization.
+That is, sharing information among competing firms can lead to more specialization.
 In hospitals, there are two distinct types of differentiation: either through services offered, or the types of patients admitted.
 Both of these behaviors are key findings in the hospital merger literature, indicating that affiliations through board members could also facilitate changes in these areas.
-In future research, one could think about measuring the likelihood of specialized services being offered, such as a Neonatal Intensive Care Unit or cardiac catheterization laboratory, where hospitals might shift investments towards services not offered by their affiliated hospital.
+In future research, one could think about measuring the likelihood of specialized services being offered, where hospitals might shift investments towards services not offered by their affiliated hospital.
 Additionally, patient composition in the hospital, measured by the payer mix of admitted patients or complexity of their conditions, could be compared among hospitals with and without shared board members.
@@
     I define a hospital as having an overlapping board member if any of the individuals on their board are also on the board of a hospital in the same hospital referral region (HRR).
-HRRs are market designations created by Dartmouth Atlas using patient referral data which are meant to capture geographic regions of the average patient's choice set of health care facilities.
+HRRs are market designations created by Dartmouth Atlas using patient referral data which are meant to capture geographic regions of the average patient's choice set of hospitals.
 I make this restriction first because board overlap is predicted to matter among competing firms, and hospitals compete for patients within the same geographic region.
@@
 Therefore, I rely on variation in board overlap that does not include system affiliation.
-Therefore, any hospitals with board overlap in my sample are otherwise unaffiliated.
+Thus, any hospitals with board overlap in my sample are otherwise unaffiliated.
 There still may exist board overlap across systems.
 That is, I still consider hospital A belonging to System 1 as eligible to share board members with hospital B in System 2.
+The final sample consists of approximately 1500 nonprofit hospitals from 2017-2022.
 
@@
 
-    First, I present a summary statistics table showing characteristics of all hospitals in the final sample, which consists of 1,466 nonprofit hospitals, almost all of which are adult general hospitals.
+    First, I present a summary statistics table showing characteristics of all hospitals in the final sample, which consists of 1,467 nonprofit hospitals, almost all of which are adult general hospitals.
 They have, on average, 178 beds and 14 board members.
@@
     Eight percent of hospitals in the sample have overlapping board members with an otherwise unaffiliated hospital.
-In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decrease from 129 affiliated hospitals in 2017 to 114 affiliated hospitals in 2021.
+In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decline from 129 affiliated hospitals in 2017 to 114 affiliated hospitals in 2021.
 I include 2022 values, but there are less total firms in the sample due to lag in tax forms being made available publicly.
-As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population.
-Therefore, in Figure \ref{fig:connected_HRR_percent}, I show the prevalence of having shared board members within a hospital market.
-Of the 306 HRRs in the US, approximately 17\% have at least one pair of board affiliated hospitals within the market.
-This proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation.
-This is likely due to increased system affiliations, which mechanically decreases board overlap according to the definition I use here.
 
@@
 
+    As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population.
+In Figure \ref{fig:connected_HRR_percent}, I show the prevalence of having shared board members within a hospital market.
+Of the 306 HRRs in the US, approximately 17\% have at least one pair of board affiliated hospitals within the market.
+This proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation.
+This is likely due to increased system affiliations, which mechanically decreases board overlap since I do not include hospitals in the same system as eligible for having board overlap.
+
     \begin{figure}[ht!]
@@
 
-    Next, I present geographic distribution of these pairs in Figure \ref{fig:connected_maps}.
+    Next, I present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}.
 The blue lines represent hospitals that share a common board member, and the red dots are all other hospitals.
@@
 
-    I now shift to examining characteristics of hospitals with board overlap.
-I present the proportion of the type of hospital pairs in Table \ref{tab:hospital_pair_types}.
-Each hospital is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data.
-Two-thirds of board affiliated pairs are between two adult general hospitals, and 10\% of pairs are between a specialty and a general hospital.
-Thus, for the majority of hospitals that have board interlock, the hospital they share a board member with is a direct competitor.
+    I now shift to examining characteristics of hospitals that share board members.
+I consider pairs of connected hospitals and characteristics of the pair in Table \ref{tab:hospital_pair_types}.
+Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data.
+Additionally, the median bed size of all hospitals in the sample is 103, and I define hospitals as being either small or large depending on whether they fall below or above this median.
 
@@
 
-    There is also significant variation in the ownership type of hospitals within affiliated pairs.
+    Two-thirds of board affiliated pairs are between two adult general hospitals, and 10\% of pairs are between a specialty and a general hospital.
+Thus, for the majority of hospitals that have board interlock, the hospital they share a board member with is a direct competitor.
+Additionally, there is significant variation in the ownership type of hospitals within affiliated pairs.
 Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system.
 The lowest proportion of shared board member pairs is made up of two hospitals belonging to different systems.
+
+    There is also variation in whether the hospitals are the same size or not.
+Half of the hospital pairs consist of one small and one large hospital, and the other half are split evenly between two small or two large hospitals.
+However, the average difference in the number of beds of the two hospitals in a pair is very large: an average difference of 254 beds.
+This suggests that for hospitals of opposite size, the difference is drastic.
+The average distance between hospitals that share common board members is 35 miles, or 57 kilometers.
 In total, there are 217 hospitals that share a board member with at least one other hospital at some point in the sample.
 
-    In Table
+    \subsection{Correlations with Hospital Behaviors}
 
-    I now shift towards examining summary statistics of hospitals with different types of affiliation relationships.
-First, in Table \ref{tab:hospital_general_summarystats}, I focus on general hospitals in four categories: board affiliated with another general hospital, board affiliated with a specialty hospital, unaffiliated through board members but belonging to a system, and completely unaffiliated.
-I present means of hospitals in each of these categories for the number of beds, the number of patients seen, whether it is an academic medical center, whether the hospital is located in a metropolitan area, and whether the hospital belongs to a system.
-On these dimensions, hospitals affiliated with other general hospitals (column 1) are relatively similar to unaffiliated hospitals that are part of a system (column 3).
-Unaffiliated hospitals that are not part of a system are slightly smaller in terms of bed size and number of patients seen, and less likely to be in a metropolitan area.
-The hospitals in column (2) are outliers, with a large number of beds and patients seen, and mainly in metropolitan areas.
-There are only 12 hospitals in this sample that are likely very different from the average general hospital.
+    I now merge the hospital board data to other publicly available data sets to examine descriptive differences in the hospital behaviors discussed in Section \ref{sec:hospbehaviors} for hospitals with and without board overlap.
+The broad categories of hospital behaviors that might be affected by affiliation through shared board members are price/quality, investment decisions, and differentiation, which are discussed in Section \ref{sec:hospbehaviors}.
+While I do not capture every aspect of hospital behavior, I select a number of variables that fall into these categories to examine.
 
-    \import{Objects}{hospital_general_summarystats.tex}
+    First, from the Center for Medicare and Medicaid Services (CMS) Hospital Compare data, I merge information on each hospital's overall star rating.
+This rating combines measures of quality into one index that ranges from 1 to 5: readmission rates, mortality rates, safety, timeliness of care, effectiveness of care, and patient experience.
+This is a very aggregated measure, but may capture big-picture differences in quality.
 
-    I also present means of various measures of patient demographics, include the number of Medicare and Medicaid patients normalized by total beds, the total number of Medicare patients from CMS, the average Medicare patient age, and the average Medicare patient risk score.
-These measures yield similar descriptive results, that the unaffiliated hospitals that are not part of a system are slightly smaller, but they see similar patients in terms of age and complexity.
-Finally, I show means of variables to capture aspects of the service offerings of the different types of hospitals.
-First, from the AHA survey, I include an indicator for whether the hospital offers services in a Neonatal Intensive Care Unit (NICU) or Cardiac catheterization lab, both specialized and profitable services.
-General hospitals that are affiliated with specialty hospitals are the most likely to offer these services.
-Hospitals without any board member affiliations but belonging to a system are more likely to offer these services than unaffiliated hospitals that do not belong to a system.
-Hospitals that are affiliated with another general hospitals fall somewhere in between these two, with slightly less likelihood than the unaffiliated hospitals belonging to a system.
-I also include a measure of the concentration of bed allocation, where a higher concentration indicates that a hospital's bed capacity is concentrated in fewer service categories, whereas a lower concentration suggests a more diversified service offering.
-I also compute a similar concentration metric using the percent of patients in several service categories from the CMS Provider Utilization Files.
-For both measures, there seems to be little variation in service concentration across affiliation types.
+    Next, I include data on different types of investments made into the hospital from the Healthcare Cost Reporting Information System (HCRIS) data.
+These variables include the dollar amounts spent on health information technology purchases, fixed equipment, and movable equipment.
 
-    Next, I present the same summary statistics for specialty hospitals with different types of affiliations: affiliated with a general hospital, unaffiliated through board but belonging to a system, and unaffiliated through board or system, shown in Table \ref{tab:hospital_specialty_summarystats}.
-Overall, there are less specialty hospitals included in the sample, and very few of them are affiliated through board members.
-Specialty hospitals are more concentrated than general hospitals, as expected.
-In the AHA measure of concentration, all specialty hospitals are similarly concentrated.
-However, when using the services included in the CMS files, the concentration of unaffiliated system hospitals is much lower than that of any other specialty hospital.
-This could be due to the limited services included in the CMS data.
-Since this is such a small and specialized sample of hospitals, I do not include them in the main analysis.
+    There are a plethora of ways one could measure differentiation in hospitals.
+I focus on two categories of differentiation: service offerings and patient population.
+To measure different service offerings, I first construct a measure of concentration of beds devoted to categories of services in the AHA survey.
+For services $j$ in the list of all possible services\footnote{General medical and surgical (adult), General medical and surgical (pediatric), Obstetric care, Medical/surgical intensive care, Cardiac intensive care, Neonatal intensive care, Neonatal intermediate care, Pediatric intensive care beds, Burn care, Other Special Care, Other intensive care, Physical Rehabilitation care, Alcohol/drug abuse or dependency inpatient care, Psychiatric care, Skilled nursing care, Intermediate nursing care, Acute long term care, Other long-term care, Other care}, the concentration of bed allocation is defined as
 
-    \import{Objects}{hospital_specialty_summarystats.tex}
+    $$concentration_{bed} = \sum_{j}\left(\frac{beds_j}{total beds}\right)^2.$$
+
+    \noindent That is, a higher bed concentration indicates more specialization in bed allocation, where a lower concentration indicates more services spread out in bed allocations.
+Additionally, I include a second concentration measure that relies less on the facility and more on the conditions patients are treated with.
+The CMS Provider Utilization files contain measures of the percent of patients seen for different conditions.
+Therefore, for services $k$ in the list of conditions in the CMS data\footnote{Percent Of Beneficiaries Identified With Combined Cancer Flag For 6 Cancer Indicators, Percent Of Beneficiaries Identified With Chronic Obstructive Pulmonary Disease, Percent Of Beneficiaries Identified With Chronic Kidney Disease, Percent Of Beneficiaries Identified With Heart Failure And Non-Ischemic Heart Disease}, I define the concentration of patient conditions in each hospital as
+
+    $$HHI_{patient} = \sum_{k}\left(\frac{patients_k}{total patients}\right)^2.$$
+
+    Finally, I include indicators for whether the hospital has a NICU or Cath Lab, both profitable services for hospitals to offer.
+I also consider differentiation by the patient population admitted to the hospital.
+There is information in the HCRIS data about how many discharges are for patients on Medicaid or Medicare.
+There is also information in the CMS Provider Utilization Files about the average risk score of Medicare patients seen in the hospital.
+I include all of these variables to capture something about the aggregate patient population being treated in different hospitals.
+
+    In Table \ref{size_characteristics}, I present averages of all of these variables for four categories of hospitals: whether they are small (less than 103 beds) or large (greater than 103 beds), and whether they share board members with another competing hospital.
+I limit this sample to only general hospitals for relevant comparison.
+
+   \import{Objects}{size_characteristics}
+
+   Small hospitals with and without shared board members have a very similar average star rating of 3.4, while large hospitals with shared board members have a slightly smaller average star rating than large hospitals without board overlap.
+In all categories of investment, small hospitals are relatively similar, with slightly more investment in health IT and fixed equipment for board affiliated small hospitals.
+Large hospitals are also generally similar in how they invest in health IT and fixed equipment relative to how much they spend on average, but hospitals with shared board members spend significantly more on movable equipment than hospitals without board overlap.
+
+   There do not seem to be meaningful differences in the concentration of services offered or conditions treated among hospitals with and without board overlap, whether they are small or large.
+Similarly, there are no clear differences in the patient composition of hospitals with and without board overlap.
+This is consistent even among the facilities of these hospitals, with no distinct differences in whether hospitals have a NICU or Cath lab based on board overlap.
+
+   Another important level of variation in this sample of hospitals could be whether they belong to a system or not, as this often dictates a number of the behaviors discussed.
+Therefore, I present a similar table of means for hospitals with and without board overlap that do or do not have a system affiliation in Table \ref{sys_characteristics}.
+
+   \import{Objects}{system_characteristics}
+
+    Among hospitals that do not belong to a system, those without board overlap have a slightly higher quality rating than those without board overlap.
+However, the opposite is true among hospitals that do belong to a system, where those with board overlap have a higher quality rating.
+The most significant difference in investment is among hospitals that belong to a system for investment in movable equipment.
+Those with board overlap invest over 20 million in movable equipment, while those without board overlap only invest 8.3 million.
+The other comparisons of investment show no drastic differences.
+In-system and out-of-system hospitals both show higher levels of concentration of services when they do not have board overlap.
+However, the patient composition is similar within the categories of system affiliation.
+Independent hospitals that do not have board overlap are more likely to have a Cath lab, but otherwise hospitals are similar in in service offerings.
+
+    While these statistics do not establish a causal link between board overlap and hospital behavior, they highlight key differences across hospital categories.
+Notably, hospitals with overlapping board members exhibit distinct patterns in quality and investment decisions.
+These differences underscore the need for further research to explore board overlap as a potential mechanism for affiliation and behavioral changes in health care.
+
+    \section{Conclusion}
+
+    This study contributes to the growing literature on hospital governance and market structure by examining the extent and implications of board member overlap across hospitals.
+By identifying the prevalence of shared board membership, I provide new insights into how governance ties may influence competition, decision-making, and resource allocation in the hospital sector.
+This paper sheds light on the role of board interlocks in potentially shaping hospital behavior, offering an important perspective on the non-ownership mechanisms that can link hospitals within the same market.
+
+    Using a comprehensive dataset of board members in nonprofit hospitals from 2017-2022, I document the characteristics of hospitals with shared board members and analyze how these overlaps have evolved over time.
+I identify connections between hospitals and examine the patterns of overlap by hospital type, market, and geographic region.
+The results indicate that a significant share of hospitals are connected through board interlocks, and a significant portion of hospital markets contain board overlap within the market.
+Further, I show geographic distributions of overlap and characterize connected hospital pairs by important characteristics such as ownership, size, and specialty.
+
+    These findings hold important implications for policymakers and future research.
+Regulators and antitrust authorities should consider the potential competitive effects of board member overlap when assessing hospital market dynamics, particularly in regions with high consolidation or limited provider choice.
+Future research could further investigate how these governance connections influence hospital pricing, quality of care, and patient outcomes.
+Expanding this work to explore causal relationships and the broader economic consequences of board interlocks would provide valuable insights for competition in health care.
 

### 2025-11-25 Paper/paper_2025.tex
Updates from Overleaf

--- 2025-11-25 Paper/descriptive_paper.tex
@@
 
-    Consolidation in health care remains a forefront concern for regulators in the US, as both horizontal and vertical integration have substantially increased over the last several decades (\cite{levinson2024ten}).
+    % Consolidation in health care remains a forefront concern for regulators in the US, as both horizontal and vertical integration have substantially increased over the last several decades (\cite{levinson2024ten}).
 Researchers have documented higher prices (\cite{brot2024pays}), and mixed effects on quality (\cite{kessler2000hospital}; \cite{cooper2011does}) as a result of health care becoming more concentrated through mergers and acquisitions.
@@
 In this paper, I document the extent of overlapping board members in nonprofit hospitals in the US, and explore avenues of future research to investigate
-    potential anti-competitive effects of this type of affiliation.
+    % potential anti-competitive effects of this type of affiliation.
 
-    The presence of overlapping board members across firms, also referred to as board interlock, has been heavily studied in corporate finance, particularly for large publicly traded firms.
+    % The presence of overlapping board members across firms, also referred to as board interlock, has been heavily studied in corporate finance, particularly for large publicly traded firms.
 Board overlap can facilitate indirect collusion on pricing and output strategies, and thus by the Clayton Act of 1914, board overlap among sufficiently large competing firms is illegal.
@@
 
-    In this paper, I form networks of hospital board of director members using Tax Form 990s, which are yearly declarations of the financial state of nonprofits in the US that also include information about who is leading the firm.
+    % In this paper, I form networks of hospital board of director members using Tax Form 990s, which are yearly declarations of the financial state of nonprofits in the US that also include information about who is leading the firm.
 Extracting these identities and matching to other sources of hospital information, I show that approximately 8\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations.
@@
 
-    Informed by established research on board overlap and hospital consolidation, I describe categories of hospital behaviors that may be affected by board overlap: price/quality, investment decisions, and differentiation.
+    % Informed by established research on board overlap and hospital consolidation, I describe categories of hospital behaviors that may be affected by board overlap: price/quality, investment decisions, and differentiation.
 I construct various measures of behavior in these categories, and show how they are correlated with board overlap for small/large and system/independent hospitals.
@@
 
-     This paper contributes to our understanding of board overlap in a setting not previously considered and our understanding of hospital affiliations apart from formal mergers and acquisitions.
+    %  This paper contributes to our understanding of board overlap in a setting not previously considered and our understanding of hospital affiliations apart from formal mergers and acquisitions.
 The findings have implications for policymakers in assessing potential antitrust violations in certain nonprofit settings, and for future research in identifying the causal effects of board overlap in hospitals.
@@
 
-    \section{Documenting Overlapping Board Members in US Hospitals}\label{sec:document}
+    \section{Background/Setting}
 
-    \subsection{Data Construction}
+    \section{Data Construction}
 
-   I compile data on nonprofit hospital board of directors from publicly available Tax Form 990s, which all sufficiently large nonprofits must file with the Internal Revenue Service (IRS) each year.
+   I compile data on nonprofit hospital board of directors in years 2017-2022 from publicly available Tax Form 990s, which all sufficiently large nonprofits must file with the Internal Revenue Service (IRS) each year.
 These forms contain a section in which nonprofits declare their board, executives, and highest compensated employees.
@@
 I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital.
-PDF versions of the 990s are made publicly available by Nonprofit Explorer dating back at least 20 years.
-However, using optical character recognition text extraction methods on these PDFs is both time consuming and can lead to small errors in names extracted.
-Since I rely heavily on the accuracy of matching individuals, I analyze the data from 2017 onward, when the IRS began publishing XML versions of these files, which ensures accurate names that can be matched across hospitals and time.
-This also provides a more current picture of the prevalence of board overlap in this industry.
 I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees".
 I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics.
-There are 2,096 hospitals in the initial extraction of firms and identities.
+The initial extraction yields 2,096 nonprofit hospitals.
 
@@
 Specifically, I compute the Jaro-Winkler distance between hospital names in each data set.
-Then, I record matches with sufficiently similar names and addresses.
-I have a sample of 1,623 hospitals in the tax data that can be linked to the AHA survey.
+Then, I record matches with sufficiently similar names and addresses, yielding a sample of 1,836 matched hospitals.
 
     I limit the individuals in the sample to those on the board of directors, excluding executives that also serve on the board.
-I clean the board member names by removing common titles (Dr., Reverand, etc.), and combining names within the same firm that have slight spelling differences.
-For example, if John Matthews is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person.
-Finally, to avoid issues in whether first or last names are listed first, I sort each individual's names alphabetically.
+I then record self-reported information about the board members in the data, including whether they include a title for medical doctor, nurse, or having a degree in health administration.
+I also record whether the firm records that they are a non-voting member of the board (though this is very rare).
+Several steps are necessary to ensure the names are clean and consistent across firms to identify people across hospitals.
+First, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency.
+Second, I combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common.
+For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person.
+I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).
 
-    I define a hospital as having an overlapping board member if any of the individuals on their board are also on the board of a hospital in the same hospital referral region (HRR).
-HRRs are market designations created by Dartmouth Atlas using patient referral data which are meant to capture geographic regions of the average patient's choice set of hospitals.
-I make this restriction first because board overlap is predicted to matter among competing firms, and hospitals compete for patients within the same geographic region.
-Additionally, hospital mergers have been shown to have a more significant effect when the entities involved are located geographically close (\cite{cooper2019price}).
-Therefore, the relevant measure for board overlap in hospitals should be those within the same market.
+    I create a list of overlapping board member candidates by identifying any names that occur on multiple hospital boards in the same year.
+However, it is very common for hospitals within a system to largely share the same board.
+This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated.
+Therefore, I define the relevant type of overlap as someone being on the board of multiple hospitals that are otherwise \textit{unaffiliated}.
+Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system as they do.
+Further, In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close \cite{cooper2019price}.
+Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market.
+All other hospitals have no board overlap with otherwise unaffiliated hospitals.
+I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals.
 
-    Another restriction that I place for two hospitals to share overlapping board members is that they must not be affiliated by system or network.
-It is common for within-system hospitals to either share common board members or be governed by the same over-arching board.
-However, systems are already formally affiliated, and this affiliation can dictate many of the behaviors I discuss in Section \ref{sec:hospbehaviors}.
-Therefore, I rely on variation in board overlap that does not include system affiliation.
-Thus, any hospitals with board overlap in my sample are otherwise unaffiliated.
-There still may exist board overlap across systems.
-That is, I still consider hospital A belonging to System 1 as eligible to share board members with hospital B in System 2.
-The final sample consists of approximately 1500 nonprofit hospitals from 2017-2022.
+    The final sample limitation I make for simplicity is to remove hospitals that gain and then lose overlapping board members in the sample.
+Since treatment will be turning on overlapping board member, I remove those who "lose" their treatment.
+
+    \subsection{Documenting Overlapping Board Members}
+
+    The full sample of hospitals
 
@@
 
-    \import{Objects}{all_summarystats.tex}
+    %\import{Objects}{all_summarystats.tex}
 

### 2025-11-26 Paper/paper_2025.tex
Updates from Overleaf

@@
 
+    \import{Objects}{board_people.tex}
+
     I create a list of overlapping board member candidates by identifying any names that occur on multiple hospital boards in the same year.
@@
 Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system as they do.
-Further, In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close \cite{cooper2019price}.
+Further, In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}).
 Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market.

### 2025-12-05 Paper/paper_2025.tex
Updates from Overleaf

@@
 
-	\linespread{1.2}\title{\vspace{-0.5in} Overlapping Board Members in the Hospital Industry}
+	\linespread{1.2}\title{\vspace{-0.5in} Unobserved Firm Affiliations: Evidence from Hospital Board Overlap}
 
@@
 		{\small
-    Consolidation in health care has been a heavily discussed topic in research and regulation, as more concentration is linked to higher prices and lower quality.
-Aside from formal mergers and acquisitions, there is opportunity for hospitals to form affiliations with other competing hospitals through sharing common board members, which has mainly been studied in the context of large public firms.
-This paper establishes the existence of overlapping board members in competing nonprofit hospitals in the US.
-Using a data set constructed from IRS Tax Form 990s, I show that 8\% of this sample of nonprofit hospitals have overlapping board members with an otherwise unaffiliated hospital, and 17\% of hospital referral regions have overlapping boards within the market.
-Having an overlapping board member is correlated with quality metrics and investment decisions, though I do not draw a causal link.
-These findings highlight an understudied mechanism through which hospitals form affiliations, potentially affecting patient and market outcomes.
+
 		} 
@@
 
-    % Consolidation in health care remains a forefront concern for regulators in the US, as both horizontal and vertical integration have substantially increased over the last several decades (\cite{levinson2024ten}).
-Researchers have documented higher prices (\cite{brot2024pays}), and mixed effects on quality (\cite{kessler2000hospital}; \cite{cooper2011does}) as a result of health care becoming more concentrated through mergers and acquisitions.
-There is significant antitrust scrutiny in this sector, and proposed mergers are regularly denied due to expected harmful effects to competition (\cite{Meyer2022}).
-Firms in other industries have been shown to engage in methods of informal affiliation besides formal mergers and acquisitions.
-For example, instances of board overlap in public firms have been cited to violate the Clayton Act due to their anti-competitive nature.
-Yet, despite concerns of low competition, this type of connectedness among firms has not been given much consideration in health care.
-In this paper, I document the extent of overlapping board members in nonprofit hospitals in the US, and explore avenues of future research to investigate
-    % potential anti-competitive effects of this type of affiliation.
+    Antitrust enforcement in the US has been particularly relevant in recent years as price gouging and consolidation has increased.
 
-    % The presence of overlapping board members across firms, also referred to as board interlock, has been heavily studied in corporate finance, particularly for large publicly traded firms.
-Board overlap can facilitate indirect collusion on pricing and output strategies, and thus by the Clayton Act of 1914, board overlap among sufficiently large competing firms is illegal.
-Yet, it is known to exist even among the largest publicly traded firms.
-Researchers have found that board overlap is correlated with governance practices, performance, investment in research and development, and product differentiation (\cite{cai2014board}; \cite{lamb2016ties}; \cite{geng2021does}).
-However, it is still unclear whether board overlap really is anti-competitive, as it could also facilitate valuable knowledge sharing and allow multiple firms to gain from expertise of high quality board members.
-While the Department of Justice has increased scrutiny of board overlap, their focus has remained on tech and private firms (\cite{Morse2023}).
-Board overlap does occur in health care with instances of organizations denied merger permission due to boards not acting independently (\cite{huberfeld2006tackling}), of hospitals declaring informal affiliations through leadership (\cite{barnett_babcock_2012}), and of life sciences companies with substantial documented overlap (\cite{manjunath2024illegal}).
-However, little is known about how prevalent board overlap is in the hospital industry, and whether it is correlated with other hospital characteristics.
+    Antitrust legislation creates an additional hurdle for large-scale anti-competitive behavior, but there are other types of firm affiliations that typically elude antitrust scrutiny such as common ownership, joint ventures, and other types of affiliation/partnerships.
+By and large, firms form partnerships with other firms, often competitors, without needing to seek approval.
+This has started to raise concern recently, and states have even begun regulating smaller-scale affiliations in recent years.
+Horizontal mergers and consolidation have rightfully received a lot of attention in the literature, often linked to something??
+However, we know little about the effects of smaller-scale affiliations, which could carry some of the same implications as consolidation proposals scrutinized for their ant-competitive nature.
 
-    % In this paper, I form networks of hospital board of director members using Tax Form 990s, which are yearly declarations of the financial state of nonprofits in the US that also include information about who is leading the firm.
-Extracting these identities and matching to other sources of hospital information, I show that approximately 8\% of the nonprofit hospitals in the sample are affiliated with another hospital in the same market through shared board members, even excluding system or network affiliations.
-Among hospitals that share board members, two thirds of the hospital pairs are between two general adult hospitals.
-That is, a majority of the shared board members serve on directly competing hospitals in the same geographic market.
-In addition, half of the hospital pairs are among one large and one small hospital, and half of the hospital pairs are between two hospitals that do not belong to a system.
-
-    % Informed by established research on board overlap and hospital consolidation, I describe categories of hospital behaviors that may be affected by board overlap: price/quality, investment decisions, and differentiation.
-I construct various measures of behavior in these categories, and show how they are correlated with board overlap for small/large and system/independent hospitals.
-While the purpose of this paper is not to establish a causal relationship between board overlap and hospital behavior, there are key observable differences in the average quality and investment decisions of hospitals with and without board overlap that motivate future research on the topic.
-
-    %  This paper contributes to our understanding of board overlap in a setting not previously considered and our understanding of hospital affiliations apart from formal mergers and acquisitions.
-The findings have implications for policymakers in assessing potential antitrust violations in certain nonprofit settings, and for future research in identifying the causal effects of board overlap in hospitals.
-
-    \section{Related Literature}\label{sec:relatedlit}
-
-    \subsection{Overlapping Board Members}\label{sec:boardoverlaplit}
-
-    One of the first studies of board overlap argues that firms are not independent, self-sufficient entities as typically assumed in economics, but are actually heavily interconnected via overlapping board members (\cite{dooley1969interlocking}).
-Dooley frames board interlock as ``the root of many evils", and finds several characteristics to be correlated with the likelihood of sharing a board member with another corporation: size, financial relationships with other companies, and management control.
-This led to a number of studies focusing on how board interlock affects corporate governance, including whether board overlap stretches individual board members too thin (\cite{ferris2003too}; \cite{field2013busy}), contributes to increases in backdating stock options (\cite{bizjak2009option}), or led to convergence in governance strategy and behavior (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}).
-
-    The literature thus establishes a strong link between board interlock and governance practices.
-However, the relationship between board overlap and firm behavior, particularly its impact on competition, remains less understood.
-Theoretically, the direction of this effect is ambiguous.
-On one hand, board interlock is often viewed as an anti-competitive practice that enables implicit collusion.
-Alternatively, firms may seek highly experienced and connected individuals for their boards, leading to valuable knowledge sharing and enhanced decision-making.
-Most studies in this area have focused on the financial performance implications of board overlap, with firm behavior considered primarily as a mediating mechanism.
-
-    Two main studies analyze of the effect of board overlap on financial performance.
-First, \citeauthor{baran2017director} (\citeyear{baran2017director}) leverages entry and exit from the S\&P 500, along with variation in the local supply of directors, and finds that well-connected board members contribute positively to firm value (\cite{baran2017director}).
-Similarly, leveraging policies in the US that remove barriers to board overlap, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that board overlap leads to higher profitability (\cite{geng2021does}).
-\citeauthor{schonlau2009board} (\citeyear{schonlau2009board}) and \citeauthor{cai2012board} (\citeyear{cai2012board}) analyze financial performance after mergers and acquisitions of publicly traded firms with connected boards, finding that firms with board overlap are more likely to be involved in acquisitions, and that specific types of board connections lead to better financial performance for the acquirer after the transaction (\cite{schonlau2009board}; \cite{cai2012board}).
-
-    Beyond financial performance, several studies explore the relationship between board overlap and firm investment decisions as a mechanism for changes in financial performance.
-One paper examines whether interconnected firms are more likely to invest in the same information technology, and finds a positive association between board overlap and investment in IT among firms with active boards, indicating that timely communication is an important factor in whether board overlap matters (\cite{cheng2021social}).
-Finally, \citeauthor{geng2021does} (\citeyear{geng2021does}) finds that lower research and development expenditure and more product differentiation are key mechanisms through which overlapping firms improve profitability (\cite{geng2021does}).
-
-    Despite these insights, significant gaps remain in our understanding of board overlap.
-Research has largely focused on governance or firm-level performance, while firm behavior and the broader market implications of board overlap remain underexplored.
+    In this paper, I examine the effects of soft consolidation in the form of overlapping board members in nonprofit hospitals.
+Existing research establishes a strong link between board overlap and governance practices, but we know little about whether there should be anti-competitive concerns when firms share common board members.
 Additionally, most studies have been limited to large, publicly traded firms.
@@
 However, this assumption is flawed, particularly in industries like health care, where nonprofit hospitals dominate and have been shown to engage in anti-competitive practices (\cite{hulver2023ftc}).
-Further research is needed to assess the competitive effects of board overlap across different market settings, particularly in nonprofit sectors where such dynamics may have substantial implications for industry competition and consumer welfare.
-I draw from this literature to inform potential outcomes of interest when considering board overlap in hospitals: financial performance, mergers and acquisitions, investment, and product differentiation.
-I discuss in Section \ref{sec:hospbehaviors} how these outcomes relate to hospitals specifically.
+Despite these insights, gaps persist: research rarely addresses market-level effects or nonprofit sectors, assuming nonprofits pose minimal antitrust risk (\cite{baer2014clayton}; \cite{aai2013section7}).
+This assumption is problematic in health care, where nonprofit hospitals dominate and exhibit anti-competitive behavior (\cite{hulver2023ftc}).
+Further work should explore competitive implications of board overlap beyond governance and financial metrics, particularly in nonprofit contexts.
 
-    \subsection{Common Ownership}\label{sec:commonownerlit}
-
-    As consolidation in investment companies and in many industries has increased over time, so has the prevalence of common ownership among publicly traded firms.
-That is, firms within the same industry and market are increasingly owned by the same investment companies.
-While a separate occurrence and literature than that of board overlap, there are numerous parallels between the two concepts.
-Unlike publicly traded firms, nonprofits do not have shareholders that receive dividend payments.
-However, in some ways, a board functions similarly to shareholders: they influence the strategic direction of the organization through information sharing and voting on firm decisions without actively managing day-to-day operations.
-Further, the literature on common ownership in relation to market concentration and economic incentives is more developed than that of board overlap.
-Thus, I draw on this literature to inform outcomes that are potentially related to board overlap in the hospital industry.
-
-    Whether common ownership affects competition is a widely debated theme in the economics literature.
-Theoretical models predict that common ownership will lead to lower equilibrium output and higher markups through firms placing non-zero weight on competitors' profit (\cite{rubinstein1983competitive}; \cite{rotemberg1984financial}; \cite{azar2012new}).
-Empirical evidence has grown tremendously since the seminal studies by \citeauthor{he2017product} (\citeyear{he2017product}) and \citeauthor{azar2018anticompetitive} (\citeyear{azar2018anticompetitive}), who found that common ownership leads to greater consolidation (acquisitions, joint ventures, strategic alliances), and that common ownership led to higher prices in the airline industry, respectively (\cite{he2017product}; \cite{azar2018anticompetitive}).
-However, these findings have been critiqued for how they measure common ownership and endogeneity concerns (\cite{kennedy2017competitive}; \cite{lewellen2021does}).
-
-    A number of studies have since attempted to estimate the effect of common ownership on a range of outcomes including prices, research and development, merger activity, and patent durability (\cite{gerardi2023critical}).
-Many of which document a statistically significant relationship between common ownership and the relevant outcome.
-However, recent studies using alternative identification strategies find that common ownership can actually increase competition through lower insider trading profits, more product development, and higher investments (\cite{chen2023does}; \cite{kini2024common}).
-Additionally, recent literature has considered common ownership in the context of private firms with venture capitalists, where consolidated ownership is even more prevalent, and has found that competition decreases and information sharing increases in this setting (\cite{lindsey2008blurring}; \cite{gonzalez2020exchanges}; \cite{li2023common}; \cite{eldar2024common}).
-
-    The literature on common ownership yields several conclusions for a study on board overlap.
-The first is a theoretical foundation confirming the importance of particular outcomes relevant to competition and consumer welfare, including merger and acquisition activity and investment decisions.
-Second, the common ownership literature focuses much more on how prices are impacted, which is also relevant in the health care setting.
-Finally, this literature introduces important concerns about the endogenous choice of owners or board members.
-While I do not seek to causally estimate the effects of board overlap in hospitals in this paper, this is an important consideration for future research.
-
-    \subsection{Consolidation in Hospitals}\label{sec:hospconsol}
-
-    There is a rich empirical literature examining the effects of consolidation in health care that includes both vertical and horizontal integration between hospitals, practices, and physicians.
-As the focus of this paper is affiliation among competing hospitals, I draw on the literature examining hospital consolidation in particular through mergers and systems.
-This will inform hospital behaviors that might be correlated with board overlap but are unique to the hospital setting and thus weren't discussed in Section \ref{sec:boardoverlaplit}.
-The main outcomes of interest in this area of research have centered around price and quality of care.
-
-    Early studies on hospital consolidation found very little evidence of impacts on operating costs, efficiency, or quality of care (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}).
-From 2000-2020, the occurrence of hospital mergers increased dramatically, where a quarter of hospital markets included zero independent hospitals by 2020 (\cite{ElevanceHealth2023}).
-This increase in concentration has led to a vast number of studies estimating its effect on quality, operating costs, and prices.
-
-    Studies typically document cost and efficiency gains experienced by hospitals who merge to become part of a larger system (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}), even if the effects are small (\cite{craig2021mergers}).
-Naturally, this leads to the question of whether such efficiency gains are passed through to patients by lower prices.
-Across various settings, data sources, and methodology, hospital mergers are associated with significant increases in price due to hospitals having greater negotiating power with insurance companies (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}).
-
-    There is mixed evidence on whether such gains in efficiency translate to better quality of care.
-Several studies comparing hospitals that became a part of a system with observably similar hospitals that did not merge find little to no effect on readmission or mortality rates on average (\cite{haas2011mergers}; \cite{beaulieu2020changes}).
-However, one study finds that among heart disease patients, treatment intensity and mortality rates both increase after hospital consolidation (\cite{hayford2012impact}), and another finds increased readmission rates across all types of patients (\cite{andreyeva2024corporatization}).
-
-    A select number of studies investigate other dimensions that could be affected by more system affiliation over time.
-In a slightly different setting of acquisitions of dialysis centers, \citeauthor{eliason2020acquisitions} (\citeyear{eliason2020acquisitions}) finds that decreases in quality are driven primarily by the convergence of treatment styles among acquired centers and their acquirer (\cite{eliason2020acquisitions}).
-Additional studies have suggested that consolidated service offerings is a potential mechanism for effects on quality (\cite{mariani2022impact}).
-Finally, \citeauthor{desai2023hospital} (\citeyear{desai2023hospital}) examine admission patterns of low-income patients as a result of increased hospital concentration, and find that Medicaid admissions decrease as a result of increased concentration.
-That is, highly concentrated markets exhibit redistribution patterns of patients within the market (\cite{desai2023hospital}).
-
-    This literature informs ways that hospital affiliations might affect behaviors and outcomes specific to hospitals.
-In addition to the behaviors discussed in Sections \ref{sec:boardoverlaplit} and \ref{sec:commonownerlit}, measures of quality of care, treatment styles, service offerings, and admission patterns could also be important aspects driven by affiliation with other hospitals.
-I discuss these in more detail in Section \ref{sec:hospbehaviors}.
-
-    \subsection{Informing Hospital Behaviors}\label{sec:hospbehaviors}
-
-    The main purpose of this paper is to document the prevalence of board overlap in US hospitals in recent years.
-While I do not establish any causal relationship between board overlap and hospital behaviors, I consider the literature on board overlap, common ownership, and hospital consolidation discussed in Sections \ref{sec:boardoverlaplit}-\ref{sec:hospconsol} to inform a set of hospital behaviors that may be relevant for future research.
-I divide these behaviors into three categories: price/quality, investment decisions, and differentiation.
-
-    The hospital consolidation literature highlights price and quality of care as key factors in the industry.
-Mergers often grant hospitals greater market power, enabling them to negotiate higher prices with insurance companies.
-It is unclear whether shared board members would lead to similar price increases.
-If board overlap facilitates information sharing about negotiations, it could indirectly influence pricing.
-However, direct board involvement in pricing decisions seems less likely.
-Nevertheless, with access to data on actual prices charged by hospitals, this remains an interesting outcome for future research to explore.
-
-    It may be more likely that quality of care could be impacted by overlapping board members among hospitals.
-There are direct actions that could lead to changes in quality such as convergence in treatment styles or investment in technology, which I will discuss in detail in the following sections.
-There is also an indirect path through which overlapping board members might impact quality, which is that hospitals might compete for the same pool of experienced board members.
-If the board members with the most expertise are in high demand, sharing that expertise among multiple hospitals could facilitate greater improvements in quality.
-
-    In both board overlap and common ownership, several investment decisions are shown to be associated with established firm affiliation.
-First, hospitals with board shared board members might be more likely to merge in the future.
-Second, sharing board members could facilitate engagement in the same types of technology or capital.
-Therefore, an area of future research is to look at outcomes such as health information technology investments, specific medical technology advancements in the hospital, and capital investments.
-
-    Finally, previous research has shown that product differentiation is associated with board overlap in other industries.
-That is, sharing information among competing firms can lead to more specialization.
-In hospitals, there are two distinct types of differentiation: either through services offered, or the types of patients admitted.
-Both of these behaviors are key findings in the hospital merger literature, indicating that affiliations through board members could also facilitate changes in these areas.
-In future research, one could think about measuring the likelihood of specialized services being offered, where hospitals might shift investments towards services not offered by their affiliated hospital.
-Additionally, patient composition in the hospital, measured by the payer mix of admitted patients or complexity of their conditions, could be compared among hospitals with and without shared board members.
+    In 2024, the FTC banned a case of interlocking board (\cite{ftc2025sevita}).
 
@@
 
+    \subsection{Nonprofit Hospital Governance}
+
+    Nonprofits are governed at a high level by a board of directors: a group of unpaid community members that gather to oversee the broad direction and strategies of the firm.
+The board of directors does not mange the day-to-day activity of the hospital, but provides oversight to the highest level of manager, the Chief Executive Officer, whom the board also selects.
+The practical governance of boards is highly variable across different hospitals.
+State-level policies might influence details such as the regularity of board meetings or board member term limits.
+However, the main function of the board remains consistent: to provide strategic direction and oversight.
+In research focused on hospital governance, several factors have been found to be correlated with strategic change or hospital-level behaviors, including more influence over the CEO or top management (\cite{golden2001will}; \cite{alexander2008governance}; \cite{jiang2012enhancing}).
+Additionally, since the Affordable Care Act, boards have been actively involved in quality oversight (\cite{jha2010hospital}; \cite{prybil2014board}; \cite{prybil2010board}).
+Therefore, while boards are not usually involved in the day-to-day operations of hospitals, they do influence strategy and long-run hospital behavior.
+
+    The board of directors selects a Chief Executive Officer, who then selects other members of the executive team such as a Chief Financial Officer and/or Chief Medical Officer.
+This leadership team manages the day-to-day operations of the hospital.
+An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize these teams.
+Some hospital executives specialize in health care administration by earning a degree in health care management, or an MBA specific to health care.
+
+    In this paper, I consider the effects of board and executive interlock, someone sitting as an executive or on the board of multiple organizations.
+Board overlap does occur in health care with instances of organizations denied merger permission due to boards not acting independently (\cite{huberfeld2006tackling}), of hospitals declaring informal affiliations through leadership (\cite{barnett_babcock_2012}), and of life sciences companies with substantial documented overlap (\cite{manjunath2024illegal}).
+For many years, oversight of this activity in nonprofits was considered second-order to that of for-profits.
+However, in 2024, the FTC raised concern over board interlock of two prominent health care organizations (\cite{ftc2025sevita}).
+Additionally, board will sometimes ask executives from other organizations to sit on their board (\cite{gordon2025ftc}; \cite{alston2024governance}).
+
+    Early research on board interlock highlights that firms are not isolated entities but interconnected through shared directors (\cite{dooley1969interlocking}), with interlocks linked to firm size, financial ties, and management control.
+Research has examined governance implications, including whether overlapping boards overextend directors (\cite{ferris2003too}; \cite{field2013busy}), increase stock option backdating (\cite{bizjak2009option}), or promote convergence in governance strategies (\cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}).
+Empirical evidence focuses largely on financial outcomes: well-connected directors can enhance firm value (\cite{baran2017director}), and board overlap is associated with higher profitability (\cite{geng2021does}), greater acquisition activity, and improved post-merger performance (\cite{schonlau2009board}; \cite{cai2012board}).
+Mechanisms include shared IT investments (\cite{cheng2021social}), reduced R\&D, and increased product differentiation (\cite{geng2021does}).
+
+    \subsection{Hospital Affiliations}\label{sec:hospaffil}
+
+    Board interlock is often happening in conjunction with affiliation agreements: any relationship between otherwise independent health care providers designed to create shared advantage and value (\cite{aha2018affiliations}).
+Some examples of affiliation agreements include joint ventures for service provision, clinical agreements among hired health care workers, or joint management.
+It is common for these types of agreements to involve some sort of shared governance (\cite{aha2018affiliations}; \cite{natlawreview2023ftc}).
+While some have raised anti-trust concerns over affiliation agreements, they are largely undocumented, and therefore their effects are unknown (\cite{rand2023consolidation}).
+Organizations do not require pre-authorization to form affiliation agreements, but there are several states that now require notice of any type of affiliation (\cite{ncsl2024healthcare}).
+
+    The more documented and studied type of hospital affiliation is that of traditional mergers and acquisitions.
+I will discuss the literature surrounding this topic in order to inform relevant behaviors that could also be affected by affiliation agreements.
+Hospital consolidation has accelerated over the past two decades, with mergers becoming so prevalent that by 2020, a quarter of hospital markets had no independent hospitals (\cite{ElevanceHealth2023}).
+Early studies found little evidence that consolidation improved operating costs, efficiency, or quality (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}), but more recent work documents modest efficiency gains for hospitals joining larger systems (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}; \cite{craig2021mergers}).
+These gains, however, rarely translate into lower prices; instead, mergers consistently lead to significant price increases due to enhanced bargaining power with insurers (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}).
+Evidence on quality is mixed: most studies find little change in readmission or mortality rates (\cite{haas2011mergers}; \cite{beaulieu2020changes}), though some report higher treatment intensity and mortality for heart disease patients (\cite{hayford2012impact}) and increased readmissions overall (\cite{andreyeva2024corporatization}).
+Other research highlights mechanisms such as convergence in treatment styles (\cite{eliason2020acquisitions}), changes in service offerings (\cite{mariani2022impact}), and shifts in admission patterns, including reduced Medicaid admissions in concentrated markets (\cite{desai2023hospital}).
+These findings underscore that consolidation can influence not only costs and prices but also care delivery and patient access.
+
     \section{Data Construction}
 
-   I compile data on nonprofit hospital board of directors in years 2017-2022 from publicly available Tax Form 990s, which all sufficiently large nonprofits must file with the Internal Revenue Service (IRS) each year.
-These forms contain a section in which nonprofits declare their board, executives, and highest compensated employees.
+   To create a measure of board/executive interlock, I compile data on nonprofit hospital board of directors and executives in years 2017-2022 from publicly available Tax Form 990s.
+All sufficiently large nonprofits must file these forms with the Internal Revenue Service (IRS) each year.
+They contain a section in which nonprofits declare their board, executives, and highest compensated employees.
 This paper focuses only on the board of director identities.
@@
 Specifically, I compute the Jaro-Winkler distance between hospital names in each data set.
-Then, I record matches with sufficiently similar names and addresses, yielding a sample of 1,836 matched hospitals.
+Then, I record matches with sufficiently similar names and addresses, yielding a sample of 1,824 matched hospitals.
 
-    I limit the individuals in the sample to those on the board of directors, excluding executives that also serve on the board.
-I then record self-reported information about the board members in the data, including whether they include a title for medical doctor, nurse, or having a degree in health administration.
-I also record whether the firm records that they are a non-voting member of the board (though this is very rare).
-Several steps are necessary to ensure the names are clean and consistent across firms to identify people across hospitals.
-First, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency.
-Second, I combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common.
+     I record self-reported information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration.
+To clean the names, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency.
+I also combine names within the same firm that have either slight misspelling differences or, if middle names are only sometimes listed, they have at least 2 names in common.
 For example, if John Lee Matthew is a board member of hospital A in 2017, and John Matthew is a board member of hospital A in 2018, I assume these names represent the same person.
-I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).
+Finally, I use the gender package in R to predict each individual's gender using historical data (\cite{gender}).
 
-    \import{Objects}{board_people.tex}
+    \import{Objects}{boardand_people.tex}
 

### 2025-12-05 Paper/paper_2025.tex

@@
 
-    \section{Data Construction}
+    \section{Data and Summary Statistics}
 
@@
 
-    \import{Objects}{boardand_people.tex}
+    I present summary statistics of characteristics of the people in the data in Table \ref{tab:board_people}.
+There are just under 37,000 people in the data, 63\% of which are board members and the remaining 37\% are executives.
+Board members are slightly moe likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%.
+Nurses are consistent across board and executives, but only 3\% of the sample are nurses.
+A negligible number of people claim to have a health administration degree.
+Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 32\% of board members estimated to be female.
+I also include average team composition of hospital boards and executive teams, where these trends largely hold at the hospital level.
+On average, hospital boards contain 7 members and executive teams contain 10 members.
+
+    \import{Objects}{boardandexec_people.tex}
 
@@
 Therefore, I define the relevant type of overlap as someone being on the board of multiple hospitals that are otherwise \textit{unaffiliated}.
-Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system as they do.
-Further, In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}).
+Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system.
+I exclude hospitals that gain or lose overlapping board members several times in the sample period.
+
+    In the hospital merger literature, mergers have been found to be more significant when the two hospitals involved are located geographically close (\cite{cooper2019price}).
 Therefore, I define two types of overlap: overlap with an otherwise non-affiliated hospital in the same market, or overlap with an otherwise unaffiliated hospital in a different market.
-All other hospitals have no board overlap with otherwise unaffiliated hospitals.
 I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals.
-
-    The final sample limitation I make for simplicity is to remove hospitals that gain and then lose overlapping board members in the sample.
-Since treatment will be turning on overlapping board member, I remove those who "lose" their treatment.
 
@@
 
-    The full sample of hospitals
-
-    \subsection{Summary Statistics}
-
-    First, I present a summary statistics table showing characteristics of all hospitals in the final sample, which consists of 1,467 nonprofit hospitals, almost all of which are adult general hospitals.
-They have, on average, 178 beds and 14 board members.
-Almost 40\% of the hospitals in the sample are academic medical centers, almost 70\% of the sample are located in metropolitan areas, and 56\% belong to a system.
-
-    %\import{Objects}{all_summarystats.tex}
-
-    Eight percent of hospitals in the sample have overlapping board members with an otherwise unaffiliated hospital.
-In Figure \ref{fig:connected_percent}, I show that this percentage remains relatively constant over time, with a slight decline from 129 affiliated hospitals in 2017 to 114 affiliated hospitals in 2021.
-I include 2022 values, but there are less total firms in the sample due to lag in tax forms being made available publicly.
+    In Figure \ref{fig:connected_percent} I show the prevalence of overlapping board members or executives over time.
+In 2017, approximately 5\% of hospitals have an overlapping board members or executive with another hospital in the same market that is not formally affiliated otherwise.
+That is, these hospital are not under common ownership.
+This percentage increases over time, with 10\% of hospitals sharing a key leader in their market in 2022.
+I also graph the percent of HRR markets where hospitals within that market share board members or executives.
+Approximately 15\% of HRRs are potentially affected by this type of affiliation.
+Sentence here about putting this into perspective relative to hospital consolidation as a whole.
 
@@
         \vspace{9mm}
-        \caption{Percent of hospitals that have a board affiliation}
-        \includegraphics[width=.8\textwidth]{Objects/connected_percent.pdf}
+        \caption{Prevalence of Board/Executive Interlock}
+        \includegraphics[width=.9\textwidth]{Objects/connected_percent.pdf}
         \label{fig:connected_percent}
-    \end{figure}
-
-    As mentioned previously, common board members should be more likely to affect behavior within a hospital market, where hospitals compete for the same patient population.
-In Figure \ref{fig:connected_HRR_percent}, I show the prevalence of having shared board members within a hospital market.
-Of the 306 HRRs in the US, approximately 17\% have at least one pair of board affiliated hospitals within the market.
-This proportion decreases slightly over time, with 15\% of HRRs having connected hospitals in 2021, a nominal decrease of 8 markets losing hospitals with affiliation.
-This is likely due to increased system affiliations, which mechanically decreases board overlap since I do not include hospitals in the same system as eligible for having board overlap.
-
-    \begin{figure}[ht!]
-        \centering
-        \caption{Percent of HRRs containing board affiliated hospitals}
-        \includegraphics[width=.8\textwidth]{Objects/connected_HRR_percent.pdf}
-        \label{fig:connected_HRR_percent}
     \end{figure}
@@
 Over time, the connections remain fairly consistent geographically.
-Taken together, these show that board overlap is a prevalent occurrence among nonprofit hospitals in the US from 2017 onward.
 
@@
         \centering
-        \caption{Geographic distribution of board affiliated hospitals over time}
+        \caption{Geographic Distribution of Board/Executive Interlock}
         \includegraphics[width=\textwidth]{Objects/connected_maps.pdf}
@@
 
-    I now shift to examining characteristics of hospitals that share board members.
-I consider pairs of connected hospitals and characteristics of the pair in Table \ref{tab:hospital_pair_types}.
+    Hospitals in the sample are either always connected, never connected, or become connected at some point in time.
+I show characteristics of hospitals in these categories in Table \ref{tab:hosp_group_stats}.
+
+    \import{Objects}{hosp_group_stats.tex}
+
+    I now present characteristics of pairs of hospitals that are connected by board members or executives in Table \ref{hospital_pair_types}.
+In the first column, I show pair characteristics of hospitals connected within the same HRR.
+In the second column, I present characteristics of hospitals connected to otherwise unaffiliated hospitals located in a different HRR.
+Therefore, all hospital pairs in the table are connected, but only the first column are connected to hospitals that are potentially direct competitors.
 Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data.
-Additionally, the median bed size of all hospitals in the sample is 103, and I define hospitals as being either small or large depending on whether they fall below or above this median.
+Additionally, the median bed size of all hospitals in the sample is 101, and I define hospitals as being either small or large depending on whether they fall below or above this median.
+There are typically 1.5 connections in the same HRR and almost 6 connections in different HRRs.
 
@@
 
-    Two-thirds of board affiliated pairs are between two adult general hospitals, and 10\% of pairs are between a specialty and a general hospital.
-Thus, for the majority of hospitals that have board interlock, the hospital they share a board member with is a direct competitor.
-Additionally, there is significant variation in the ownership type of hospitals within affiliated pairs.
-Forty-five percent of pairs are between two non-system affiliated hospitals, and 36\% of pairs are between an independent hospital affiliated with a hospital belonging to a system.
-The lowest proportion of shared board member pairs is made up of two hospitals belonging to different systems.
+    Most pairs, 76-78\% are two general hospitals.
+Thus, for the majority of hospitals that have board/executive interlock in the same HRR, the hospital they share a leader with is a direct competitor.
+I also include characteristics of system versus independent ownership.
+For there is a nearly even spread of the type of ownership of the pair.
+42\% of these pairs are made up of one independent hospital and one system-owned hospital, while 37\% are made up of two independent hospitals and the rest are made up of two hospitals that are both system owned but belong to two different systems.
+These figures are slightly different for connections in different HRRs, where a larger portion of connections, 45\%, are made up of two system-owned hospitals of two different systems.
 
-    There is also variation in whether the hospitals are the same size or not.
-Half of the hospital pairs consist of one small and one large hospital, and the other half are split evenly between two small or two large hospitals.
-However, the average difference in the number of beds of the two hospitals in a pair is very large: an average difference of 254 beds.
-This suggests that for hospitals of opposite size, the difference is drastic.
-The average distance between hospitals that share common board members is 35 miles, or 57 kilometers.
-In total, there are 217 hospitals that share a board member with at least one other hospital at some point in the sample.
+    In both same and different HRR connected pairs, 50\% of the connections are such that one hospital is small in bed size and one is large in bed size.
+The average bed difference is over 200, indicating that the size difference of the average hospital connection is drastic.
+what's the implication of this?
+Finally, connected hospitals in the same HRR are approximately 62 kilometers, or 38 miles, away from the other.
 
-    \subsection{Correlations with Hospital Behaviors}
-
-    I now merge the hospital board data to other publicly available data sets to examine descriptive differences in the hospital behaviors discussed in Section \ref{sec:hospbehaviors} for hospitals with and without board overlap.
-The broad categories of hospital behaviors that might be affected by affiliation through shared board members are price/quality, investment decisions, and differentiation, which are discussed in Section \ref{sec:hospbehaviors}.
-While I do not capture every aspect of hospital behavior, I select a number of variables that fall into these categories to examine.
-
-    First, from the Center for Medicare and Medicaid Services (CMS) Hospital Compare data, I merge information on each hospital's overall star rating.
-This rating combines measures of quality into one index that ranges from 1 to 5: readmission rates, mortality rates, safety, timeliness of care, effectiveness of care, and patient experience.
-This is a very aggregated measure, but may capture big-picture differences in quality.
-
-    Next, I include data on different types of investments made into the hospital from the Healthcare Cost Reporting Information System (HCRIS) data.
-These variables include the dollar amounts spent on health information technology purchases, fixed equipment, and movable equipment.
-
-    There are a plethora of ways one could measure differentiation in hospitals.
-I focus on two categories of differentiation: service offerings and patient population.
-To measure different service offerings, I first construct a measure of concentration of beds devoted to categories of services in the AHA survey.
-For services $j$ in the list of all possible services\footnote{General medical and surgical (adult), General medical and surgical (pediatric), Obstetric care, Medical/surgical intensive care, Cardiac intensive care, Neonatal intensive care, Neonatal intermediate care, Pediatric intensive care beds, Burn care, Other Special Care, Other intensive care, Physical Rehabilitation care, Alcohol/drug abuse or dependency inpatient care, Psychiatric care, Skilled nursing care, Intermediate nursing care, Acute long term care, Other long-term care, Other care}, the concentration of bed allocation is defined as
-
-    $$concentration_{bed} = \sum_{j}\left(\frac{beds_j}{total beds}\right)^2.$$
-
-    \noindent That is, a higher bed concentration indicates more specialization in bed allocation, where a lower concentration indicates more services spread out in bed allocations.
-Additionally, I include a second concentration measure that relies less on the facility and more on the conditions patients are treated with.
-The CMS Provider Utilization files contain measures of the percent of patients seen for different conditions.
-Therefore, for services $k$ in the list of conditions in the CMS data\footnote{Percent Of Beneficiaries Identified With Combined Cancer Flag For 6 Cancer Indicators, Percent Of Beneficiaries Identified With Chronic Obstructive Pulmonary Disease, Percent Of Beneficiaries Identified With Chronic Kidney Disease, Percent Of Beneficiaries Identified With Heart Failure And Non-Ischemic Heart Disease}, I define the concentration of patient conditions in each hospital as
-
-    $$HHI_{patient} = \sum_{k}\left(\frac{patients_k}{total patients}\right)^2.$$
-
-    Finally, I include indicators for whether the hospital has a NICU or Cath Lab, both profitable services for hospitals to offer.
-I also consider differentiation by the patient population admitted to the hospital.
-There is information in the HCRIS data about how many discharges are for patients on Medicaid or Medicare.
-There is also information in the CMS Provider Utilization Files about the average risk score of Medicare patients seen in the hospital.
-I include all of these variables to capture something about the aggregate patient population being treated in different hospitals.
-
-    In Table \ref{size_characteristics}, I present averages of all of these variables for four categories of hospitals: whether they are small (less than 103 beds) or large (greater than 103 beds), and whether they share board members with another competing hospital.
-I limit this sample to only general hospitals for relevant comparison.
-
-   \import{Objects}{size_characteristics}
-
-   Small hospitals with and without shared board members have a very similar average star rating of 3.4, while large hospitals with shared board members have a slightly smaller average star rating than large hospitals without board overlap.
-In all categories of investment, small hospitals are relatively similar, with slightly more investment in health IT and fixed equipment for board affiliated small hospitals.
-Large hospitals are also generally similar in how they invest in health IT and fixed equipment relative to how much they spend on average, but hospitals with shared board members spend significantly more on movable equipment than hospitals without board overlap.
-
-   There do not seem to be meaningful differences in the concentration of services offered or conditions treated among hospitals with and without board overlap, whether they are small or large.
-Similarly, there are no clear differences in the patient composition of hospitals with and without board overlap.
-This is consistent even among the facilities of these hospitals, with no distinct differences in whether hospitals have a NICU or Cath lab based on board overlap.
-
-   Another important level of variation in this sample of hospitals could be whether they belong to a system or not, as this often dictates a number of the behaviors discussed.
-Therefore, I present a similar table of means for hospitals with and without board overlap that do or do not have a system affiliation in Table \ref{sys_characteristics}.
-
-   \import{Objects}{system_characteristics}
-
-    Among hospitals that do not belong to a system, those without board overlap have a slightly higher quality rating than those without board overlap.
-However, the opposite is true among hospitals that do belong to a system, where those with board overlap have a higher quality rating.
-The most significant difference in investment is among hospitals that belong to a system for investment in movable equipment.
-Those with board overlap invest over 20 million in movable equipment, while those without board overlap only invest 8.3 million.
-The other comparisons of investment show no drastic differences.
-In-system and out-of-system hospitals both show higher levels of concentration of services when they do not have board overlap.
-However, the patient composition is similar within the categories of system affiliation.
-Independent hospitals that do not have board overlap are more likely to have a Cath lab, but otherwise hospitals are similar in in service offerings.
-
-    While these statistics do not establish a causal link between board overlap and hospital behavior, they highlight key differences across hospital categories.
-Notably, hospitals with overlapping board members exhibit distinct patterns in quality and investment decisions.
-These differences underscore the need for further research to explore board overlap as a potential mechanism for affiliation and behavioral changes in health care.
+    add outcome graphs
 

### 2025-12-07 Scripts/clean_matched_names_updated.R
add everything to move to mac

@@ -10,7 +10,7 @@ source("./Scripts/paths.R")
 all_people_data <- readRDS(paste0(created_data_path, "/all_people_data_scheduleH.rds"))
 cw <- readRDS(paste0(created_data_path, "/updated_ein_aha_cw.rds"))
 
-# only keep observations where EIN is found in either ein_hosp or ein_sys
+# only keep observations where EIN is found in the crosswalk
 all_people_data <- all_people_data %>%
   filter(Filer.EIN %in% cw$Filer.EIN)
 

### 2025-12-07 Scripts/create_connections_boardandexec.R

@@ -258,6 +258,9 @@ identify_common_members <- function(data) {
 
 
 
+
+
+
 # Clean up the names ----------------------------------------------------------------------------------------
 
 # Read in cleaned up version of names found in 990 tax forms
@@ -307,19 +310,67 @@ people <- assign_person_id(people)
 # Create variables of board members connected to multiple EINs ---------------------------------------------------
 connections <- identify_common_members(people)
 
-# separate other_eins into multiple columns
-connections <- connections %>%
-  separate_wider_delim(other_eins, delim = ",", names_sep = "", too_few="align_start")
+# only keep EINs that are present in enough years
+ein_keep <- connections %>%
+  distinct(Filer.EIN, TaxYr) %>%
+  mutate(count=ifelse(TaxYr %in% c(2017,2018,2019,2020,2021),1,0)) %>%
+  group_by(Filer.EIN) %>%
+  filter(sum(count)>=5) %>%
+  ungroup() %>%
+  distinct(Filer.EIN)
 
-# wide to long in other_eins
 connections <- connections %>%
-  pivot_longer(cols = starts_with("other_eins"), names_to = "num_board", values_to = "other_ein")
+  filter(Filer.EIN %in% unique(ein_keep$Filer.EIN))
 
-# Remove columns with NA
+# pivot longer
+connections <- connections %>%
+  separate_wider_delim(cols = "other_eins", delim=",", names_sep = "", too_few = "align_start")
+
+connections <- connections %>%
+  pivot_longer(cols=starts_with("other_eins"), names_to="other", values_to="other_ein")
+
+# Remove columns with NA, only keep actual connections
 connections <- connections %>%
   mutate(other_ein = str_trim(other_ein)) %>%
-  filter(!is.na(other_ein)) %>%
-  select(TaxYr, name_cleaned, gender, position, doctor, nurse, other_doctor, ha, nonvoting, Filer.EIN, other_ein)
+  filter(!is.na(other_ein))
+
+# calculate initial overlap of boards for hospital pairs
+filer_board <- connections %>%
+  group_by(TaxYr, Filer.EIN) %>%
+  summarise(board = list(unique(name_cleaned)))
+
+connections <- connections %>%
[…]

### 2025-12-07 Scripts/create_connections_boardonly.R

@@ -349,28 +349,28 @@ connections <- connections %>%
 AHA <- read_csv(paste0(raw_data_path, "\\AHAdata_20052023.csv")) 
 
 AHA_hrr <- AHA %>%
-  select(ID, YEAR, HRRCODE, SYSID, MNAME, SERV) %>%
+  select(ID, YEAR, HRRCODE, SYSID, MNAME, SERV, LAT, LONG) %>%
   filter(YEAR>=2015 & YEAR<=2023) %>%
   mutate(YEAR = as.character(YEAR))
 
 connections <- connections %>%
   left_join(AHA_hrr, by=c("Filer.ID"="ID", "TaxYr"="YEAR")) %>%
-  rename(filer.hrr=HRRCODE, filer.sysid=SYSID, filer.name=MNAME, filer.type=SERV)
+  rename(filer.hrr=HRRCODE, filer.sysid=SYSID, filer.name=MNAME, filer.type=SERV, filer.lat=LAT, filer.long=LONG)
 
 # fill missing variables when possible
 connections <- connections %>%
   group_by(Filer.ID) %>%
-  fill(filer.hrr, filer.type, filer.name, .direction="downup") %>%
+  fill(filer.hrr, filer.type, filer.name, filer.lat, filer.long, .direction="downup") %>%
   ungroup()
 
 connections <- connections %>%
   left_join(AHA_hrr, by=c("other.id"="ID", "TaxYr"="YEAR")) %>%
-  rename(other.hrr=HRRCODE, other.sysid=SYSID, other.name=MNAME, other.type=SERV)
+  rename(other.hrr=HRRCODE, other.sysid=SYSID, other.name=MNAME, other.type=SERV, other.lat=LAT, other.long=LONG)
 
 # fill missing variables when possible
 connections <- connections %>%
   group_by(other.id) %>%
-  fill(other.hrr, other.type, other.name, .direction="downup") %>%
+  fill(other.hrr, other.type, other.name, other.lat, other.long, .direction="downup") %>%
   ungroup()
 
 # Create classifications for each connection type
@@ -415,6 +415,13 @@ people_connections <- people_connections %>%
 saveRDS(people_connections, file=paste0(created_data_path, "people_connections_boardonly.rds"))
 
 # summarize connections at the hospital level -----------------------------------------------
+# create hospital-level connections data set for making a map showing the connections
+hospital_connections <- connections %>%
[…]

### 2025-12-07 Scripts/descriptive_tables_and_figures.R

@@ -15,106 +15,166 @@ library(did)
 
 source("./Scripts/paths.R")
 
-people_data_board <- readRDS(paste0(created_data_path, "people_connections_boardonly.rds"))
 people_data <- readRDS(paste0(created_data_path, "people_connections_boardandexec.rds"))
 hospital_data <- readRDS(paste0(created_data_path, "hospital_data_boardandexec.rds"))
 hospital_connections <- readRDS(paste0(created_data_path, "hospital_connections_boardandexec.rds"))
+AHA <- read_csv(paste0(raw_data_path, "\\AHAdata_20052023.csv")) 
 
-# table: summary stats of the people who are connected versus not connected --------------------------------------------------------
-# I think I should do this for a snapshot in time
-n_conn <- people_data %>%
-  filter(TaxYr==2019 & sameHRR_competing==1) %>%
-  distinct(name_cleaned) %>%
+# filter out hospitals that turn their overlap on and off
+hospital_data <- hospital_data %>%
+  group_by(Filer.EIN) %>%
+  arrange(TaxYr) %>%
+  mutate(lag_treat = dplyr::lag(any_partnership_sameHRR)) %>%
+  mutate(lead_treat = dplyr::lead(any_partnership_sameHRR)) %>%
+  mutate(any_partnership_sameHRR = ifelse(!is.na(lag_treat) & !is.na(lead_treat) & lag_treat==1 & lead_treat==1,1,any_partnership_sameHRR)) %>%
+  mutate(drop = ifelse(lag_treat==1 & any_partnership_sameHRR==0,1,NA)) %>%
+  fill(drop, .direction="downup") %>%
+  ungroup() %>%
+  filter(is.na(drop)) 
+
+# drop 2016 because the sample of hospitals is smaller
+hospital_data <- hospital_data %>%
+  filter(TaxYr!=2016)
+
+# create variables for always, become, and never connected in the same and different HRRs (for partnerships)
+hospital_data <- hospital_data %>%
+  mutate(always_same = ifelse(minyr_sameHRR_part==2017,1,0),
+         become_same = ifelse(minyr_sameHRR_part>2017,1,0),
+         never_same = ifelse(minyr_sameHRR_part==0,1,0)) %>%
+  mutate(always_diff = ifelse(minyr_diffHRR_part==2017,1,0),
+         become_diff = ifelse(minyr_diffHRR_part>2017,1,0),
+         never_diff = ifelse(minyr_diffHRR_part==0,1,0))
+
[…]

### 2025-12-07 Scripts/updated_match_hosp.R

@@ -146,8 +146,11 @@ AHA <- AHA %>%
   mutate(MLOCZIP = str_sub(MLOCZIP, 1, 5)) %>%
   mutate(MLOCZIP = ifelse(nchar(MLOCZIP)==4, paste("0", MLOCZIP, sep=""), MLOCZIP))
 
-# limit the AHA data to 2015 and later and then remove year element
+# limit to hospitals present in the AHA data in the years we care about
 AHA_match <- AHA %>%
+  mutate(count = ifelse(YEAR>2016 & YEAR<2023,1,0)) %>%
+  group_by(ID) %>%
+  filter(sum(count)>0) %>%
   distinct(ID, MNAME, MLOCADDR, FSTCD, MLOCCITY, MLOCZIP)
 
 # do the same for the tax data
@@ -184,8 +187,6 @@ joined_data <- joined_data %>%
   distinct()
 
 # Only keep matches in the same state
-observe <- joined_data %>%
-  filter(Filer.Stfips!=FSTCD)
 joined_data <- joined_data %>%
   filter(Filer.Stfips==FSTCD | is.na(FSTCD)) %>%
   distinct()
@@ -261,10 +262,11 @@ sample <- joined_data %>%
   select(-allyears)
 write.csv(sample, file = paste0(created_data_path, "sample_matches.csv"))
   # first of these had 100% match rate
-  # 7300 observations total
+  # 8600 observations total
 
 joined_data <- joined_data %>%
   distinct(Filer.EIN, ID)
+  # 1558
 
 # Merge the joined data with the original EIN data
 ein_data <- ein_data %>%
@@ -296,7 +298,7 @@ distinct_matches %>%
   filter(!is.na(ID)) %>%
   distinct(Filer.EIN) %>%
   nrow()
-  #1568
+  #1558
 
 
 
@@ -338,6 +340,7 @@ no_match_join <- no_match_join %>%
   distinct(Filer.EIN, ID)
 
 distinct_matches <- bind_rows(distinct_matches, no_match_join)
+# 2793 obs
 
 # remove observations with NA if they have a non-NA value
 distinct_matches <- distinct_matches %>%
@@ -348,7 +351,7 @@ distinct_matches <- distinct_matches %>%
 # look at those with no matches
 no_matches2 <- distinct_matches %>%
   filter(is.na(ID)) 
-  # 756 observations
+  # 767 observations
 
 
 
[…]

### 2025-12-09 Scripts/descriptive_tables_and_figures.R
update script

@@ -15,6 +15,8 @@ library(did)
 
 source("./Scripts/paths.R")
 
+options(knitr.kable.NA = '')
+
 people_data <- readRDS(paste0(created_data_path, "people_connections_boardandexec.rds"))
 hospital_data <- readRDS(paste0(created_data_path, "hospital_data_boardandexec.rds"))
 hospital_connections <- readRDS(paste0(created_data_path, "hospital_connections_boardandexec.rds"))
@@ -142,7 +144,6 @@ knitr::kable(stats,
              align=c("l","c","c","c","c", "c", "c"),
              position="ht!",
              linesep = "") %>%
-  add_header_above(c(" "=2, "Percent" =5)) %>%
   pack_rows(group_label = "Individuals", start_row = 1, end_row = 2) %>%
   pack_rows(group_label = "Hospital Teams", start_row = 3, end_row = 4) %>%
   write(file = paste0(objects_path, "boardandexec_people.tex"))
@@ -560,33 +561,184 @@ knitr::kable(stat,
   pack_rows("", 5, 5, indent = FALSE, latex_gap_space = "-0.3em") %>%
   write(file = paste0(objects_path, "hosp_group_stats.tex"))
 
-# graph: outcome variables by group------------------------------------------------------
+# table: outcome variables by group broken down by pre- and post------------------------------------------------------
 # define the groups
 hospital_data <- hospital_data %>%
-  mutate(group = ifelse(minyr_sameHRR_part>2017, "Become Connected Same", NA)) %>%
-  mutate(group = ifelse(is.na(group) & minyr_diffHRR_part>2017, "Become Connected Different", group)) %>%
-  mutate(group = ifelse(is.na(group) & always_diff==1, "Always Connected Different", group)) %>%
-  mutate(group = ifelse(is.na(group) & never_same==1 & never_diff==1, "Never Connected", group))
+  mutate(group = ifelse(become_same==1 & (minyr_diffHRR_part>minyr_sameHRR_part | minyr_diffHRR_part==0), "Become Connected Same", NA)) %>%
+  mutate(group = ifelse(never_same==1 & never_diff==1, "Never Connected", group)) %>%
+  mutate(group = ifelse(become_diff==1 & is.na(group), "Become Connected Different", group))
 
 # create relative year variable
[…]

### 2025-12-09 Scripts/descriptive_tables_and_figures.R
start adding outcome graphs

@@ -678,20 +678,18 @@ outcome_graphs_relative <- outcome_graphs %>%
                  "movableequipment_purch", "tot_pat_rev", "tot_operating_exp"), list(m=mean), na.rm=T) %>%
   pivot_longer(cols = ends_with("_m"), names_to = "variable", values_to = "mean") %>%
   filter(rel_year<4 & rel_year>-4) 
-outcome_table_other <- outcome_table %>%
+outcome_graphs_other <- outcome_graphs %>%
   filter(is.na(rel_year)) %>%
   group_by(TaxYr, group) %>%
   summarise_at(c("perc_mcaid", "perc_mcare", "tot_discharges", "bed_conc", "build_purch", "fixedequipment_purch",
                  "movableequipment_purch", "tot_pat_rev", "tot_operating_exp"), list(m=mean), na.rm=T) %>%
-  pivot_longer(cols = ends_with("_m"), names_to = "variable", values_to = "mean") %>%
-  group_by(group, variable) %>%
-  summarise(Pre=mean(mean))
+  pivot_longer(cols = ends_with("_m"), names_to = "variable", values_to = "mean") 
 
-
+graphs <- bind_rows(outcome_graphs_relative, outcome_graphs_other)
 
 
 # graph for operating expenses
-tot_operating_exp <- graphs %>%
+graphs %>%
   filter(variable=="tot_operating_exp_m") %>%
   mutate(mean = mean/100000) %>%
   mutate(
@@ -722,10 +720,112 @@ tot_operating_exp <- graphs %>%
     data = data.frame(axis_type = "Relative Year", x = 0, xend = 0, y = 0, yend = 3000),
     aes(x = x, xend = xend, y = y, yend = yend),
     linetype = "dashed", color = "gray", inherit.aes = FALSE
-  )
+  ) %>%
+  ggsave(filename = paste0(objects_path, "tot_oper_exp_graph.pdf"), width = 6, height = 4, units = "in")
+
+# graph for percent medicaid and percent medicare
+
+library(dplyr)
+library(ggplot2)
+
+# Build plot
+(graphs %>%
+  filter(variable %in% c("perc_mcaid_m", "perc_mcare_m")) %>%
+  mutate(variable = ifelse(variable == "perc_mcaid_m", "Percent Medicaid", "Percent Medicare")) %>%
+  mutate(
+    axis_type = case_when(
[…]

### 2025-12-10 Paper/paper_2025.tex
Updates from Overleaf

@@
 
-    add outcome graphs
+    \import{Objects}{hosp_outcomes_table.tex}
+
+    \section{Estimating Anti-competitive Effects of Board Overlap}
+
+    One of the biggest challenges to understanding
 
@@
 
+    \appendix
+
+% First part of the figure (page 1)
+\begin{figure}[p] % 'p' places figure on a dedicated float page
+    \centering
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{bed_conc}
+        \caption{Caption for image 1}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image2}
+        \caption{Caption for image 2}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image3}
+        \caption{Caption for image 3}
+    \end{subfigure}
+
+    \caption{Overall caption for stacked subfigures (part 1)}
+\end{figure}
+
+% Second part of the figure (page 2)
+\begin{figure}[p]
+    \ContinuedFloat % continues numbering from previous figure
+    \centering
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image4}
+        \caption{Caption for image 4}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image5}
+        \caption{Caption for image 5}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image6}
+        \caption{Caption for image 6}
+    \end{subfigure}
+
+    \caption{Overall caption for stacked subfigures (part 2)}
+\end{figure}
+
+% Third part of the figure (page 3)
+\begin{figure}[p]
+    \ContinuedFloat
+    \centering
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image7}
+        \caption{Caption for image 7}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image8}
+        \caption{Caption for image 8}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{image9}
+        \caption{Caption for image 9}
+    \end{subfigure}
+
+    \caption{Overall caption for stacked subfigures (part 3)}
+\end{figure}
+
 \end{document}

### 2025-12-10 Paper/paper_2025.tex

--- 2025-12-05 Paper/paper_2025.tex
@@
 
+    \appendix
+
+% First part of the figure (page 1)
+\begin{figure}[p] % 'p' places figure on a dedicated float page
+    \centering
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{Objects/bed_conc_graph.pdf}
+        \caption{Bed Concentration}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{Objects/tot_discharges_graph.pdf}
+        \caption{Total Discharges}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{Objects/perc_patients_graphs.pdf}
+        \caption{Patient Type Proportions}
+    \end{subfigure}
+
+    \caption{ }
+\end{figure}
+
+% Second part of the figure (page 2)
+\begin{figure}[p]
+    \ContinuedFloat % continues numbering from previous figure
+    \centering
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{Objects/tot_oper_exp_graph.pdf}
+        \caption{Operating Expenses}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{Objects/purchases_graphs.pdf}
+        \caption{Purchases}
+    \end{subfigure}
+
+    \begin{subfigure}{\textwidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{Objects/pat_rev_graph.pdf}
+        \caption{Patient Revenue}
+    \end{subfigure}
+
+    \caption{ }
+\end{figure}
+
 \end{document}

### 2025-12-18 Paper/paper_2025.tex
Updates from Overleaf

--- 2025-12-05 Paper/paper_2025.tex
@@
 
-	\linespread{1.2}\title{\vspace{-0.5in} Unobserved Firm Affiliations: Evidence from Hospital Board Overlap}
+	\linespread{1.2}\title{\vspace{-0.5in} Non-Merger Affiliations in Already Consolidated Health Care Markets}
 
@@
 
-    Antitrust enforcement in the US has been particularly relevant in recent years as price gouging and consolidation has increased.
+    Market concentration has increased across industries in the last two decades (\cite{grullon2019us}), sparking renewed debate and, more recently, a shift toward stricter antitrust enforcement.
+Modern antitrust policy primarily focuses on mergers and acquisitions among large firms.
+Since the Hart–Scott–Rodino Act of 1976, the U.S.
+government has reviewed certain transactions that may threaten competition, defined by regulators as “a process of rivalry that incentivizes businesses to offer lower prices, improve wages and working conditions, enhance quality and resiliency, innovate, and expand choice” (\cite{DOJFTC2023MergerGuidelines}).
+Formal challenges remain rare: since 2015, roughly 3\% of proposed mergers have faced extended review, and 32 proposals were dissolved or restructured (\cite{ConversableEconomist2025Antitrust}).
 
-    Antitrust legislation creates an additional hurdle for large-scale anti-competitive behavior, but there are other types of firm affiliations that typically elude antitrust scrutiny such as common ownership, joint ventures, and other types of affiliation/partnerships.
-By and large, firms form partnerships with other firms, often competitors, without needing to seek approval.
-This has started to raise concern recently, and states have even begun regulating smaller-scale affiliations in recent years.
-Horizontal mergers and consolidation have rightfully received a lot of attention in the literature, often linked to something??
-However, we know little about the effects of smaller-scale affiliations, which could carry some of the same implications as consolidation proposals scrutinized for their ant-competitive nature.
+    Careful scrutiny of large-scale mergers is important.
+However, growing empirical and policy concern centers on unreported, small-scale mergers and non-merger affiliations.
+Research shows that small-scale mergers, often called “stealth mergers”, can reduce competition (\cite{majerovitz2021consolidation}; \cite{aggarwal2025stealth}; \cite{wollmann2019stealth}; \cite{wollmann2020get}; \cite{kepler2023stealth}).
+Non-merger affiliations, by contrast, encompass structural ties such as joint ventures, management consolidation, and overlapping boards of directors that do not trigger merger review.
+Among these, board interlocks are particularly noteworthy because they create direct governance links between firms, are explicitly addressed under the Clayton Act, and have drawn increasing policy attention.
+While board interlocks are documented (\cite{LemleyVanLoo2025OverlappingDirectors}; \cite{Poberejsky2025InterlocksInnovation}), their competitive effects remain largely hypothesized.
 
-    In this paper, I examine the effects of soft consolidation in the form of overlapping board members in nonprofit hospitals.
-Existing research establishes a strong link between board overlap and governance practices, but we know little about whether there should be anti-competitive concerns when firms share common board members.
-Additionally, most studies have been limited to large, publicly traded firms.
-This focus stems from the assumption that nonprofits engage in less anti-competitive behavior (\cite{baer2014clayton}; \cite{aai2013section7}), making board overlap in such organizations less concerning.
-However, this assumption is flawed, particularly in industries like health care, where nonprofit hospitals dominate and have been shown to engage in anti-competitive practices (\cite{hulver2023ftc}).
-Despite these insights, gaps persist: research rarely addresses market-level effects or nonprofit sectors, assuming nonprofits pose minimal antitrust risk (\cite{baer2014clayton}; \cite{aai2013section7}).
-This assumption is problematic in health care, where nonprofit hospitals dominate and exhibit anti-competitive behavior (\cite{hulver2023ftc}).
-Further work should explore competitive implications of board overlap beyond governance and financial metrics, particularly in nonprofit contexts.
+    This issue is especially relevant in health care, where consolidation is a persistent concern and several states have begun regulating small-scale mergers and non-merger affiliations (\cite{ncsl2024healthcare}).
+Additionally, in September 2025, the FTC announced the resignation of three directors from Sevita Health after uncovering board overlap with Beacon Specialized Living Services, both providers of residential care for individuals with intellectual and developmental disabilities, citing concerns under the Clayton Act and urging firms to proactively avoid such conflicts (\cite{ftc2025sevita}).
+Board interlocks are no longer abstract governance issues but active antitrust targets.
+Thus, in this paper, I construct a measure of non-merger affiliation among U.S.
+nonprofit hospitals that captures both interlocking boards and executive cross-board service, and I examine how these affiliations influence hospital behavior.
 
-    In 2024, the FTC banned a case of interlocking board (\cite{ftc2025sevita}).
+    I collect data on nonprofit hospital executives and board members from publicly available IRS Form 990 filings and merge these records with the American Hospital Association (AHA) Annual Survey.
+Using this combined dataset, I classify hospital affiliations based on system membership and geographic market.
+I define a non-merger affiliation as the presence of a common board member or an executive serving on another hospital’s board, where the two hospitals are otherwise unaffiliated.
+That is, they do not belong to the same system or network but share leadership ties.
+Because the hypothesized effects of non-merger affiliations are anti-competitive, whether the affiliated hospitals compete for the same patients is a critical distinction.
+To capture this, I separate affiliations into those occurring within the same Hospital Referral Region (HRR) and those across different HRRs, where HRRs represent geographic hospital markets (\cite{Wennberg1973SmallAV}).
+I show that [X]\% of nonprofit hospitals in my sample share a leader with an otherwise unaffiliated hospital in the same market, and [Y]\% of HRRs contain at least one such affiliation.
+I consider several hospital outcomes as behaviors that are informed by the corporate governance literature and hospital consolidation literature.
+I group these outcomes into categories of investment, patient composition, service offerings, and operating costs.
 
-    \section{Background/Setting}
+    Similarly to the merger literature, selection into the decision to become affiliated complicates causal inference.
+I begin with a staggered difference-in-differences design comparing hospitals that form an affiliation to those that do not, and I show that [INSERT FINDINGS].
+While event-study estimates reveal no evidence of differential pre-trends, unobserved factors may still influence affiliation decisions.
+Therefore, my preferred specification employs a triple-differences approach that leverages variation across time, affiliation status, and whether the affiliation occurs within the same geographic market.
+This design isolates the incremental effect of affiliating with a hospital in the same market where competitive effects are most salient.
+Using this approach, I find that [INSERT FINDINGS].
+
+    [IMPLICATIONS OF FINDINGS] Firms are responsive to antitrust regulation.
+That is, enforcement is a deterrant for anti-competitive behavior.
+Thus, it is important to think about whether regulation should increase even for small-scale affiliations.
+(something like this for policy implications?)
+
+    Prior work establishes that corporations are deeply interconnected through overlapping directors and that these ties shape governance practices and firm performance (\cite{dooley1969interlocking}; \cite{ferris2003too}; \cite{field2013busy}; \cite{bizjak2009option}; \cite{bouwman2011corporate}; \cite{chiu2013board}; \cite{cai2014board}; \cite{baran2017director}; \cite{geng2021does}; \cite{schonlau2009board}; \cite{cai2012board}; \cite{cheng2021social}).
+That literature largely centers on publicly traded firms and frames overlap as either a governance friction (busy directors, option backdating, strategy convergence) or a source of private benefits (director networks that raise value, profitability, and post‑M\&A performance).
+In contrast, I study non‑merger affiliations in nonprofit hospitals, an arena often presumed less prone to anti-competitive behavior (\cite{baer2014clayton}; \cite{aai2013section7}) despite mounting evidence of competitive risks in health care (\cite{hulver2023ftc}).
+I broaden the construct beyond traditional board interlock to include executive cross‑board service, and I move the focus from governance proxies and firm‑level valuation to market‑relevant behaviors that are informed by the literature on both governance and consolidation (\cite{baran2017director}; \cite{geng2021does}; \cite{schonlau2009board}; \cite{cai2012board}; \cite{cheng2021social}).
+Methodologically, I leverage a triple‑differences design that distinguishes affiliations formed within versus across geographic markets to isolate the incremental, competition‑salient effect of overlap, rather than inferring behavior from financial outcomes alone.
+
+    A large empirical literature examines the effects of consolidation in health care, particularly through hospital mergers and system integration.
+Early work found little evidence that consolidation improved operating costs, efficiency, or quality of care (\cite{alexander1996short}; \cite{ho2000hospital}; \cite{dranove2003hospital}).
+Yet from 2000 to 2020, hospital mergers surged, leaving a quarter of markets with no independent hospitals by 2020 (\cite{ElevanceHealth2023}).
+This trend spurred extensive research on price, quality, and efficiency.
+Studies consistently document modest cost and efficiency gains for hospitals joining larger systems (\cite{schmitt2017hospital}; \cite{andreyeva2024corporatization}; \cite{craig2021mergers}), but these gains rarely translate into lower prices; instead, mergers typically lead to significant price increases as hospitals gain bargaining power with insurers (\cite{gaynor2012impact}; \cite{boozary2019association}; \cite{cooper2019price}; \cite{andreyeva2024corporatization}).
+Evidence on quality is mixed: most studies find little effect on readmissions or mortality (\cite{haas2011mergers}; \cite{beaulieu2020changes}), though some report increased treatment intensity and mortality for heart disease patients (\cite{hayford2012impact}) or higher readmission rates overall (\cite{andreyeva2024corporatization}).
+Beyond price and quality, research highlights mechanisms such as convergence in treatment styles (\cite{eliason2020acquisitions}), consolidated service offerings (\cite{mariani2022impact}), and redistribution of patients within markets, including declines in Medicaid admissions in highly concentrated areas (\cite{desai2023hospital}).
+While this literature focuses on mergers and system affiliation, my study examines a different form of structural connection: non-merger affiliations through board and executive overlap.
+
+    The paper proceeds as follows.
+In Section \ref{sec:background}, I provide institutional details on nonprofit hospital governance and hospital affiliations.
+In Section \ref{sec:data}, I outline the data collection process and provide summary statistics.
+In Section \ref{sec:simpledid}, I estimate a simple staggered difference-in-differences model leveraging affiliations over time.
+Finally, in Section \ref{sec:tripledid}, I leverage variation in both affiliation and geography to estimate a triple differences specification.
+
+    \section{Background/Setting}\label{sec:background}
 
@@
 
-    \section{Data and Summary Statistics}
+    \section{Data and Summary Statistics}\label{sec:data}
 
@@
 
-    \subsection{Documenting Overlapping Board Members}
+    \subsection{Documenting Leadership Overlap}
 
@@
 
-    add outcome graphs
+    \import{Objects}{hosp_outcomes_table.tex}
+
+    \section{Is Leadership Overlap Correlated with Consolidation?}
+
+    While hospitals often publicly declare that their leadership affiliations are not intended to lead to mergers or other  consolidation (add citations), the relationship between leadership overlap and consolidation is an open empirical question.
+Theoretically, there are three ways that leadership overlap and consolidation could be related.
+First, leadership overlap may act as a substitute for formal mergers.
+In this case, hospitals leverage shared governance to achieve some benefits of consolidation, such as strategic alignment, resource coordination, and cost efficiencies, without incurring the regulatory scrutiny and transaction costs associated with a merger.
+If this hypothesis holds, we would expect leadership overlap to reduce the likelihood of subsequent merger activity.
+Second, leadership overlap may serve as a precursor to merging.
+Hospitals might begin by sharing board members or executives to build trust, harmonize decision-making, and prepare for operational integration.
+Under this interpretation, overlap is an intermediate step toward full consolidation, and its presence should increase the probability of a future merger.
+Third, leadership overlap may result from intentions to merge.
+Here, overlap is not an independent strategic choice but rather a signal of an already-decided merger plan.
+This scenario introduces endogeneity concerns, as leadership overlap and merger intentions are jointly determined.
+In addition to its possible role in mergers, leadership overlap may also influence the likelihood of hospital closure.
+Shared leadership could provide access to broader networks, financial resources, and managerial expertise, which might help struggling hospitals remain viable and avoid closure.
+Conversely, overlap could signal strategic realignment within a system, where weaker hospitals are deliberately phased out to concentrate resources in stronger facilities.
+In this sense, leadership overlap could either reduce closure risk by stabilizing operations or increase it if consolidation strategies involve selective shutdowns.
+
+    To evaluate these possibilities, I estimate the effect of gaining or losing leadership overlap on two outcomes: (1) the likelihood that a hospital closes in the future, and (2) the likelihood that it merges in the future.
+I employ a stacked difference-in-differences (DiD) approach, which constructs event-time observations for each treated hospital around the time of gaining or losing overlap and stacks these into a single panel.
+This method accommodates variation in treatment timing across hospitals and enables estimation of dynamic effects.
+
+    \begin{equation}
+    Y_{it} = \alpha + \sum_{k \neq -1} \beta_k \cdot D_{i,t+k} + \gamma_i + \delta_t + \varepsilon_{it}
+    \end{equation}
+
+    \noindent where $Y_{it}$ is the outcome for hospital $i$ at time $t$ (closure or merger indicator), $D_{i,t+k}$ is an indicator for event time $k$ relative to the change in leadership overlap (with $k = -1$ as the omitted category), $\gamma_i$ and $\delta_t$ are hospital and time fixed effects, respectively, and $\beta_k$ captures the effect at each event time $k$.
+
+    \begin{figure}[ht!]
+        \centering
+        \vspace{9mm}
+        \caption{Effect of Overlap on Closure}
+        \includegraphics[width=.9\textwidth]{Objects/closure.pdf}
+        \label{fig:closure}
+    \end{figure}
+
+    \begin{figure}[ht!]
+        \centering
+        \vspace{9mm}
+        \caption{Effect of Overlap on in-System Overlap}
+        \includegraphics[width=.9\textwidth]{Objects/any_formal_sameHRR.pdf}
+        \label{fig:any_formal_sameHRR}
+    \end{figure}
+
+    \begin{figure}[ht!]
+        \centering
+        \vspace{9mm}
+        \caption{Effect of Overlap on Being Independent}
+        \includegraphics[width=.9\textwidth]{Objects/independent.pdf}
+        \label{fig:independent}
+    \end{figure}
 
@@
 
+    \appendix
+
+\begin{figure}
+    \centering
+    \caption{Outcome Variables Over Time, Patient and Service Allocation}
+    \includegraphics[width=0.75\linewidth]{Objects/outcome_perc_dis_bed_graph.pdf}
+    \label{fig:outcome_means1}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \caption{Outcome Variables Over Time, Purhcases, Costs, and Revenue}
+    \includegraphics[width=0.75\linewidth]{Objects/outcome_purch_oper_rev_graph.pdf}
+    \label{fig:outcome_means2}
+\end{figure}
+
 \end{document}

### 2025-12-18 Paper/paper_2025.tex

@@
 
-    \section{Is Leadership Overlap Correlated with Consolidation?}
+    \section{Estimation and Identification}
 
-    While hospitals often publicly declare that their leadership affiliations are not intended to lead to mergers or other  consolidation (add citations), the relationship between leadership overlap and consolidation is an open empirical question.
+    I estimate the effect of gaining or losing leadership overlap on several hospital outcomes.
+Since hospitals become connected via leadership at different times and their effects could be heterogeneous, I employ a stacked difference-in-differences (DiD) approach.
+Specifically, I construct event-time observations for each treated hospital around the time of gaining or losing overlap and stacks these into a single panel.
+This method accommodates variation in treatment timing across hospitals and enables estimation of dynamic effects.
+On the stacked panel, I estimate the following equation:
+
+    \begin{equation}\label{eq:stacked}
+    Y_{it} = \alpha + \sum_{k \neq -1} \beta_k \cdot D_{i,t+k} + \gamma_i + \delta_t + \varepsilon_{it}
+    \end{equation}
+
+    \noindent where $Y_{it}$ is the outcome for hospital $i$ at time $t$ (closure or merger indicator), $D_{i,t+k}$ is an indicator for event time $k$ relative to the change in leadership overlap (with $k = -1$ as the omitted category), $\gamma_i$ and $\delta_t$ are hospital and time fixed effects, respectively, and $\beta_k$ captures the effect at each event time $k$.
+
+    \section{Leadership Overlap and Consolidation}
+
+    While hospitals often publicly declare that their leadership affiliations are not intended to lead to mergers or other consolidation (add citations), the underlying nature of overlapping leadership with regards to other types of hospital affiliations is an open empirical question.
 Theoretically, there are three ways that leadership overlap and consolidation could be related.
@@
 
-    To evaluate these possibilities, I estimate the effect of gaining or losing leadership overlap on two outcomes: (1) the likelihood that a hospital closes in the future, and (2) the likelihood that it merges in the future.
-I employ a stacked difference-in-differences (DiD) approach, which constructs event-time observations for each treated hospital around the time of gaining or losing overlap and stacks these into a single panel.
-This method accommodates variation in treatment timing across hospitals and enables estimation of dynamic effects.
-
-    \begin{equation}
-    Y_{it} = \alpha + \sum_{k \neq -1} \beta_k \cdot D_{i,t+k} + \gamma_i + \delta_t + \varepsilon_{it}
-    \end{equation}
-
-    \noindent where $Y_{it}$ is the outcome for hospital $i$ at time $t$ (closure or merger indicator), $D_{i,t+k}$ is an indicator for event time $k$ relative to the change in leadership overlap (with $k = -1$ as the omitted category), $\gamma_i$ and $\delta_t$ are hospital and time fixed effects, respectively, and $\beta_k$ captures the effect at each event time $k$.
+    The estimates from Equation \ref{eq:stacked} where hospital closure is the outcome variable are shown in Figure \ref{fig:closure}.
+I consider three types of variation in leadership overlap: gaining overlap with a hospital in the same HRR or different HRR, and losing overlap with a hospital in the same HRR or different HRR.
+While gaining leadership overlap with another hospital in the same or different HRR is not correlated with closing in the future, hospitals that lose overlap are .15 percentage points more likely to close 1-4 years after losing connectedness compared to (who are the comparison hospitals here??).
+Part of this may is mechanical, as hospitals that close (drop out of the data) necessarily lose board overlap.
+(I don't think theres any way to get rid of this mechanical relationship...)
 
@@
 
-    \begin{figure}[ht!]
-        \centering
-        \vspace{9mm}
-        \caption{Effect of Overlap on in-System Overlap}
-        \includegraphics[width=.9\textwidth]{Objects/any_formal_sameHRR.pdf}
-        \label{fig:any_formal_sameHRR}
-    \end{figure}
+    I also estimate Equation \ref{eq:stacked} where the outcome variable is an indicator for whether or not the hospital is independent, where independent means the hospital does not belong to a system.
+I present the results of this estimation in Figure \ref{fig:independent}.
+In this case, losing leadership overlap is not associated with any changes in system status.
+However, gaining leadership overlap in the same HRR is slightly negatively associated with being independent in the first year after becoming connected via leadership.
 
@@
         \label{fig:independent}
+    \end{figure}
+
+    This does not necessarily mean that leadership overlap causes closure or mergers, as the relationship could be more complex than that.
+Add more to this paragraph about how this is not causal but indicates that the nature of board overlap is more than just expertise sharing??
+
+    \section{Leadership Overlap and Hospital Outcomes}
+
+    If leadership overlap, encompassing informal hospital affiliations, is a substitute for more formal mergers, there are several outcomes that we typically see affected in the hospital merger literature that we can test whether they are also associated with leadership overlap.
+These outcomes fall into categories of patient population, service offerings, and operating expenses.
+
+    \subsection{Patient Population}
+
+    \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Medicaid Discharges}
+        \includegraphics[width=\textwidth]{Objects/mcaid_discharges.pdf}
+        \label{fig:mcaid}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Medicare Discharges}
+        \includegraphics[width=\textwidth]{Objects/mcare_discharges.pdf}
+        \label{fig:mcare}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Total Discharges}
+        \includegraphics[width=\textwidth]{Objects/tot_discharges.pdf}
+    \end{subfigure}
+    \label{fig:mcaid_mcare}
+    \end{figure}
+
+    \subsection{Service Offerings}
+
+    \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Bed Concentration}
+        \includegraphics[width=\textwidth]{Objects/bed_conc.pdf}
+        \label{fig:bed_conc}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Number of Services}
+        \includegraphics[width=\textwidth]{Objects/num_services.pdf}
+        \label{fig:num_services}
+    \end{subfigure}
+    \label{fig:services}
+    \end{figure}
+
+    \subsection{Investment and Operating Expenses}
+
+    \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Purchases (Building or Equipment)}
+        \includegraphics[width=\textwidth]{Objects/any_purch.pdf}
+        \label{fig:any_purch}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.47\textwidth}
+        \caption{Total Operating Expenses}
+        \includegraphics[width=\textwidth]{Objects/tot_operating_exp.pdf}
+        \label{fig:tot_oper_exp}
+    \end{subfigure}
+    \label{fig:expenses}
     \end{figure}

### 2025-12-19 Scripts/DownloadXMLs.R
(hopefully) one last graph update for the draft

@@ -1,0 +1,204 @@
+
+# NONPROFIT EXPLORER: SEARCH → SCRAPE ORG PAGE → DOWNLOAD ALL XMLs
+# Parallel across EINs with future.apply (polite: 2–4 workers)
+
+suppressPackageStartupMessages({
+  library(httr)         # HTTP requests
+  library(jsonlite)     # JSON parsing
+  library(stringr)      # String helpers
+  library(rvest)        # HTML scraping
+  library(xml2)         # url_absolute for absolute URLs
+  library(fs)           # Files/dirs
+  library(glue)         # String interpolation
+  library(curl)         # Streaming downloads
+  library(future.apply) # Parallel across EINs
+})
+
+# ------------ Configuration ------------
+BASE_SEARCH <- "https://projects.propublica.org/nonprofits/api/v2/search.json"
+ORG_PAGE    <- "https://projects.propublica.org/nonprofits/organizations"
+UA          <- "Hanna-Glenn-Research/1.0 (+R httr; Nonprofit XML downloader)"
+
+# Output root directory
+out_root <- "data/nonprofit_xml"
+fs::dir_create(out_root)
+
+# Concurrency (adjust politely)
+workers <- 6  # start with 2–4; raise only if stable
+
+# ------------ Helpers ------------
+
+# Normalize EIN to digits-only (expect 9 digits)
+normalize_ein <- function(x) {
+  d <- stringr::str_replace_all(x, "[^0-9]", "")
+  if (nchar(d) != 9) {
+    warning(glue::glue("EIN '{x}' normalizes to '{d}' (not 9 digits)."))
+  }
+  d
+}
+
+# Format EIN with hyphen for search q=XX-XXXXXXX
+format_ein_hyphen <- function(d) {
+  if (nchar(d) == 9) paste0(substr(d, 1, 2), "-", substr(d, 3, 9)) else d
+}
+
+# GET with UA and simple retry/backoff
+api_get <- function(url, query = list(), retries = 3, sleep_sec = 1) {
+  for (i in seq_len(retries)) {
+    resp <- httr::GET(url, query = query, httr::user_agent(UA), httr::timeout(30))
+    if (!httr::http_error(resp)) return(resp)
+    message(glue::glue("HTTP {httr::status_code(resp)} for {url}; retry {i}/{retries}"))
+    Sys.sleep(sleep_sec * i)
+  }
[…]

### 2025-12-19 Scripts/Read990_scheduleH.R

@@ -15,11 +15,11 @@ replace_null <- function(x) {
 
 
 # make a list of all folders to search through
-folders <- list.files(path = paste0(drive_path, "IRS_XML_990s/"), full.names = FALSE)
+#folders <- list.files(path = paste0(drive_path, "IRS_XML_990s/"), full.names = FALSE)
 
 # create empty data frames to store data
-all_ein_data <- data.frame()
-all_people_data <- data.frame()
+#all_ein_data <- data.frame()
+#all_people_data <- data.frame()
 
 # read in temp data sets
 all_ein_data <- readRDS(paste0(created_data_path, "all_ein_data_scheduleH.rds"))
@@ -119,6 +119,103 @@ for (folder in folders) {
     }
   }
 }
+
+
+additional_folders <- list.files(path = "./data/", full.names = FALSE)
+
+# loop through each folder from the webscraping exercise to get older XMLs as well (see DownloadXMLs.R)
+for (folder in additional_folders) {
+  # make a list of all files in the folder
+  files <- list.files(path = paste0("./data/",folder), full.names = FALSE)
+  
+  # loop through files
+  for (file in files){
+    # read in the xml file
+    # See if the file is non-empty
+    if (file.size(paste0("./data/",folder,"/",file)) > 0){
+      xml <- xmlParse(paste0("./data/",folder,"/",file))
+      xml_list <- xmlToList(xml, addAttributes = TRUE)
+      
+        ein <- xml_list[["ReturnHeader"]][["Filer"]][["EIN"]]
+        # convert to list format
+        xml_list <- xmlToList(xml, addAttributes = TRUE)
+        # fill all missing values with NA
+        xml_list <- replace_null(xml_list)
+        
+        header <- as.data.frame(xml_list[["ReturnHeader"]])
+        
+        data <- list_flatten(xml_list[["ReturnData"]], name_repair = "minimal")
+        data <- list_flatten(data, name_repair = "minimal")
+        data <- list_flatten(data, name_repair = "minimal")
+        data <- list_flatten(data, name_repair = "minimal")
+        data <- list_flatten(data, name_repair = "minimal")
+        data <- keep(data, ~length(.x) == 1)
+        data <- as.data.frame(data)
+        
+        # convert from wide to long
+        data <- data %>%
+          gather(key = "variable", value = "value") %>%
+          distinct() %>%
+          filter(!stringr::str_detect(variable, "attrs$"))
+        
+        simple_vars <- data %>%
[…]

### 2025-12-19 Scripts/Read_XMLs.R

@@ -1,0 +1,546 @@
+
+# ============================================================
+#  IRS 990 XML -> Firm-level + People (name + title only)
+#  Schema-agnostic XML + robust text fallback
+#  Foreign address handling + Option A filtering
+#  Parallel processing (furrr) + diagnostics
+#  Recursively scans nested folders under nonprofit_xmls
+# ============================================================
+
+# ---------------------------
+# CONFIG
+# ---------------------------
+INPUT_DIR      <- "./data/nonprofit_xml"       # top-level folder with nested subfolders
+FIRMS_OUT_CSV  <- "firms_990.csv"
+PEOPLE_OUT_CSV <- "people_990.csv"       # ONLY person_name + title (+ ein, year keys)
+DIAG_OUT_CSV   <- "parse_diagnostics.csv"
+ERRORS_OUT_CSV <- "parse_errors.csv"
+
+WORKERS <- max(1L, parallel::detectCores() - 1L)
+
+# ---------------------------
+# Packages
+# ---------------------------
+# install.packages(c("xml2","dplyr","purrr","stringr","tidyr","janitor","readr","furrr","future"))
+library(xml2)
+library(dplyr)
+library(purrr)
+library(stringr)
+library(tidyr)
+library(janitor)
+library(readr)
+library(furrr)
+library(future)
+
+# ---------------------------
+# Helpers
+# ---------------------------
+
+`%||%` <- function(x, y) {
+  if (is.null(x) || length(x) == 0 || (length(x) == 1 && is.na(x))) y else x
+}
+
+xml_text_or_na <- function(doc_or_node, xpath) {
+  node <- xml2::xml_find_first(doc_or_node, xpath)
+  if (inherits(node, "xml_missing")) return(NA_character_)
+  val <- xml2::xml_text(node)
+  if (length(val) == 0) NA_character_ else stringr::str_squish(val)
+}
+
+find_all <- function(doc_or_node, xpath) xml2::xml_find_all(doc_or_node, xpath)
+strip_ns <- function(doc) { xml2::xml_ns_strip(doc); doc }
+
+# Normalize raw text before regex/line parsing
[…]

### 2025-12-19 Scripts/analysis_new.R

@@ -1,0 +1,551 @@
+source("./Scripts/paths.R")
+
+library(tidyr)
+library(dplyr)
+library(purrr)
+library(fixest)
+library(ggplot2)
+library(did)
+library(forcats)
+
+# ----------------------------------------------------------------------------------
+#                                     Helper Functions
+# ----------------------------------------------------------------------------------
+# Pre period is fixed at g-1; Post period is g + post_offset (event time k).
+build_stack <- function(df, treat, pre_offset = -1, post_offset = 0) {
+  stopifnot(is.character(treat), length(treat) == 1)  
+  if (!treat %in% names(df)) stop("`treat` column not found in `df`.")  
+  if (!is.numeric(dplyr::pull(df, .data[[treat]]))) stop("`treat` must be numeric.")  
+  
+  cohort_years <- df %>%    
+    filter(.data[[treat]] > 0) %>%    
+    distinct(.data[[treat]]) %>%    
+    pull() %>%    
+    sort()
+  
+  stacked <- bind_rows(lapply(cohort_years, function(g) {
+    pre_year  <- g + pre_offset       # g-1
+    post_year <- g + post_offset      # g + k
+    
+    df %>%
+      filter(TaxYr %in% c(pre_year, post_year)) %>%
+      mutate(
+        Treated       = as.integer(.data[[treat]] == g),
+        clean_control = (.data[[treat]] == 0 | .data[[treat]] > g),
+        Post          = as.integer(TaxYr == post_year),
+        stack_id      = g
+      ) %>%
+      filter(Treated == 1 | clean_control) %>%
+      group_by(Filer.EIN) %>%
+      # keep units present in both pre and post within this slice
+      filter(n_distinct(Post) == 2) %>%
+      ungroup()
+  }))
+  
+  stacked %>%
+    mutate(
+      id_stack   = interaction(stack_id, Filer.EIN, drop = TRUE, lex.order = TRUE),
+      time_stack = interaction(stack_id, TaxYr,     drop = TRUE, lex.order = TRUE)
+    )
+}
+
+# ------------------ Helper: pooled DiD with dynamic outcome ------------------
+estimate_stacked <- function(stacked_df, outcome_var, cluster_var = "Filer.EIN") {
+  if (nrow(stacked_df) == 0) {
+    return(tibble(att = NA_real_, se = NA_real_, n_units = 0L, n_rows = 0L))
+  }
+  
[…]

### 2025-12-19 Scripts/clean_matched_names_updated.R

@@ -10,6 +10,10 @@ source("./Scripts/paths.R")
 all_people_data <- readRDS(paste0(created_data_path, "/all_people_data_scheduleH.rds"))
 cw <- readRDS(paste0(created_data_path, "/updated_ein_aha_cw.rds"))
 
+new_people_data <- readRDS(paste0(created_data_path, "new_xmls_people_data.rds")) %>%
+  rename(PersonNm = person_name, TitleTxt = title, TaxYr = end_year, Filer.EIN=ein) %>%
+  filter(PersonNm != "NONE" & !is.na(PersonNm))
+
 # only keep observations where EIN is found in the crosswalk
 all_people_data <- all_people_data %>%
   filter(Filer.EIN %in% cw$Filer.EIN)
@@ -35,30 +39,35 @@ observe <- all_people_data %>%
 all_people_data <- all_people_data %>%
   filter(!is.na(TitleTxt))
 
-# tabulate the number of unique EINs for every value of TaxYr
-all_people_data %>%
-  group_by(TaxYr) %>%
-  summarise(n = n_distinct(Filer.EIN))
-
-# Remove 2015 and 2023
-all_people_data <- all_people_data %>%
-  filter(TaxYr!=2015 & TaxYr!=2023)
-
 # create data set with only relevant variables
 people_data <- all_people_data %>%
   select(TaxYr, Filer.EIN, PersonNm, TitleTxt)
 
+# join to updated people_data from XMLs that I webscraped
+people_data <- people_data %>%
+  rbind(new_people_data)
+
 # convert to all lowercase
 people_data <- people_data %>%
   mutate(PersonNm = tolower(PersonNm),
-         TitleTxt = tolower(TitleTxt))
+         TitleTxt = tolower(TitleTxt)) %>%
+  distinct()
+
+# tabulate the number of unique EINs for every value of TaxYr
+people_data %>%
+  group_by(TaxYr) %>%
+  summarise(n = n_distinct(Filer.EIN))
+
+# Remove 2024
+people_data <- people_data %>%
+  filter(TaxYr!=2024)
 
 # GOAL 1: GET ANY RELEVANT TITLES FROM THE NAME COLUMN #################
 # I got these lists by going through all of the names with at least 3 words. Don't do this again unless you add more names!!
 doctor_list <- c("md", "do", "urologist", "mdmba")
[…]

### 2025-12-19 Scripts/create_connections_boardandexec.R

@@ -264,7 +264,8 @@ identify_common_members <- function(data) {
 # Clean up the names ----------------------------------------------------------------------------------------
 
 # Read in cleaned up version of names found in 990 tax forms
-people <- read_rds(paste0(created_data_path, "cleaned_people_data.rds"))
+people <- read_rds(paste0(created_data_path, "cleaned_people_data_new.rds")) %>%
+  filter(TaxYr!=2025)
 
 # map nicknames to longer versions
 people <- normalize_nicknames(people, nickname_map)
@@ -307,20 +308,10 @@ people <- people %>%
 # assign unique identifiers to names that share at least 2 tokens across firm, years
 people <- assign_person_id(people)
 
+saveRDS(people, paste0(created_data_path, "people.rds"))
+
 # Create variables of board members connected to multiple EINs ---------------------------------------------------
 connections <- identify_common_members(people)
-
-# only keep EINs that are present in enough years
-ein_keep <- connections %>%
-  distinct(Filer.EIN, TaxYr) %>%
-  mutate(count=ifelse(TaxYr %in% c(2017,2018,2019,2020,2021),1,0)) %>%
-  group_by(Filer.EIN) %>%
-  filter(sum(count)>=5) %>%
-  ungroup() %>%
-  distinct(Filer.EIN)
-
-connections <- connections %>%
-  filter(Filer.EIN %in% unique(ein_keep$Filer.EIN))
 
 # pivot longer
 connections <- connections %>%
@@ -385,28 +376,28 @@ connections <- connections %>%
 hospital_data <- connections %>%
   distinct(TaxYr, Filer.EIN) %>%
   left_join(cw) %>%
-  rename(Filer.ID=ID)
+  rename(Filer.ID=ID) 
 
 # Get rid of duplicates
 connections <- connections %>%
-  distinct()
-
+  select(-other) %>%
+  distinct() 
 
 # join AHA information to the filer and other id
-AHA <- read_csv(paste0(raw_data_path, "\\AHAdata_20052023.csv")) 
+AHA <- read_csv(paste0(raw_data_path, "/AHAdata_20052023.csv")) 
 
 AHA_hrr <- AHA %>%
[…]

### 2025-12-19 Scripts/descriptive_tables_and_figures.R

@@ -20,32 +20,68 @@ options(knitr.kable.NA = '')
 people_data <- readRDS(paste0(created_data_path, "people_connections_boardandexec.rds"))
 hospital_data <- readRDS(paste0(created_data_path, "hospital_data_boardandexec.rds"))
 hospital_connections <- readRDS(paste0(created_data_path, "hospital_connections_boardandexec.rds"))
-AHA <- read_csv(paste0(raw_data_path, "\\AHAdata_20052023.csv")) 
+AHA <- read_csv(paste0(raw_data_path, "AHAdata_20052023.csv")) 
 
-# filter out hospitals that turn their overlap on and off
+# generic filtering
+hospital_data <- readRDS(paste0(created_data_path, "hospital_data_boardandexec.rds"))
+
+# Filter to either being present in 2014-2021 at least or dropping out of the data
+# this gets rid of hospitals who come in and out of the data due to missing tax records
+hospital_data <- hospital_data %>%
+  filter(any_sameHRR_nonmiss_2014_2021 | dropout_both_ind) %>%
+  group_by(Filer.EIN) %>%
+  fill(minyr_sameHRR_part, minyr_diffHRR_part, maxyr_sameHRR_part, maxyr_diffHRR_part, .direction="downup") %>%
+  ungroup()
+
+# add indicator for the time the hospital closes
+hospital_data <- hospital_data %>%
+  mutate(closure = ifelse(TaxYr>=dropout_year,1,0)) %>%
+  mutate(closure = ifelse(is.na(closure),0,closure))
+
+# convert variables to numeric
+hospital_data <- hospital_data %>%
+  mutate(minyr_sameHRR_part = as.numeric(minyr_sameHRR_part),
+         minyr_diffHRR_part = as.numeric(minyr_diffHRR_part),
+         maxyr_sameHRR_part = as.numeric(maxyr_sameHRR_part),
+         maxyr_diffHRR_part = as.numeric(maxyr_diffHRR_part),
+         minyr_sameHRR_formal = as.numeric(minyr_sameHRR_formal),
+         maxyr_sameHRR_formal = as.numeric(maxyr_sameHRR_formal),
+         Filer.EIN = as.numeric(Filer.EIN),
+         TaxYr = as.numeric(TaxYr))
+
+# create variables for always, become, and never connected in the same and different HRRs
[…]

### 2025-12-19 Paper/paper_2025.tex

@@
 
-   To create a measure of board/executive interlock, I compile data on nonprofit hospital board of directors and executives in years 2017-2022 from publicly available Tax Form 990s.
+   To create a measure of board/executive overlap, I compile data on nonprofit hospital board of directors and executives in years 2013-2023 from publicly available Tax Form 990s.
 All sufficiently large nonprofits must file these forms with the Internal Revenue Service (IRS) each year.
 They contain a section in which nonprofits declare their board, executives, and highest compensated employees.
-This paper focuses only on the board of director identities.
+This paper focuses only on board members and executives.
 I limit to firms that fill out Schedule H of the tax form, an indicator that the firm operates a hospital.
 I extract all names and positions of individuals listed in the section titled ``Officers, Directors, Trustees, Key Employees, and Five Highest Compensated Employees".
-I also record information on the hospital that filed the tax form: the Employee Identification Number (EIN), organization name and location, and other characteristics.
+I also record information on the hospital entity: the Employee Identification Number (EIN), organization name and location, and other characteristics.
 The initial extraction yields 2,096 nonprofit hospitals.
@@
 
-    I record self-reported information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration.
+    I maintain information about the individuals in the data, including whether they have a title for medical doctor, nurse, or having a degree in health administration.
 To clean the names, I remove any titles or credentials, and I convert common nicknames to their longer versions for consistency.
@@
     I present summary statistics of characteristics of the people in the data in Table \ref{tab:board_people}.
-There are just under 37,000 people in the data, 63\% of which are board members and the remaining 37\% are executives.
-Board members are slightly moe likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%.
+There are just under 52,000 people in the data at some point over the course of 10 years, 63\% of which are board members and the remaining 37\% are executives.
+Board members are slightly more likely to be medical doctors than executives, but the fraction of doctor board members is still relatively low at 18\%.
 Nurses are consistent across board and executives, but only 3\% of the sample are nurses.
 A negligible number of people claim to have a health administration degree.
-Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 32\% of board members estimated to be female.
+Executives are slightly more likely to be female according to the name algorithms, with 38\% of executives estimated to be female and 30\% of board members estimated to be female.
 I also include average team composition of hospital boards and executive teams, where these trends largely hold at the hospital level.
-On average, hospital boards contain 7 members and executive teams contain 10 members.
+On average, hospital boards contain 19 members and executive teams contain 14 members.
 
@@
 
-    I create a list of overlapping board member candidates by identifying any names that occur on multiple hospital boards in the same year.
-However, it is very common for hospitals within a system to largely share the same board.
+    I create a list of overlapping board member or executive candidates by identifying any names associated with multiple hospital in the same year.
+However, it is very common for hospitals within a system to largely share the same board members and/or executives.
 This is not the type of variation I seek to investigate in this paper, since hospitals within systems are already heavily affiliated.
-Therefore, I define the relevant type of overlap as someone being on the board of multiple hospitals that are otherwise \textit{unaffiliated}.
+Therefore, I define the relevant type of overlap as someone being on the leadership team of multiple hospitals, where the hospitals are otherwise unaffiliated by system or network.
 Hospitals within systems are still eligible to have overlap with other hospitals that do not belong to the same system.
@@
 I use the market definitions developed by Dartmouth Atlas, Hospital Referral Regions (HRRs), which are meant to capture geographic regions of the average patient's choice set of hospitals.
+From these overlap definitions, I characterize hospitals by the first and last year they are connected in the same HRR, and the first and last year they are connected in a different HRR.
 
@@
 
-    In Figure \ref{fig:connected_percent} I show the prevalence of overlapping board members or executives over time.
-In 2017, approximately 5\% of hospitals have an overlapping board members or executive with another hospital in the same market that is not formally affiliated otherwise.
-That is, these hospital are not under common ownership.
-This percentage increases over time, with 10\% of hospitals sharing a key leader in their market in 2022.
-I also graph the percent of HRR markets where hospitals within that market share board members or executives.
-Approximately 15\% of HRRs are potentially affected by this type of affiliation.
-Sentence here about putting this into perspective relative to hospital consolidation as a whole.
+    In Figure \ref{fig:connected_percent}, I illustrate the prevalence of overlapping board members or executives over time.
+The fraction of hospitals with overlapping leaders rises gradually from 10\% in 2014 to 15\% in 2017, then declines slightly after 2017.
+The peak in 2017 lacks an obvious explanation, but the subsequent decline may reflect growing consolidation in health care, which reduces opportunities to connect with hospitals outside the same system.
+I also plot the fraction of markets that include at least two hospitals with overlapping leaders.
+This measure increases from 15\% in 2014 to nearly 30\% in 2017, then falls to roughly 20\% by 2023.
+For comparison, about 60\% of hospitals in my sample begin as part of a system, and this fraction rises steadily over time.
+Although informal affiliations through overlapping leaders occur less often than formal system membership, they remain a notable feature of the U.S.
+hospital landscape.
+I also present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}.
+The blue lines represent hospitals that share a common board member or executive, and the red dots are all other hospitals.
+Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population.
 
@@
 
-    Next, I present the geographic distribution of these pairs in Figure \ref{fig:connected_maps}.
-The blue lines represent hospitals that share a common board member, and the red dots are all other hospitals.
-Connected hospitals are distributed fairly evenly across the US, apart from relatively few connections on the west coast relative to the population.
-Over time, the connections remain fairly consistent geographically.
-
     \begin{figure}[ht!]
@@
         \caption{Geographic Distribution of Board/Executive Interlock}
-        \includegraphics[width=\textwidth]{Objects/connected_maps.pdf}
+        \includegraphics[width=.85\textwidth]{Objects/connected_maps.pdf}
         \label{fig:connected_maps}
@@
 
-    Hospitals in the sample are either always connected, never connected, or become connected at some point in time.
-I show characteristics of hospitals in these categories in Table \ref{tab:hosp_group_stats}.
+    \subsection{Characteristics of Hospitals with Overlap}
+
+    Hospitals in the sample fall into three categories: always connected, never connected, or connected at some point in time.
+Table \ref{tab:hosp_group_stats} reports characteristics for each group.
+Hospitals that never exhibit leadership overlap, whether within or across markets, differ markedly from connected hospitals, with substantially fewer beds and nurses.
+In contrast, hospitals that become connected during the sample period resemble connected hospitals much more closely.
+Over 300 hospitals gain overlap within the same HRR, and almost 400 gain overlap across different HRRs.
+These groups are not mutually exclusive, so some hospitals experience overlap in both settings.\footnote{I account for this in the estimation.} Nearly 600 hospitals remain connected across different HRRs throughout the sample, making this form of overlap the most common.
 
@@
 
-    I now present characteristics of pairs of hospitals that are connected by board members or executives in Table \ref{hospital_pair_types}.
+    I now present characteristics of connected pairs of hospitals in Table \ref{hospital_pair_types}.
 In the first column, I show pair characteristics of hospitals connected within the same HRR.
@@
 Each hospital in a pair is general or specialty (I only consider adult hospitals for these categories), children or adult, and independent or belonging to a system, all characteristics in the AHA data.
-Additionally, the median bed size of all hospitals in the sample is 101, and I define hospitals as being either small or large depending on whether they fall below or above this median.
-There are typically 1.5 connections in the same HRR and almost 6 connections in different HRRs.
+Additionally, the median bed size of all hospitals in the sample is 119, and I define hospitals as being either small or large depending on whether they fall below or above this median.
+There are typically 1.7 connections in the same HRR and almost 8 connections in different HRRs.
 
@@
 
-    Most pairs, 76-78\% are two general hospitals.
-Thus, for the majority of hospitals that have board/executive interlock in the same HRR, the hospital they share a leader with is a direct competitor.
-I also include characteristics of system versus independent ownership.
-For there is a nearly even spread of the type of ownership of the pair.
-42\% of these pairs are made up of one independent hospital and one system-owned hospital, while 37\% are made up of two independent hospitals and the rest are made up of two hospitals that are both system owned but belong to two different systems.
-These figures are slightly different for connections in different HRRs, where a larger portion of connections, 45\%, are made up of two system-owned hospitals of two different systems.
+    I will first discuss hospital pairs located in the same HRR.
+Most of these pairs, 64\%, consist of two general hospitals, and are thus competitors in terms of typical services provided.
+An additional 12\% of these pairs are between a general and a specialty hospital, and the remaining 1\% are between two specialty hospitals.
+Similarly, three quarters of the pairs in the same HRR are between two adult hospitals.
+While the largest fraction of these pairs, 44\%, is between two independent hospitals, there is also a significant portion, 40\%, that occur between one independent hospital and one system-affiliated hospital.
+Only 16\% of these pairs are between two hospitals belonging to different systems.
+Almost half of the pairs are between one small and one large hospitals, and the average bed difference is over 200, indicating a significant size difference on average.
+Finally, pairs within the same HRR are typically 35 miles apart.
 
-    In both same and different HRR connected pairs, 50\% of the connections are such that one hospital is small in bed size and one is large in bed size.
-The average bed difference is over 200, indicating that the size difference of the average hospital connection is drastic.
-what's the implication of this?
-Finally, connected hospitals in the same HRR are approximately 62 kilometers, or 38 miles, away from the other.
+    Now, I discuss hospitals that share leaders with hospitals in different markets.
+80\% of these pairs are between two general hospitals, and 90\% are between two adult hospitals.
+Thus, it is likely that hospitals are seeking expertise from other hospitals similar to theirs moreso than in the same HRR.
+There is a fairly even spread of whether the hospitals in the pair are independent.
+Still, 50\% of pairs are between one large and one small hospital, and the bed size difference is 238 on average.
+As expected, the hospitals in these pairs are located much farther apart geographically than those in the same HRR.
 
-    \import{Objects}{hosp_outcomes_table.tex}
+    Hospital decisions to become connected are obviously endogenous.
+Thus, when thinking about how overlap affects hospital behavior, the relevant comparison group is important.
+To think about this further, I present trends over time for a select number of outcomes for four different groups of hospitals: (1) those that gain overlap in the same HRR, (2) those that gain overlap in a different HRR, (3) those that are always connected in a different HRR, and (4) those that are never connected.
+I show these trends for outcomes of number of discharges, expenses, and number of services offered by the hospital in Figure \ref{fig:outcome_desc_graph}.
+
+    \begin{figure}[ht!]
+        \centering
+        \caption{Select Outcomes by Overlap Status}
+        \includegraphics[width=.7\textwidth]{Objects/outcome_desc_graph.pdf}
+        \label{fig:outcome_desc_graph}
+    \end{figure}
+
+  These trends show that not all hospitals serve as a suitable comparison group for those that gain a connection within the same HRR.
+In the top panel, which plots the average number of services over time, hospitals that remain unconnected or always connected in a different HRR follow a different trajectory than hospitals that become connected.
+A similar pattern appears for discharges: hospitals that gain overlap track a comparable path before the connection, then diverge, while hospitals that never gain overlap exhibit a distinct trend of declining Medicare and Medicaid discharges.
+Finally, total operating expenses reveal entirely different slopes for hospitals that never gain overlap compared to those that do, whether in the same or a different market.
+For this reason, when estimating the effect of gaining overlap in the same market, I compare these hospitals only to those that become connected in a different market.
+This approach allows me to isolate the effect of connecting with a competitor from the broader effect of becoming connected.
 
